### 2020-05-13 FastLinAlg/Paper/FastLinAlgPaper.tex

\documentclass[11pt]{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,fullpage    }
%\usepackage{mathtools}
\RequirePackage[dvipsnames,usenames, table]{xcolor}
\usepackage{hyperref}
\hypersetup{
bookmarks,
bookmarksdepth=3,
bookmarksopen,
bookmarksnumbered,
pdfstartview=FitH,
colorlinks,backref,hyperindex,
linkcolor=Sepia,
anchorcolor=BurntOrange,
citecolor=MidnightBlue,
citecolor=OliveGreen,
filecolor=BlueViolet,
menucolor=Yellow,
urlcolor=OliveGreen
}
\usepackage{xypic}
\usepackage{alltt}
\usepackage{multicol}  
%\usepackage{etex}
\usepackage{xspace}
\usepackage{rotating}
\interfootnotelinepenalty=100000

\usepackage{mabliautoref}
\usepackage{colonequals}
\frenchspacing
\input{kmacros3.sty}
\usepackage{stmaryrd}

\usepackage{verbatim}
\usepackage{enumerate}
\begin{document}
\title{{FastLinAlg} package for \emph{Macaulay2}}
\author{Boyana Martinova}
\email{u1056124@utah.edu}
\author{Marcus Robinson}
\email{robinson@math.utah.edu}
\author{Karl Schwede}
\email{schwede@math.utah.edu}
\author{Yuhui Yao}
\email{ywei018@gmail.com}
\address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
\date{\today}

\begin{abstract}
In this article, we present {\tt FastLinAlg.m2}, a package in \emph{Macaulay2} designed to
 introduce new methods focused on computations in function field linear algebra. 
Some key functionality that our 
package offers includes: finding a submatrix of a given rank in a provided matrix (when present), 
verifying that a ring is regular in codimension n, recursively computing the ideals of minors in a 
matrix, and finding an upper bound of the projective dimension of a module.
\end{abstract}

\keywords{FastLinAlg, Macaulay2}
\thanks{Martinova was supported by a University of Utah Mathematics REU fellowship and by the University of Utah ACCESS program.  
Robinson was supported by NSF RTG grant \#1840190.
Schwede was supported by NSF CAREER grant \#1501102 and NSF grant \#1801849.  
Yao was supported by a University of Utah Mathematics REU fellowship.}

\maketitle

\section{Introduction}

We start with some motivation.  
Suppose that $I = (f_1, \dots, f_m) \subseteq k[x_1, \dots, x_n]$ is a prime ideal.  The corresponding variety $X := V(I)$ is singular if and only if $I$ plus the ideal generated by the determinants of the $n-\dim X$ submatrices of the Jacobian matrix 
\[
\mathrm{Jac}(X) = \left( \begin{array}{c} {\partial f_i \over \partial x_j}\end{array}\right),
\]
generate the unit ideal.  
Unfortunately, even for relatively small values of $m$ and $n$, the number of such submatrices is prohibitive.  Suppose for instance that $n = 10, m = 15$ and $\dim X = 5$.  Then there are  
\[
  {10 \choose 5} \cdot {15 \choose 5} =  756756
\]
such submatrices.

We cannot reasonably compute all of their determinants (this computation is often time intensive).  This package attempts to fix this in several ways.  
\begin{enumerate}
  \item We try to compute just a portion of the determinants, in a relatively smart way.
  \item We offer some tools for computing determinants which are sometimes faster.
\end{enumerate}
Of course, computing the singular locus is not the only potential application. This technique has been applied to the related problem of showing that the singular locus has a certain codimension (for example, checking that a variety is R1 in order to prove normality).  We provide a function for giving a better upper bound on projective dimension of a non-homogeneous module.  Finally, this package has also been applied in the {\tt RationalMaps} Macaulay2 package.

We provide the following functions:

\begin{itemize}
  \item{} {\tt getSubmatrixOfRank}, this tries to find a submatrix of a given rank, \autoref{sec.GetSubmatrixOfRank}.
  \item{} {{\tt isRankAtLeast}}, this uses {\tt getSubmatrixOfRank} to try to find lower bounds for the rank of a matrix, \autoref{sec.IsRankAtLeast}.
  \item{}  {\tt Rn}, tries to verify if an integral domain is regular in dimension n, \autoref{sec.Rn}.
  \item{}  {\tt projDim}, tries to find upper bounds for the projective dimension of a non-homogeneous module, 
  \autoref{sec.ProjDim}.
  \item{}  {\tt recursiveMinors}, computes the ideal of minors of a matrix via a recursive cofactor algorithm, as opposed to the included non-recursive cofactor algorithm, \autoref{sec.RecursiveMinors}.
\end{itemize}

The latest version of this package is available at:
\begin{center}
  {\tt https://github.com/kschwede/M2/blob/master/M2/Macaulay2/packages/FastLinAlg.m2}
\end{center}

%We include a final discussion of applications of this package to the {\tt RationalMaps} and {\tt SymbolicPowers} packages.

\subsection*{Acknowledgements:}  The authors thank David Eisenbud, Eloisa Grifo and Zhuang He for valuable conversations.

\section{Finding interesting submatrices}
\label{sec.FindingInterestingSubmatrices}

A lot of the speedups available in the package come down to finding interesting square submatrices of a given matrix. For example, it is often useful to compute a square submatrix whose determinant has small degree as this is less likely to vanish.

\subsection{How are the submatrices chosen?}

Consider the following matrix defined over $\mathbb{Q}[x,y]$.  
\[
  \left[\begin{array}{ccc}
    x & xy & 0 \\
    xy^2 & x^6 & 0 \\
    0 & x^2 y^3 & xy^4
  \end{array} \right]
\]
Suppose we want to choose a submatrix of size $2 \times 2$.  
Consider the monomial order {\tt Lex} where $x < y$.  We find, in the matrix, the nonzero element of smallest order.  
In this case, that is $x$.  We choose this element to be in part of our submatrix.
\[
  \left[\begin{array}{>{\columncolor{red!20}}ccc}
    \rowcolor{red!20}
    \cellcolor{red!40}{\bf x} & xy & 0 \\
    xy^2 & x^6 & 0 \\
    0 & x^2 y^3 & xy^4
  \end{array}\right]
\]
Hence we discard that row and column containing this term and find that the next smallest 
element with respect to our monomial order is $xy^4$.
\[
  \left[\begin{array}{cc}
    x^6 & 0 \\
    x^2 y^3 & x y^4
  \end{array}\right]
  \;\;\;\;
  \left[\begin{array}{c>{\columncolor{red!20}}c}    
    x^6 & 0 \\    
    \rowcolor{red!20}
    x^2 y^3 & \cellcolor{red!40} {\bf x y^4}
  \end{array}\right]
\]
Since we are only looking for a $2 \times 2$ submatrix, we stop here. 
We have selected the submatrix with rows $0$ and $2$ and columns $0$ and $2$. 
\[
  \left[\begin{array}{ccc}
    x &  0 \\
    0 & x y^4 \\
  \end{array} \right]
\]
The determinant of which is $x^2 y ^4$.  This happens to be the smallest $2 \times 2$ minor with respect to the given monomial order (which frequently happens, although it is certainly not always the case).

If we chose a different monomial order, we get a different submatrix, with a different determinant.

For example, 
\begin{description}  
  \item[{\tt Lex}, $x > y$] We yield the submatrix with rows $0$ and $1$ and columns $0$ and $1$, whose determinant is $x^7 - x y^3$
  \item[{\tt GRevLex}, $x < y$] We yield the submatrix with rows $0$ and $2$ and columns $0$ and $1$, whose determinant is $x^3 y^3$
\end{description}
For any of these strategies, in this package we randomize the order of the variables before choosing a submatrix.

The strategies we implement in this package are a bit more complicated, however.  
If we have a matrix whose entries are not monomial, then we could reasonably either pick the 
submatrix of smallest entries with respect to our monomial order ({\tt LexSmallest} or {\tt GrevLexSmallest}), or the submatrix whose 
entries have the smallest terms ({\tt LexSmallestTerm} or {\tt GRevLexSmallestTerm}).  Both of these strategies are implemented.

For example, consider the matrix
\[
  \left[\begin{array}{ccc}
    x^2 + y^2 & 0 & xy + 2x \\
    y^4 - x & 0 & 3 x^5 \\
    x^3 & x^4 y^5  - y^8 & 0
  \end{array} \right]
\]
In this case, if we are choosing the entries with smallest terms, we first replace each entry in the matrix 
with the smallest term.  For example, if we are using {\tt Lex} with $x < y$ we would obtain:
\[
  \left[\begin{array}{ccc}
    x^2 & 0 & 2x \\
    -x & 0 & 3 x^5 \\
    x^3 & x^4 y^5 & 0
  \end{array} \right]
\]
Then we proceed as before.  Notice that if there is a tie, it is broken randomly.  We also have the strategies {\tt GRevLexLargest} and {\tt LexLargest} available to the user.

Finally, we have two random strategies available.  {\tt Random} and {\tt RandomNonzero}.  
\begin{description}
  \item[{\tt Random}]  With this strategy, a random submatrix is chosen.
  \item[{\tt RandomNonzero}]  With this strategy, a random nonzero element is chosen in each step following the method used by the other strategies. This guarantees a submatrix where no row or column is zero which can be very useful when dealing with relatively sparse matrices.
\end{description}

\begin{remark}
  Different strategies work differently on different examples.  When you have a non-homogeneous matrix, with some entries that have constant terms, those entries will always be chosen first when picking smallest terms, regardless of the monomial order.  On the other hand, for homogeneous matrices, choosing the smallest term is frequently very effective.
\end{remark}

\subsubsection{Modifying the underlying matrix when using {\tt GRevLex*}}
Finally, when using {\tt GRevLex*} orders, we periodically change the underlying matrix 
by replacing terms of small order with terms of larger order in order to avoid computing the same submatrix.  For example, in the following matrix, after 
a couple iterations, we might replace the $x^2$ term with 
\[x^2 \cdot (\text{a random degree 1 polynomial}). \]
It might look something like the following.
\[
  \left[\begin{array}{ccc}
    x^2 & 0 & xy \\
    y^4 & 0 & x^5 \\
    x^3 & x^4 y^5 & 0
  \end{array} \right]
  \rightarrow
  \left[\begin{array}{ccc}
    x^2 (2x-7y) & 0 & xy \\
    y^4 & 0 & x^5 \\
    x^3 & x^4 y^5 & 0
  \end{array} \right].
\]
This forces the algorithm to make different choices to be made.  After several iterates are done, the matrix is reset again to its original form.

\subsection{The strategy options}  
The core features included in the package allow the user to choose which 
strategy should be used when selecting submatrices. This is done by setting a 
{\tt Strategy} option to one of the following.  
%{\tt GRevLexLargest, GRevLexSmallest,} and {\tt GRevLexSmallestTerm} use graded 
%reverse lexicographical order to find submatrices where each row and column have a large entry, small 
%entry, and entry with small term, respectively. {\tt LexLargest, LexSmallest,} and {\tt LexSmallestTerm} 
%use lexicographical order to select submatrices that fulfill the same criteria. {\tt Random} selects 
%submatrices at random.

%Generally though, most functions will randomly use one of the strategies.  
\begin{itemize}
  \item {\tt StrategyDefault}: This strategy uses  
  {\tt LexSmallest, LexSmallestTerm, GRevLexSmallest, GRevLexSmallestTerm, Random,} and {\tt RandomNonzero} with equal probability.
  \item {\tt StrategyDefaultNonRandom}: This uses  
  {\tt LexSmallest, LexSmallestTerm, GRevLexSmallest, GRevLexSmallestTerm, Random,} and {\tt RandomNonzero} with equal probability.
  \item {\tt StrategyLexSmallest}: chooses 50\% of the submatrices using {\tt LexSmallest} and 50\% 
  using {\tt LexSmallestTerm}.
  \item {\tt StrategyGRevLexSmallest}: chooses 50\% of the submatrices using {\tt GRevLexSmallest} 
  and 50\% using {\tt GRevLexLargest}.
  \item {\tt StrategyRandom}: chooses submatrices by using 50\% {\tt Random} and 50\% {\tt RandonNonzero}.
  \end{itemize}

The user can create their own custom strategy.  One creates a {\tt HashTable} which has the keys 
{\tt LexLargest, LexSmallestTerm, LexSmallest, GRevLexSmallestTerm, GrevLexSmallest, GRevLexLargest, Random, RandomNonzero} each with value an integer (the values need not sum to 100).  If one value is twice the size of another, that strategy will be employed twice as often.  For example, {\tt StrategyDefaultNonRandom} was created by the command:

{{\small\color{blue}
\begin{verbatim}
StrategyDefaultNonRandom = new HashTable from {
    LexLargest => 0,
    LexSmallestTerm => 25,
    LexSmallest => 25,
    GRevLexSmallestTerm => 25,
    GRevLexSmallest => 25,
    GRevLexLargest => 0,
    Random => 0,
    RandomNonzero => 0
};
\end{verbatim}}
}
\section{Find a submatrix of a given rank: {\tt getSubmatrixOfRank}}
\label{sec.GetSubmatrixOfRank}

This method examines the submatrices of an input matrix and attempts to find one of a 
given rank. If a submatrix with the specified rank is found, a list of two lists is returned. 
The first is the list of rows, the second is the list of columns, which describe the desired submatrix of desired rank. 
If no such submatrix is found, the function will return {\tt null}.

\par The option {\tt MaxMinors} allows the user to control how many minors to consider. If left {\tt null}, 
the number considered is based on the size of the matrix. This method will choose the indicated amount of 
minors using one of the strategy options described below. If one of the chosen submatrices has the desired 
rank, the function will terminate and return its rows and columns. This process continues until a submatrix 
is found or {\tt MaxMinors} submatrices have been unsuccessfully checked.  The strategy can be controlled using the {\tt Strategy} option as described above, the default value is {\tt  StrategyDefaultNonRandom}.

\subsection{Examples of {\tt getSubmatrixOfRank}}
In the following example, we first create a $3 \times 4$ rational matrix, $M$. We execute two calls 
to {\tt getSubmatrixOfRank}, the first has no Strategy parameter and the second utilizes 
{\tt StrategyGRevLexSmallest}. Note that these calls return different indices, but both find 
valid rank $3$ submatrices. We then create a larger $9 \times 10$ matrix over the $3$-dimensional 
space of integers modulo $103$. We display the time needed for the {\tt rank} function to return, 
followed by the time elapsed during a call to {\tt getSubmatrixOfRank} when searching for a rank $7$
submatrix. We repeat these calculations on a new matrix with the same parameters, execting the call
to {\tt getSubmatrixOfRank} first, where we search for a rank $9$ submatrix, and then calling 
{\tt rank}. In both trials we found that {\tt getSubmatrixOfRank} significantly outperformed {\tt rank}.

{\small\color{blue}
\begin{verbatim}
  i1 : loadPackage "FastLinAlg";

  i2 : R = QQ[x,y];

  i3 : M = random(R^{2,2,2},R^4)

  o3 = {-2} | x2+2/3xy+9/2y2    3/10x2+2/3xy+1/5y2 2x2+5/3xy+7/5y2    4/3x2+1/3xy+10/9y2 |
       {-2} | 3/2x2+2/3xy+2y2   1/2x2+3/2xy+3/4y2  6x2+5xy+4y2        9/5x2+1/5xy+7/2y2  |
       {-2} | 1/4x2+1/7xy+5/6y2 7/5x2+4xy+4/5y2    10/9x2+3/7xy+5/9y2 5/2x2+xy+7/6y2     |
               3       4
  o3 : Matrix R  <--- R

  i4 : getSubmatrixOfRank(3,M)

  o4 = {{2, 0, 1}, {0, 1, 3}}

  o4 : List

  i5 : getSubmatrixOfRank(3, M, Strategy=>StrategyGRevLexSmallest)

  o5 = {{0, 2, 1}, {1, 2, 0}}

  o5 : List

  i6 : Q = ZZ/103[x,y,z];

  i7 : N = random(Q^{7,7,7,7,7,7,7,8,8},Q^10); 
                9       10
  o7 : Matrix Q  <--- Q

  i8 : elapsedTime rank N;
       -- 17.701 seconds elapsed

  i9 : elapsedTime getSubmatrixOfRank(7,N);
       -- 0.0373561 seconds elapsed

  i10 : O = random(Q^{7,7,7,7,7,7,7,8,8},Q^10);
                9       10
  o10 : Matrix Q  <--- Q

  i11 : elapsedTime getSubmatrixOfRank(9,O);
  -- 0.76357 seconds elapsed

  i12 : elapsedTime rank O;
       -- 14.6581 seconds elapsed
  \end{verbatim}
}
In one of the core examples from the {\tt RationalMaps} package, before using this package a function would look at several thousands of submatrices (randomly) typically before finding a submatrix of the desired rank whereas this package finds one after looking at fewer than half a dozen.  Using this package sped up the computation of that example by more than one order of magnitude, see 
\cite[Page 7]{BottHassanzadehSchwedeSmolkinRationalMaps}, the non-maximal linear rank example.

\section{Finding lower bounds for matrix ranks: {\tt isRankAtLeast}}
\label{sec.IsRankAtLeast}

This method is a direct implementation of {\tt getSubmatrixOfRank}. This function returns a 
boolean indicating whether the rank of an input matrix, $M$, is greater than or equal to an 
input integer, $n$. In order to do so, the function first performs some basic checks 
to ensure a rank of $n$ is possible given $M$'s dimensions, then executes a call to 
{\tt getSubmatrixOfRank}. If {\tt getSubmatrixOfRank} returns a matrix, then this function 
will return true. However, if {\tt getSubmatrixOfRank} does not return a matrix, a conclusive 
answer can not be reached. As such, the method will then evaluate the rank of $M$ and return 
the appropriate boolean value.

\par {\tt isRankAtLeast} is efficient when {\tt getSubmatrixOfRank} returns quickly, however may be 
costly if the results are inconclusive and a rank evaluation is necessary. As such, the 
described implementation is not optimized. In order to lead to time improvements, we developed a 
multithreaded version of this function that concurrrently evaluates the rank of $M$ and invokes 
{\tt getSubmatrixOfRank}. Once a thread has terminated, the method cancels the other and returns 
the appropriate value. During the implementation of this functionality, we discovered that 
\emph{Macaulay2} becomes unstable when cancelling threads and thus do currently not allow users 
to invoke the multithreaded version. However, this functionality is included in the package and can 
be made easily accessible once the stability issue is resolved.

\subsection{Example of {\tt isRankAtLeast}}
The following example first creates and displays a smaller $3 \times 3$ matrix, $M$. We call
{\tt isRankAtLeast} to determine if its rank is at least 3. We then create a larger $9 \times 9$ 
matrix, $N$, and call {\tt isRankAtLeast} to determine if its rank is at least 7. Directly calling 
{\tt rank N} on a matrix of this size would take multiple seconds, whereas {\tt isRankAtLeast} 
returns in a fraction of the time.

{\small\color{blue}
\begin{verbatim} 
  i1 : loadPackage "FastLinAlg";

  i2 : R = QQ[x,y,z];

  i3 : M = random(R^{2,2,2},R^3)

  o3 = {-2} | 7/2x2+2/3xy+y2+2/9xz+7/10yz+3/5z2     x2+5xy+1/2y2+2xz+7/9yz+7/5z2
      {-2} | 8/7x2+1/10xy+y2+2/3xz+3yz+1/2z2       3x2+7/8xy+1/9y2+3xz+10yz+5z2
      {-2} | 1/10x2+5/2xy+3/4y2+2/3xz+3/2yz+10/7z2 3x2+10/7xy+7/2y2+7/10xz+2yz+1/4z2
      ------------------------------------------------------------------------------
      6/7x2+9/7xy+3/10y2+2/5xz+2/5yz+5/4z2 |
      2/9x2+5/4xy+3y2+1/2xz+yz+7/8z2       |
      8/9x2+3/4xy+y2+10xz+9yz+4/5z2        |
              3       3
  o3 : Matrix R  <--- R

  i4 : isRankAtLeast(3,M)

  o4 = true

  ii5 : rank M

  oo5 = 3

  i6 : N = random(R^{6,6,6,6,6,6,6,7,7},R^9);
                9       9
  o6 : Matrix R  <--- R

  i7 : elapsedTime isRankAtLeast(7,N)
  -- 0.0654172 seconds elapsed

  o7 = true

\end{verbatim}
}

\section{Regular in codimension $n$: {\tt Rn}}
\label{sec.Rn}

Using the {\tt getSubmatrixOfRank} routines, we provide a function for checking if a variety is regular in codimension $n$, or $Rn$.  The default strategy is {\tt Strategy=>Default}.

The function {\tt Rn(ZZ, Ring)} returns {\tt true} if it verifies that the ring is regular in codimension $n$.  Note, this only works if the ring is equidimensional, as it is using a Jacobian criterion.  If it cannot make a determination, it returns {\tt null}.  If it ended up computing all minors of the matrix, and it still doesn't have the desired codimension, it will return {\tt false} (note this will likely only occur for small matrices).

\subsection{Example of {\tt Rn}}
We begin with an example of a 3 dimensional ring that is regular in codimension 1, but not in codimension 2.  It is generated by 12 equations in 7 variables.

{\small\color{blue}
\begin{verbatim}
i3 :  T = ZZ/101[x1,x2,x3,x4,x5,x6,x7];

i4 :  I =  ideal(x5*x6-x4*x7,x1*x6-x2*x7,x5^2-x1*x7,x4*x5-x2*x7,x4^2-x2*x6,x1*x4-x2*x5,
x2*x3^3*x5+3*x2*x3^2*x7+8*x2^2*x5+3*x3*x4*x7-8*x4*x7+x6*x7,x1*x3^3*x5+3*x1*x3^2*x7
+8*x1*x2*x5+3*x3*x5*x7-8*x5*x7+x7^2,x2*x3^3*x4+3*x2*x3^2*x6+8*x2^2*x4+3*x3*x4*x6
-8*x4*x6+x6^2,x2^2*x3^3+3*x2*x3^2*x4+8*x2^3+3*x2*x3*x6-8*x2*x6+x4*x6,x1*x2*x3^3
+3*x2*x3^2*x5+8*x1*x2^2+3*x2*x3*x7-8*x2*x7+x4*x7,x1^2*x3^3+3*x1*x3^2*x5+8*x1^2*x2
+3*x1*x3*x7-8*x1*x7+x5*x7);

o4 : Ideal of T

i5 : S = T/I; dim S

o6 = 3

i7 : time Rn(1, S)
     -- used 0.150734 seconds

o7 = true

i8 : time Rn(2, S)
     -- used 2.12777 seconds

i9 : time singularLocus S;
     -- used 8.29746 seconds

i10 : time dim o9
-- used 23.2483 seconds

o10 = 1
\end{verbatim}
}
You can see that the function {\tt Rn} verified that $S$ was regular in codimension 1 in a fraction of a second.  When we ran {\tt Rn(2, S)} nothing was returned, indicating that nothing was found.  Computing the jacobian ideal however took more than 8 seconds and verifying that it had dimension 1 took more than 23 seconds.

\subsection{The {\tt Strategy} option}

Let us look in slightly more detail at the same example but this time using some different strategies.  For instance, you might think that it might be just as effective to choose random matrices, and sometimes it is. 
{\small\color{blue}
\begin{verbatim}
i11 :  time Rn(1, S, Strategy=>StrategyRandom, Verbose=>true)
  Rn: ring dimension =3, there are 17325 possible minors, we will compute up to 324 of them.
  Rn: About to enter loop
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 7, and computed = 7
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 9, and computed = 9
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 11, and computed = 11
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 14, and computed = 14
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 18, and computed = 18
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 24, and computed = 24
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 31, and computed = 31
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 40, and computed = 40
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 52, and computed = 52
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 67, and computed = 67
  Rn:  partial singular locus dimension computed, = 2
  Rn:  Loop step, about to compute dimension.  Submatrices considered: 87, and computed = 87
  Rn:  partial singular locus dimension computed, = 1
  Rn:  Loop completed, submatrices considered = 87, and computed = 87.  
       singular locus dimension appears to be = 1
       -- used 0.758961 seconds

o11 = true
\end{verbatim}
}
You can see that the {\tt StrategyRandom} option looked at 87 submatrices in this particular example.  Note it does not check to see if we have obtained the desired outcome with each submatrix considered, it does this periodically with the space between checks increasing.  The {\tt considered} values on each line tells how many submatrices have been selected.  Computed tells how many were not repeats (this will be very low with a random strategy).

Running {\tt Rn(1, S, Strategy=>StrategyRandom, Verbose=>true)} 50 times yielded:
\begin{enumerate}
  \item 61.3 average number of submatrices matrices considered.
  \item a median value of 40 or 52 sumatrices considered.
  \item a minimum value of 7 submatrices considered (one time).
  \item a maximum value of 248 submatrices considered (one time).
\end{enumerate}
Note 7 is the minimum number of submatrices that will be considered when checking regular in codimension 1 for this variety (the minimum number depends on the dimension).

On the other hand, the default strategy {\tt Rn(1, S, Strategy=>StrategyDefaultNonRandom, Verbose=>true)} run 50 times yields
\begin{enumerate}
  \item 12.1 average number of submatrices considered.
  \item a median value of 7 or 9 submatrices considered.
  \item a minimum value of 7 submatices considered (25 times).
  \item a maximum value of 40 submatrices considered (one time).
\end{enumerate}

In this particular example, {\tt Strategy=>StrategyLexSmallest} appears to be optimal.  Note that larger matrices tend to create even larger disparities.

\subsection{Notes on implementation}

As mentioned above, this function computes minors (based on the passed {\tt Strategy}) option until either it finds that the singular locus has the desired dimension, or until it has considered too many minors.  By default, it considers up to:
\[
  10\cdot (\text{Minimum number of minors needed}) + 8 \cdot \log_{1.3}(\text{possible minors}).
\]
This was simply chosen by experimentation.
Note that if you are trying to show your singular locus has a certain codimension, you will need a minimum number of minors.  The 10 multiplying it is because our default strategy uses multiple strategies, but only one might work well on a given matrix.  You can provide an alternate function of $x = (\text{minors needed})$ and $y = (\text{possible minors})$ passing it to the option {\tt MaxMinors}.  You can also pass {\tt MaxMinors} a number.

These matrices are considered in a loop.  We begin with computing a constant number of minors, by default $2 \cdot (\text{Minimum number of minors needed}) + 3$ and check whether the output has the right dimension.  You can provide a different function of $x = (\text{minors needed})$ via the option {\tt MinMinorsFunction}.
After that, we compute additional minors, checking periodically (based on an exponential function, $1.3^k$ minors considered before the next reset) whether our minors define a subset of the desired desired codimension.  You can provide your own function via the option {\tt CodimCheckFunction}.  If in this loop, a submatrix is considered again, it is not recomputed, but the counter is still increased.

\subsection{Other options}

This function also includes other options including the option {\tt ModP} which handles switching the coefficient field for a field of characteristic $p > 0$ (which you specify with {\tt ModP => p}. )

One can also control how determinants are computed with the {\tt DetStrategy} option, valid values are {\tt Bareiss}, {\tt Cofactor} and {\tt Recursive}.

\section{Projective dimension: {\tt projDim}}
\label{sec.ProjDim}

In April of 2019, it was pointed out in a thread on github
\begin{center}
  https://github.com/Macaulay2/M2/issues/936
\end{center}
that the command {\tt pdim} sometimes provides an incorrect value (an overestimate) for projective dimension for non-homogeneous modules over polynomial rings.  There it was also suggested that this could be addressed by looking at appropriate minors of the matrices in a possibly non-minimal resolution, but that in practice these matrices have too many minors to compute.  We have implemented a function {\tt projDim} that tries to address this by looking at only \emph{some} minors.  Our function does not solve the problem as it also gives only an upper bound on the projective dimension.  However, this upper bound is more frequently correct.

The idea is as follows.  Take a free resolution of a module $M$ over a polynomial ring $R$  
\[
  \xymatrix{
  0 & \ar[l] M  & \ar[l] F_0 & \ar[l]_{d_1} F_1 & \ar[l] \dots & \ar[l]_{d_{n-1}} F_{n-1}  & \ar[l]_{d_n} F_n & \ar[l] 0.  
  }
\] 
Each $d_i$ is given by a matrix.  The term $F_n$ is unnecessary (ie, $d_n$ splits) happens exactly when the $\rank F_n$ minors of $d_n$ generate the unit ideal.  In that case, we know or projective dimension is at most $n-1$.  However, we can continue in this way, we can compute the $(\rank F_{n-1} - \rank F_n)$-minors of $d_{n-1}$, and see if they generate the unit ideal.  Our algorithm of course only computes a subset of those minors.

\subsection{Example of {\tt projDim}}
In the below example, we take a monomial ideal of projective dimension 2, compute a non-homogeneous change of coordinates, and observe that {\tt pdim} returns an incorrect answer that {\tt projDim} corrects.

{\small\color{blue}
\begin{verbatim}
i1 : R = QQ[x,y,z,w];

i2 : I = ideal(x^4,x*y,w^3, y^4);

i3 : pdim module I

o3 = 2

i4 : f = map(R, R, {x+x^2+1, x+y+1, z+z^4+x-2, w+w^5+y+1});

i5 : pdim module f I

o5 = 3

i6 : time projDim module f I
    -- used 3.43851 seconds

o6 = 2

i7 : time projDim(module f I, MinDimension=>2)
    -- used 0.0503165 seconds

o7 = 2
\end{verbatim}
}

\subsection{Options}

As you can see in the previous example, setting {\tt MinDimension} will can substantially speed up the computation as it won't try to determine if the projective dimension is actually $1$.

The option {\tt MaxMinors} can either be set to be a number (the number of a minors computed at each step).  Or it can be set to be a list of numbers (one for each step in the above algorithm).  Finally, it can be set to be a function of the dimension $d$ of the polynomial ring $R$ and the number $t$ of possible minors.  This is the default option, and the function is:  $5*d + 2*\log_{1.3}(t)$.
The option {\tt Strategy} is also available and it works as above with the default value being {\tt StrategyDefault}.

\section{Computing ideals of minors: {\tt recursiveMinors}}
\label{sec.RecursiveMinors}
\emph{Macaulay2} is a unique because it contains a {\tt minors} method that returns the ideal 
of minors of a certain size, $n$,  in a given matrix, a necessary step in 
locating singularities. However, the current implementation’s default is to 
evaluate determinants using the Bareiss algorithm, which is efficient when 
the entries in the matrix have a low degree and few variables, but very slow 
otherwise. The current minors method also allows users to compute determinants 
using cofactor expansion, but this strategy performs some unnecessary calculations, 
causing it to be quite costly as well. We improved the current cofactor 
expansion method to find the determinants of minors by adding recursion and 
multithreading throughout. We also eliminated said unnecessary calculations by 
ensuring that only the required determinants are being computed at each step 
of the recursion, rather than all possible determinants of the given size.

\par In order to do so, we programmed a method in \emph{Macaulay2}’s software that 
recursively finds all $n \times n$ minors by first computing the $2 \times 2$ minors 
and storing them in a hash table. Then we use the $2 \times 2$ minors to compute the 
necessary $3 \times 3$ minors, and so forth. This process is repeated recursively until 
the minors of size $n \times n$ are evaluated. At each step, we only compute the 
determinants that will be needed when performing a cofactor expansion on the 
following size minor.

\par To allow for further time improvements, we also utilized \emph{Macaulay2}’s existing 
parallel programming methods to multithread our code so different computations 
at each step of the recursion can occur simultaneously in separate threads. We 
divide the list of all determinants to be evaluated into different available 
threads and wait for them to finish before consolidating the results in a hash 
table and proceeding with the recursion. In order to more effectively utilize 
\emph{Macaulay2}'s multithreading methods, we also created a nanosleep method that 
waits a given number of nanoseconds, rather than full seconds. This function has already 
been incorporated into the software.

\subsection{Example of {\tt recursiveMinors}}
\par Below, we first create a simple matrix, M, of one dimensional polynomials with 
rational coefficients and execute the {\tt recursiveMinors} method to find the ideal of all 
$3 \times 3$ minors. As can be seen, the result is equivalent to the output of the minors method 
when called with the same parameters. We then create a new, larger matrix, $N$, with two
dimensional rational coefficients and return the computation time for {\tt recursiveMinors} and 
minors utilizing both the Bareiss and Cofactor strategies. The {\tt recursiveMinors} method 
finished executing approximately six times faster than the Bareiss algorithm and almost seven
times faster than the Cofactor expansion, while yielding the same results.

{\small\color{blue}
\begin{verbatim}
  i1 : loadPackage "FastLinAlg";

  i2 : allowableThreads = 8

  i3 : R = QQ[x];

  i4 : M = random(R^{2,2,2}, R^4)

  o4 = {-2} | x2    3x2   5/8x2 7/10x2 |
       {-2} | 3/4x2 2x2   7/4x2 9x2    |
       {-2} | x2    2/9x2 1/2x2 4/3x2  |
               3       4
  o4 : Matrix R  <--- R

  i5 : recursiveMinors(3,M)

              1403 6  449 6    292 6  517 6
  o5 = ideal (----x , ---x , - ---x , ---x )
               60     240       45    144

  o5 : Ideal of R

  i6 : recursiveMinors(3,M) == minors(3,M)

  o6 = true

  i7 : Q = QQ[x,y];

  i8 : N = random(Q^{5,5,5,5,5,5}, Q^7);
                6       7
  o8 : Matrix Q  <--- Q

  i9 : elapsedTime minors(5,N, Strategy => Bareiss);
  -- 3.0094 seconds elapsed

  o9 : Ideal of Q

  i10 : elapsedTime minors(5,N, Strategy => Cofactor);
  -- 3.76846 seconds elapsed

  o10 : Ideal of Q

  i11 : elapsedTime recursiveMinors(5,N);
  -- 0.590152 seconds elapsed;

  o11 : Ideal of Q

  i12 : recursiveMinors(5,N) == minors(5,N)

  o12 = true

\end{verbatim}
}

\bibliographystyle{skalpha}
\bibliography{MainBib}

\end{document}

### 2020-10-02 FastLinAlg/Paper/RandomRationalPoints.tex
Paper Added

\documentclass[11pt]{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,fullpage    }
%\usepackage{mathtools}
\RequirePackage[dvipsnames,usenames, table]{xcolor}
\usepackage{hyperref,cleveref}
\hypersetup{
bookmarks,
bookmarksdepth=3,
bookmarksopen,
bookmarksnumbered,
pdfstartview=FitH,
colorlinks,backref,hyperindex,
linkcolor=Sepia,
anchorcolor=BurntOrange,
citecolor=MidnightBlue,
citecolor=OliveGreen,
filecolor=BlueViolet,
menucolor=Yellow,
urlcolor=OliveGreen
}
\usepackage{xypic}
\usepackage{alltt}
\usepackage{multicol}  
%\usepackage{etex}
\usepackage{xspace}
\usepackage{rotating}
\interfootnotelinepenalty=100000
\theoremstyle{definition}
\newtheorem{example}{Example}[section]

%\usepackage{mabliautoref}
%\usepackage{colonequals}
%\frenchspacing
%\input{kmacros3.sty}
%\usepackage{stmaryrd}
%
\usepackage{verbatim}
\usepackage{enumerate}
\begin{document}
\title{{RandomRationalPoints} package for \emph{Macaulay2}}
\author{Sankhaneel Bisui}
\email{sbisui@tulane.edu}
\address{Tulane University}
\author{Zhan Jiang}
\email{zoeng@umich.edu}
\address{University of Michigan}
\author{Sarasij Maitra}
\email{sm3vg@virginia.edu}
\address{University of Virginia}
\author{Thai Nguyen}
\email{tnguyen11@tulane.edu}
\address{Tulane University}
\author{Karl Schwede}
\email{schwede@math.utah.edu}
\address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
\date{\today}

\begin{abstract}
In this article, we present {\tt RandomRationalPoints.m2}, a package in \emph{Macaulay2} designed mainly to generate points in a variety over a finite field. We use different strategies like linear intersection and projection to obtain such points. We also present methods to obtain non-vanishing minors of some given size in a given matrix at some chosen point, which in conjunction with the FastLinAlg package should aid in determining singular locus.
\end{abstract}

\keywords{RandomRationalPoints, Macaulay2}
%\thanks{Martinova was supported by a University of Utah Mathematics REU fellowship and by the University of Utah ACCESS program.  
%%Yao was supported by a University of Utah Mathematics REU fellowship.}

\maketitle

\section{Introduction}
 Let $I$ be an ideal in a polynomial ring $k[x_1,\dots, x_n]$ over a finite field $k$. Let $X:=V(I)$ denote the corresponding set of rational points in affine $n$-space. Getting hold of these points, or at least find one such rational point, in an algorithmically efficient manner is our primary motivation. There is an existing package called RandomPoints which has some tools to aid this investigation. Our package provides further machinery along these lines. We provide some functions that apply different strategies to generate random rational points for the given variety. We also provide functions that will expedite the process of determining singular locus together with the functions provided in the package FastLinAlg.

 We provide the following core functions:
 \begin{itemize}
 	\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions})
 	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
    \item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by findANonZeroMinor. (\Cref{extendIdealByNonZeroMinor})

 \end{itemize}

All polynomial rings considered here will be over finite fields. We first briefly describe some helper functions that will be used majorly in the core functions. In the subsequent sections, we explain briefly the core functions. We shall also briefly mention the strategies that we have implemented in the execution of both helper and core functions.

\vspace{1em}
\noindent \textbf{Acknowledgements:} The authors would like to thank David Eisenbud and Mike Stillman for useful conversations and comments on the development of this package.  The authors began work on this package at the virtual Cleveland 2020 Macaulay2 workshop.

\section{helper functions}{\label{helper}}

All polynomial rings that we will consider will be over finite fields only.

We start by describing some functions that will come in handy for the main functions of this package.

\begin{itemize}
	\item {\tt Codimension}: It takes input an integer. Various functions need to know the codimension/height of the scheme/ideal it is working with.  Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.  The default value is {\tt null}, in which case the function will compute the codimension.

	\item {\tt randomCoordinateChange}: It takes a polynomial ring as an input and outputs the coordinate change map. 
	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.

	In some applications, a full change of coordinates is not desired, as it might cause code to run slowly, and so a binomial change of coordinates might be preferred.  
	These changes of coordinates can be accomplished with the following options:

	\begin{itemize}
		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial.  Sometimes, generic coordinate changes where all coordinates are modified, can be very slow.  This is a way to mitigate for that.
		This option is passed to {\tt randomCoordinateChange} by other functions that call it.

		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are changed to a general degree $1$ form.  If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications. If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
	\end{itemize}

	  This function already checks whether the map is an isomorphism by computing the Jacobian.
\end{itemize}

\section{genericProjection, projectionToHypersurface}\label{projectionfunctions}

Both of these functions provide customizable projection techniques. We describe them here.

\begin{enumerate}
\item {\tt genericProjection} 
finds a random (somewhat) generic projection of the ring or ideal.
The typical usages are as follows: 
\vspace{1em}

-- {\tt genericProjection(n, I)},

-- {\tt genericProjection(n, R)},

-- {\tt genericProjection(I)},

-- {\tt genericProjection(R)}

\vspace{1em}
\noindent where 
$I$ is an ideal 
in a polynomial ring, 
$R$ can denote a quotient of a polynomial ring and 
$n\in \mathbb{Z}$ is
an integer specifying how many dimensions to drop.

This gives the projection map from $\mathbb{A}^N \mapsto\mathbb{A}^{N-n}$ and the defining ideal of the projection of $V(I)$. If no integer $n$ is provided, then drops one dimension, in other words it treats $n = 1$.

\begin{example}
	\end{example}

Alternately, instead of {\tt I}, we may pass it a quotient ring.  It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.  It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different. This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.  It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.

\begin{example}
	\end{example}

This function makes no attempt to verify that the projection is actually generic with respect to the ideal.

\vspace{1em} 
\item {\tt projectionToHypersurface} creates a projection to a hypersurface. The typical usages are as follows:

\vspace{1em}
-- {\tt projectionToHypersurface I},

-- {\tt projectionToHypersurface R}

\vspace{1em}
\noindent where $I$ is an ideal in a polynomial ring, $R$ is a quotient of a polynomial ring. The output is
a list with two entries, the generic projection map, and the ideal if {\tt I} was provided, or the ring if {\tt R} was provided.

It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.  This can be faster, and can be useful for finding points.

If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it. The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.

\begin{example}
	content...
\end{example}
\end{enumerate}
\section{randomPoints}\label{randomPoints}
{\tt randomPoints} is a function to find rational points in a variety. The typical usages are as follows:

\vspace{1em}
-- {\tt randomPoints(I)},

-- {\tt randomPoints(n, I)}

\vspace{1em}
\noindent where $n$ is a positive integer denoting the number of desired points, and 
$I$ is an ideal inside a polynomial ring.

\subsection{Additional Options}\label{strategydetails}

The user may also choose to provide some additional information depending on the context which may help in faster computations.

\begin{enumerate}
	\vspace{1em}

\item {\tt Strategy => $\cdot$} where the $\cdot$ can be \textit{Default, BruteForce, LinearIntersection, GenericProjection and  HybridProjectionIntersection}.

\begin{itemize}
	\item {\tt BruteForce} simply tries random points and sees if they are on the variety.

	\item {\tt GenericProjection} projects to a hypersurface, via 
	{\tt projectionToHypersurface} and then uses a {\tt BruteForce} strategy.

	\item {\tt LinearIntersection} intersects with an appropriately random linear space.

	\item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersection. Notice that speed, or success, varies depending on the strategy.

	\item {\tt Default} performs a sequence of different strategies, depending on the context.
\end{itemize}

\begin{example}
	content...
\end{example}

\item {\tt ProjectionAttempts => ZZ}:

When calling the Strategy {\tt GenericProjection} or {\tt HybridProjectionIntersection} from {\tt randomPoints}, this option denotes the number of trials before giving up.  This option is also passed to randomPoints by other functions.

\begin{example}
	content...
\end{example}

\item {\tt MaxCoordinatesToReplace => ZZ} (see \Cref{helper})

\vspace{1em}
\item {\tt Codimension => ZZ} (see \Cref{helper})

\vspace{1em}
\item {\tt ExtendField => Boolean}:

Various functions which produce points, or call functions which produce points, may naturally find scheme theoretic points that are not rational over the base field (for example, by intersecting with a random linear space).  Setting {\tt ExtendField => true} will tell the function that such points are valid.  Setting {\tt ExtendField => false} will tell the function ignore such points.  This sometimes can slow computation, and other times can speed it up.  In some cases, points over extended fields may also have better randomness properties for applications.

\vspace{1em}
\item {\tt IntersectionAttempts => ZZ}:

This option is used by {\tt randomPoints} in some strategies to determine the maximum number of attempts to intersect with a linear space when looking for random rational points.  Other functions pass this option through to {\tt randomPoints}.

\vspace{1em}
\item{\tt PointCheckAttempts => ZZ}:

 When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy or {\tt GenericProjection} strategy, this denotes the number of trials for brute force point checking.

 \begin{example}
 	content...
 \end{example}

\vspace{1em}
\item {\tt MaxCoordinatesToTrivialize}:

When calling {\tt randomPoints} and performing an intersection with a linear space, this is the number of defining equations of the linear space of the form $x_i - a_i$.  Having a large number of these will provide faster intersections.

\vspace{1em}
\item {\tt NumThreadsToUse => ZZ}:

When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
\end{enumerate}
%Outputs
%:List
%a list of points in the variety with possible repetitions.
%Description
%Text  
%Gives at most $n$ many point in a variety $V(I)$. 
%Example
%R = ZZ/5[t_1..t_3];
%I = ideal(t_1,t_2+t_3);
%randomPoints(3, I)
%randomPoints(4, I, Strategy => Default)
%randomPoints(4, I, Strategy => GenericProjection)
%randomPoints(4, I, Strategy => LinearIntersection)

\section{findANonZeroMinor, extendIdealByNonZeroMinor}

As mentioned in the introduction, the two functions in this section will provide further tools for computing singular locus, in addition to those available in the package FastLinAlg.

	\subsection{\tt findANonZeroMinor}\label{findANonZeroMinor}: The typical usage of this function is as follows:

	\vspace{1em}
	-- {\tt findANonZeroMinor(n,M,I)}

	\vspace{1em}
	\noindent where $I$ is an ideal
	in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime; $M$ is a matrix
	over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.

	The function outputs the following:

	-- randomly chosen point $P$ in $V(I)$ which it finds using {\tt randomPoints}.

	-- the indexes of the columns of $M$ that stay linearly independent upon plugging $P$ into $M$,

	-- the indices of the linearly independent rows of the matrix extracted from $M$ using (2),

	-- a random $n\times n$ submatrix of $M$ that has full rank at $P$.

The user may also provide the following additional information:

\begin{itemize}
	\item {\tt Strategy => Symbol} to specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).

	\item {\tt Verbose => Boolean}
		 Set the option {\tt Verbose => true} to turn on verbose output.  This may be useful in debugging or in determining why an computation is running slowly.

	\item {\tt Homogeneous => Boolean} (see \Cref{helper})

	\item {\tt MinorPointAttempts => ZZ} :
	how many points to check the rank of the matrix at
\end{itemize}

\begin{example}

\end{example}

\subsection{\tt extendIdealByNonZeroMinor}\label{extendIdealByNonZeroMinor}: The typical usage is

\vspace{1em}
-- {\tt extendIdealByNonZeroMinor(n,M,I)}

\vspace{1em}
\noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using {\tt findANonZeroMinor};  
it extracts the last entry of the output, finds its determinant and
adds it to the ideal $I$, thus extending $I$.

%\vspace{1em}
As again, we can have the following additional options:

\vspace{1em}
-- {\tt Strategy => Symbol}: specify which strategy to use when calling {\tt randomPoints}

-- {\tt Homogeneous => Boolean}: controls if the computations are homogeneous (in calls to {\tt randomPoints})

-- {\tt Verbose => Boolean}: turns on or off verbose output (see \Cref{findANonZeroMinor})

-- {\tt MinorPointAttempts => ZZ}: how many points to check the rank of the matrix at

\begin{example}

\end{example}

One use for this function can be in showing a certain rings are $R1$ (regular in codimension $1$).
Consider the following example which is $R1$ where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.  However, we can use this function to quickly find interesting minors.
\begin{example}

	\end{example}

In this particular example, there tend to be about 5 associated primes when adding the first minor to J, and so one would expect about 5 steps as each minor computed most likely will eliminate one of those primes.

There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package ``FastLinAlg".

%\end{enumerate}

\vspace{2em}
\end{document}

### 2020-10-03 FastLinAlg/Paper/RandomRationalPoints.tex
Fixed typos, also removed reference to Codimension, since it is not a function.

@@
 \begin{abstract}
-In this article, we present {\tt RandomRationalPoints.m2}, a package in \emph{Macaulay2} designed mainly to generate points in a variety over a finite field.
-We use different strategies like linear intersection and projection to obtain such points.
-We also present methods to obtain non-vanishing minors of some given size in a given matrix at some chosen point, which in conjunction with the FastLinAlg package should aid in determining singular locus.
+In this article, we present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
+We provide different strategies like linear intersection and projection to obtain such points.
+We also present methods to obtain non-vanishing minors of some given size in a given matrix, by evaluating the matrix at a point.
+This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
 \end{abstract}
@@
 Let $X:=V(I)$ denote the corresponding set of rational points in affine $n$-space.
-Getting hold of these points, or at least find one such rational point, in an algorithmically efficient manner is our primary motivation.
-There is an existing package called RandomPoints which has some tools to aid this investigation.
-Our package provides further machinery along these lines.
+Finding one such rational point or geometric point (geometric meaning a point in some finite field extension), in an algorithmically efficient manner is our primary motivation.
+
+ While there is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find all the rational points of a variety, our aim here is to find one point quickly.
  We provide some functions that apply different strategies to generate random rational points for the given variety.
-We also provide functions that will expedite the process of determining singular locus together with the functions provided in the package FastLinAlg.
+We also provide functions that will expedite the process of determining properties of the singular locus of $X$ when combined with the functions provided in the package {\tt FastLinAlg}.
 
@@
  \begin{itemize}
+	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
  	\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
- 	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
     \item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
 	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by findANonZeroMinor. (\Cref{extendIdealByNonZeroMinor})	
-
  \end{itemize}
@@
 All polynomial rings considered here will be over finite fields.
-We first briefly describe some helper functions that will be used majorly in the core functions.
+We first briefly describe some helper functions that will be used extensively in the core functions above.
 In the subsequent sections, we explain briefly the core functions.
@@
 
-\section{helper functions}{\label{helper}}
+\section{Helper functions}{\label{helper}}
 
@@
 \begin{itemize}
-	\item {\tt Codimension}: It takes input an integer.
+%	\item {\tt Codimension}: It takes input an integer.
 Various functions need to know the codimension/height of the scheme/ideal it is working with.
@@
 The default value is {\tt null}, in which case the function will compute the codimension.
-
+%%%
+%%Codimension is not a function, it is an option.
+%%%
 	\item {\tt randomCoordinateChange}: It takes a polynomial ring as an input and outputs the coordinate change map. 
@@
 	In some applications, a full change of coordinates is not desired, as it might cause code to run slowly, and so a binomial change of coordinates might be preferred.
-	These changes of coordinates can be accomplished with the following options:
+Alternately, we might only replace some coordinates by monomials.
+	Refinements of {\tt randomCoordinateChange} are controlled with the following options.
 
@@
 		This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
-
 		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are changed to a general degree $1$ form.
@@
 If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
+		\item 
 	\end{itemize}

### 2020-10-03 FastLinAlg/Paper/RandomRationalPoints.tex
Changes throughout the paper.

@@
 
-\section{Helper functions}{\label{helper}}
+\section{A helper function: randomCoordinateChange}{\label{helper}}
 
-All polynomial rings that we will consider will be over finite fields only.
+All polynomial rings that we will consider will be over finite fields.
 
-We start by describing some functions that will come in handy for the main functions of this package.
+We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
 
-\begin{itemize}
 %	\item {\tt Codimension}: It takes input an integer.
@@
 %%%
-	\item {\tt randomCoordinateChange}: It takes a polynomial ring as an input and outputs the coordinate change map. 
-	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.
+\subsection*{\tt randomCoordinateChange}: It takes a polynomial ring as an input and outputs the coordinate change map. 
+	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
 
-	In some applications, a full change of coordinates is not desired, as it might cause code to run slowly, and so a binomial change of coordinates might be preferred.
-Alternately, we might only replace some coordinates by monomials.
+	In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.
+A binomial change of coordinates might be preferred, or we might only replace some coordinates by monomials.
 	Refinements of {\tt randomCoordinateChange} are controlled with the following options.
@@
 	\begin{itemize}
-		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial.
-Sometimes, generic coordinate changes where all coordinates are modified, can be very slow.
-This is a way to mitigate for that.
+		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
 		This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
-		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are changed to a general degree $1$ form.
+		\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).
+		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
 If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
 If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
-		\item 
 	\end{itemize}
 
-	  This function already checks whether the map is an isomorphism by computing the Jacobian.
-\end{itemize}
+	{{\small\color{blue}
+	\begin{verbatim}
+	i3 : R = ZZ/11[x,y,z];
+
+	i4 : randomCoordinateChange(R)
+
+           ZZ
+o4 = map(R,--[x, y, z],{4x + 5y - 5z, 3x - 4y - 3z, 4x})                     
+           11
+
+	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1)
+
+	o5 = | x -x-4y-5z y |
+
+	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1, Homogeneous=>false)
+
+	o6 = | x-3 z-5 -x+3y-4z+2 |
+
+	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1, Replacement=>Binomial)
+
+	o7 = | y x+4z z |
+	\end{verbatim}}
+	}
 
@@
 $n\in \mathbb{Z}$ is
-an integer specifying how many dimensions to drop.
+an integer specifying how many dimensions to drop.  Note, this function makes no attempt to verify that the projection is actually generic with respect to the ideal.
 
@@
 \begin{example}
+	{{\small\color{blue}
+	\begin{verbatim}
+i1 : R=ZZ/5[x,y,z,w];
+
+i2 : I = ideal(x,y^2,w^3+x^2);
+
+i3 : genericProjection(2,I)
+
+            ZZ                                         2          2
+o6 = (map(R,--[z, w],{- x - 2y - z, - y - 2z}), ideal(z  - z*w - w ))                   
+            5
+
+o6 : Sequence
+	\end{verbatim}
+	}}
 \end{example}
@@
 It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.
+
+\begin{example}
+	{{\small\color{blue}
+	\begin{verbatim}
+i1 : R=ZZ/5[x,y,z,w];
+
+i2 : I = ideal(x,y^2,w^3+x^2);
+
+i3 : genericProjection(2,R/I)
+
+                            ZZ                          ZZ
+                            --[z, w]                    --[z, w]                        
+                R           5                           5
+o3 = (map(----------------,--------,{2y - w, - y - z}), --------)                       
+             2   3    2        4                            4
+        (x, y , w  + x )      z                            z
+
+o3 : Sequence
+	\end{verbatim}
+	}}
+\end{example}
+
 This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.
 It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.
-
-\begin{example}
-	\end{example}
-
-This function makes no attempt to verify that the projection is actually generic with respect to the ideal.
 
@@
 \begin{example}
-	content...
+	{{\small\color{blue}
+	\begin{verbatim}
+i2 : R=ZZ/5[x,y,z,w];
+
+i3 : I = ideal(x,y^2,w^3+x^2);
+
+o3 : Ideal of R
+
+i4 :genericProjection(2,I, Replacement=>Binomial)
+
+            ZZ                                   4
+o4 = (map(R,--[z, w],{x + 2z, - 2y + w}), ideal w )                                     
+            5
+
+o4 : Sequence
+\end{verbatim}
+}}
 \end{example}
+
 \end{enumerate}
 \section{randomPoints}\label{randomPoints}
-{\tt randomPoints} is a function to find rational points in a variety. The typical usages are as follows:
+{\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:
 
@@
 \noindent where $n$ is a positive integer denoting the number of desired points, and 
-$I$ is an ideal inside a polynomial ring.
+$I$ is an ideal inside a polynomial ring.  If {\tt n} is omitted, then it is assumed to be 1.
 
@@
 \begin{itemize}
+    \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.    
 	\item {\tt BruteForce} simply tries random points and sees if they are on the variety.
@@
 
-	\item {\tt Default} performs a sequence of different strategies, depending on the context.
 \end{itemize}
@@
 
-\begin{example}
-	content...
-\end{example}
-
 \item {\tt MaxCoordinatesToReplace => ZZ} (see \Cref{helper})
@@
 
-Various functions which produce points, or call functions which produce points, may naturally find scheme theoretic points that are not rational over the base field (for example, by intersecting with a random linear space).
+Intersection with a general linear space will naturally find scheme theoretic points that are not rational over the base field.  
 Setting {\tt ExtendField => true} will tell the function that such points are valid.
 Setting {\tt ExtendField => false} will tell the function ignore such points.
-This sometimes can slow computation, and other times can speed it up.
+This sometimes can slow computation, and other times can substantially speed it up when the variety has few rational points.
 In some cases, points over extended fields may also have better randomness properties for applications.
@@
 This option is used by {\tt randomPoints} in some strategies to determine the maximum number of attempts to intersect with a linear space when looking for random rational points.
-Other functions pass this option through to {\tt randomPoints}.
+%Other functions pass this option through to {\tt randomPoints}.
 
@@
 
-As mentioned in the introduction, the two functions in this section will provide further tools for computing singular locus, in addition to those available in the package FastLinAlg.
+As mentioned in the introduction, the two functions in this section will provide further tools for computing singular locus, in addition to those available in the package {\tt FastLinAlg}.
 
@@
 
--- {\tt MinorPointAttempts => ZZ}: how many points to check the rank of the matrix at
+-- {\tt MinorPointAttempts => ZZ}: the maximum number of points to check the rank of the matrix at
 
@@
 
-There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package ``FastLinAlg".
+There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}.
 

### 2020-11-10 FastLinAlg/Paper/RandomRationalPoints.tex
Examples 4.1 and 4.2 added. (Z.J.)
Various changes throughout.

@@
 \RequirePackage[dvipsnames,usenames, table]{xcolor}
-\usepackage{hyperref,cleveref}
+\usepackage{hyperref}
+\usepackage[capitalize]{cleveref}
 \hypersetup{
@@
 
- While there is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find all the rational points of a variety, our aim here is to find one point quickly.
+ There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety, our aim here is to find one point quickly, even if it is not rational.
  We provide some functions that apply different strategies to generate random rational points for the given variety.
@@
 
-All polynomial rings that we will consider will be over finite fields.
-
-We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
+%We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
 
@@
 %%%
-\subsection*{\tt randomCoordinateChange}: It takes a polynomial ring as an input and outputs the coordinate change map. 
+\subsection*{\tt randomCoordinateChange} This function takes a polynomial ring as an input and outputs the coordinate change map.  
 	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
@@
 	In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.
-A binomial change of coordinates might be preferred, or we might only replace some coordinates by monomials.
+A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.
 	Refinements of {\tt randomCoordinateChange} are controlled with the following options.
@@
 	\begin{itemize}
+		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
+If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
+If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
 		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
@@
 		\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
-		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
-If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
-If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
 	\end{itemize}
@@
 
-\begin{enumerate}
-\item {\tt genericProjection} 
-finds a random (somewhat) generic projection of the ring or ideal.
+%\begin{enumerate}
+\subsection{genericProjection} 
+This function finds a random (somewhat depending on options) generic projection of the ring or ideal.
 The typical usages are as follows: 
@@
 
-This gives the projection map from $\mathbb{A}^N \mapsto\mathbb{A}^{N-n}$ and the defining ideal of the projection of $V(I)$.
-If no integer $n$ is provided, then drops one dimension, in other words it treats $n = 1$.
+This gives the projection map from $\mathbb{A}^N \mapsto\mathbb{A}^{N-n}$ and the defining ideal of the projection of $V(I)$. If no integer $n$ is provided then it acts as if $n = 1$.
 
@@
 	\begin{verbatim}
+
 i1 : R=ZZ/5[x,y,z,w];
@@
 
-\vspace{1em} 
-\item {\tt projectionToHypersurface} creates a projection to a hypersurface. The typical usages are as follows:
-
-\vspace{1em}
--- {\tt projectionToHypersurface I},
-
--- {\tt projectionToHypersurface R}
-
-\vspace{1em}
-\noindent where $I$ is an ideal in a polynomial ring, $R$ is a quotient of a polynomial ring. The output is
-a list with two entries, the generic projection map, and the ideal if {\tt I} was provided, or the ring if {\tt R} was provided.
-
-It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.
-This can be faster, and can be useful for finding points.
-
-If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
 The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.
@@
 	\begin{verbatim}
+
 i2 : R=ZZ/5[x,y,z,w];
@@
 
-\end{enumerate}
+%\vspace{1em} 
+\subsection{\tt projectionToHypersurface} This function creates a projection to a hypersurface. The typical usages are as follows:
+
+\vspace{1em}
+-- {\tt projectionToHypersurface I},
+
+-- {\tt projectionToHypersurface R}
+
+\vspace{1em}
+\noindent where $I$ is an ideal in a polynomial ring, $R$ is a quotient of a polynomial ring. The output is
+a list with two entries, the generic projection map, and the ideal if {\tt I} was provided, or the ring if {\tt R} was provided.
+
+It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.
+This can be faster, and can be useful for finding points.
+If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
+
+%\end{enumerate}
 \section{randomPoints}\label{randomPoints}
@@
 
-\subsection{Additional Options}\label{strategydetails}
+\subsection{Options}\label{strategydetails}
 
-The user may also choose to provide some additional information depending on the context which may help in faster computations.
+The user may also choose to provide some additional information depending on the context which may help in faster computations, or whether a point is found at all.
 
-\begin{enumerate}
-	\vspace{1em}
-
-\item {\tt Strategy => $\cdot$} where the $\cdot$ can be \textit{Default, BruteForce, LinearIntersection, GenericProjection and  HybridProjectionIntersection}.
+\begin{description}		
+	\item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, GenericProjection} or {\tt HybridProjectionIntersection}.
 
@@
 
-	\item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersection. Notice that speed, or success, varies depending on the strategy.
-
+		\item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersection. 		
 	\end{itemize}
 
-\begin{example}
-	content...
+	Notice that speed, or success, varies depending on the strategy.
+
+	\begin{example}\label{BruteForce}
+		{{\small\color{blue}
+		\begin{verbatim}
+
+			i2 : R = ZZ/101[x,y,z];
+
+			i3 : J = ideal(x^3+y^2+1,z^3-x^2-y^2+2);
+
+			o3 : Ideal of R
+
+			i4 : randomPoints(J)
+				-- used 0.0181549 seconds
+
+			o4 = {{32, 37, -30}}
+
+			o4 : List
+
+			i5 : time randomPoints(J,Strategy=>BruteForce)
+				-- used 0.0146932 seconds
+
+			o5 = {}
+
+			o5 : List
+
+			i6 : randomPoints(J,Strategy=>GenericProjection)
+				-- used 0.0146932 seconds
+
+			o6 = {{2, -30, -8}}
+
+			o6 : List
+
+	\end{verbatim}
+	}}
 	\end{example}
 
-\item {\tt ProjectionAttempts => ZZ}:
+	\item[\tt ProjectionAttempts => ZZ]
 
@@
 
-\item {\tt MaxCoordinatesToReplace => ZZ} (see \Cref{helper})
+	\vspace{1em}
+	\item[\tt MaxCoordinatesToReplace => ZZ] (see \Cref{helper})
+	\vspace{1em}
+	\item[\tt Codimension => ZZ] (see \Cref{helper})
 
 	\vspace{1em}
-\item {\tt Codimension => ZZ} (see \Cref{helper})
-
-\vspace{1em}
-\item {\tt ExtendField => Boolean}:
+	\item[\tt ExtendField => Boolean]
 
@@
 	\vspace{1em}
-\item {\tt IntersectionAttempts => ZZ}:
+	\item[\tt IntersectionAttempts => ZZ]
 
@@
 	\vspace{1em}
-\item{\tt PointCheckAttempts => ZZ}:
+	\item[\tt PointCheckAttempts => ZZ]
 
@@
 	\begin{example}
- 	content...
+	We re-compute \cref{BruteForce} this time specifying more attempts.
+		{{\small\color{blue}
+		\begin{verbatim}
+	i1: R = ZZ/101[x,y,z];
+
+	o1 : PolynomialRing
+
+	i2 : J = ideal "x3+y2+1,z3-x2-y2+2";
+
+	o2 : Ideal of R
+
+	i3 : randomPoints(J,Strategy=>BruteForce,PointCheckAttempts=>20000)
+		-- used 0.84679 seconds
+
+	o3 = {{-39, -43, 28}}
+
+	o3 : List
+	\end{verbatim}
+	}}
 	\end{example}
@@
 	\vspace{1em}
-\item {\tt MaxCoordinatesToTrivialize}:
+	\item[\tt MaxCoordinatesToTrivialize]
 
@@
 	\vspace{1em}
-\item {\tt NumThreadsToUse => ZZ}:
+	\item[\tt NumThreadsToUse => ZZ]
 
 	When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
-\end{enumerate}
+\end{description}
 %Outputs
@@
 %randomPoints(4, I, Strategy => LinearIntersection)
+
+\subsection{Comments on performance}
+
+	When working over very small fields, frequently {\tt BruteForce} is most efficient.
+This is not surprising as there may not be many points to check, and a surprisingly large percentage of points can be rational in hypersurfaces and complete intersections.
+However, if the field size is larger, {\tt BruteForce} will perform poorly.
+Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
+	\begin{example}
+		{{\small\color{blue}
+		\begin{verbatim}
+
+			i2 : R = ZZ/7[x_1..x_10]
+
+			o2 = R
+
+			o2 : PolynomialRing
+
+			i3 : I = ideal(random(2, R), random(3,R));
+
+			o3 : Ideal of R
+
+			i4 : time randomPoints(I,Strategy=>BruteForce, PointCheckAttempts=>20000)
+				 -- used 0.0102075 seconds
+
+			o4 = {{-2, -1, 3, -2, -2, 2, 0, -2, -2, -2}}
+
+			o4 : List
+
+			i5 : time randomPoints(I,Strategy=>Default)
+				 -- used 0.085741 seconds
+
+			o5 = {{1, -2, 2, 2, -2, 0, 0, -1, -1, 3}}
+
+			o5 : List
+
+			i6 : S = ZZ/211[x_1..x_10]
+
+			o6 = S
+
+			o6 : PolynomialRing
+
+			i7 : J = ideal(random(2, S), random(3, S));
+
+			o7 : Ideal of S
+
+			i8 : time randomPoints(J,Strategy=>BruteForce, PointCheckAttempts=>2000000)
+				-- used 25.3479 seconds
+
+			o8 = {{-36, 26, -63, -7, -48, 78, -44, -22, 105, 39}}
+
+			i9 : time randomPoints(J,Strategy=>Default)
+			-- used 0.102918 seconds
+
+			o9 = {{0, -42, -53, -33, -65, -27, -105, -37, 99, 64}}
+
+			o9 : List
+
+			i10 : time randomPoints(J,Strategy=>LinearIntersection)
+			-- used 0.0169781 seconds
+
+			o10 = {{-51, -11, -34, -33, -65, 19, -86, 50, -44, -97}}
+
+			o10 : List
+
+			i11 : time randomPoints(J,Strategy=>GenericProjection)
+			-- used 0.364902 seconds
+
+			o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}
+
+			o11 : List
+
+			i12 : time randomPoints(J,Strategy=>LinearIntersection, ExtendField => true)
+     		-- used 0.0612306 seconds
+
+			               5      4      3      2                                      5  
+			o12 = {{54, 14a  - 48a  - 56a  + 56a  - 13a + 48, -101, 56, -42, -27, - 26a  -
+				-------------------------------------------------------------------------
+				   4      3      2
+				43a  - 70a  - 96a  + 50a - 103, -29, 0, -37}}
+
+	\end{verbatim}
+	}}
+	\end{example}
 
@@
 
-\begin{itemize}
-	\item {\tt Strategy => Symbol} to specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
+\begin{description}
+	\item[\tt Strategy => Symbol] to specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
 
-	\item {\tt Verbose => Boolean}
+	\item[\tt Verbose => Boolean]
 		 Set the option {\tt Verbose => true} to turn on verbose output.  This may be useful in debugging or in determining why an computation is running slowly.
 
-	\item {\tt Homogeneous => Boolean} (see \Cref{helper})
+	\item[\tt Homogeneous => Boolean] (see \Cref{helper})
 
-	\item {\tt MinorPointAttempts => ZZ} :
-	how many points to check the rank of the matrix at
-\end{itemize}
+	\item[\tt MinorPointAttempts => ZZ] 
+		This controls how many points at which to check the rank of the matrix.
+\end{description}
 
@@
 As again, we can have the following additional options:
-
-\vspace{1em}
--- {\tt Strategy => Symbol}: specify which strategy to use when calling {\tt randomPoints}
-
--- {\tt Homogeneous => Boolean}: controls if the computations are homogeneous (in calls to {\tt randomPoints})
-
--- {\tt Verbose => Boolean}: turns on or off verbose output (see \Cref{findANonZeroMinor})
-
--- {\tt MinorPointAttempts => ZZ}: the maximum number of points to check the rank of the matrix at
+\begin{description}
+\item[\tt Strategy => Symbol] This specifies which strategy to use when calling {\tt randomPoints}.
+\item[\tt Homogeneous => Boolean] This is passed in calls to {\tt randomPoints}.
+\item[\tt Verbose => Boolean] This turns on or off verbose output (see \Cref{findANonZeroMinor})
+\item[\tt MinorPointAttempts => ZZ] This controls how many points at which to check the rank of the matrix.
+\end{description}
 

### 2020-12-18 FastLinAlg/Paper/RandomRationalPoints.tex
RandomRationalPoints.tex checked for typos

@@
 \item[\tt Homogeneous => Boolean] This is passed in calls to {\tt randomPoints}.
-\item[\tt Verbose => Boolean] This turns on or off verbose output (see \Cref{findANonZeroMinor})
+\item[\tt Verbose => Boolean] This turns on or off verbose output (see \Cref{findANonZeroMinor}).
 \item[\tt MinorPointAttempts => ZZ] This controls how many points at which to check the rank of the matrix.
@@
 
-One use for this function can be in showing a certain rings are $R1$ (regular in codimension $1$).
-Consider the following example which is $R1$ where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
+One use for this function can be in showing that certain rings are $(R_1)$ (regular in codimension $1$).
+Consider the following example which is $(R_1)$ where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
 However, we can use this function to quickly find interesting minors.

### 2020-12-21 FastLinAlg/Paper/RandomRationalPoints.tex
Examples Added, Paper Ready?

@@
 
- There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety, our aim here is to find one point quickly, even if it is not rational.
+ There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.
  We provide some functions that apply different strategies to generate random rational points for the given variety.
@@
 	\begin{itemize}
-		\item {\tt Replacement}: setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
+		\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
 If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
@@
 
+~~
+
 \begin{example}	
+~~
+
 {{\small\color{blue}
@@
 
-\begin{example}
+~~
+
+\begin{example}~~
+
 	{{\small\color{blue}
@@
 
+~~
+
 \begin{example}
+	~~
+
 	{{\small\color{blue}
@@
 \begin{description}		
-	\item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, GenericProjection} or {\tt HybridProjectionIntersection}.
+	\item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection} or {\tt HybridProjectionIntersection}.
 
@@
 
-	Notice that speed, or success, varies depending on the strategy.
+	Notice that the speed, or success, varies depending on the strategy.
 
 	\begin{example}\label{BruteForce}
+		~~
+
 		{{\small\color{blue}
@@
 
-	When calling the Strategy {\tt GenericProjection} or {\tt HybridProjectionIntersection} from {\tt randomPoints}, this option denotes the number of trials before giving up.
-This option is also passed to randomPoints by other functions.
+	When calling the Strategy {\tt GenericProjection} or \\
+	{\tt HybridProjectionIntersection} from {\tt randomPoints}, this option denotes the number of trials before giving up.  This option is also passed to {\tt randomPoints} by other functions.
 
@@
 	\begin{example}
+
 	We re-compute \cref{BruteForce} this time specifying more attempts.
+
 		{{\small\color{blue}
@@
 Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
+
+	~
+
 	\begin{example}
+		~~
+
 		{{\small\color{blue}
@@
 
-	-- the indices of the linearly independent rows of the matrix extracted from $M$ using (2),
+	-- the indices of the linearly independent rows of the matrix extracted from $M$ in the above step,
 
-	-- a random $n\times n$ submatrix of $M$ that has full rank at $P$.
+	-- a random $n\times n$ sub-matrix of $M$ that has full rank at $P$.
 
@@
 	\item[\tt Verbose => Boolean]
-		 Set the option {\tt Verbose => true} to turn on verbose output.  This may be useful in debugging or in determining why an computation is running slowly.
+		 Set the option {\tt Verbose => true} to turn on verbose output.  This may be useful in debugging or in determining why a computation is running slowly.
 
@@
 
-\begin{example}
+\vspace{1em}
 
+\begin{example}~~
+
+	~~
+	{{\small\color{blue}
+		\begin{verbatim}
+		i3 : R = ZZ/5[x,y,z]
+
+		o3 = R
+
+		o3 : PolynomialRing
+
+		i4 : I = ideal(random(3,R)-2, random(2,R))
+
+		              3     2        2    3    2              2        2       2
+		o4 = ideal (2x  - 2x y + 2x*y  + y  + x z - 2x*y*z + y z - 2x*z  + 2y*z 
+		-------------------------------------------------------------------
+		   3                      2
+		- z  - 2, - 2x*y - x*z - z )
+
+		o4 : Ideal of R
+
+		i5 :  M = jacobian(I)
+
+		o5 = {1} | x2+xy+2y2+2xz-2yz-2z2   -2y-z |
+		     {1} | -2x2-xy-2y2-2xz+2yz+2z2 -2x   |
+		     {1} | x2-2xy+y2+xz-yz+2z2     -x-2z |
+
+		             3       2
+		o5 : Matrix R  <--- R
+
+		i6 : findANonZeroMinor(2,M,I, Strategy => GenericProjection)
+
+		o6 = ({-2, 1, 1}, {0, 1}, {0, 1}, {1} | x2+xy+2y2+2xz-2yz-2z2   -2y-z |)
+		{1} | -2x2-xy-2y2-2xz+2yz+2z2 -2x   |
+
+		o6 : Sequence
+		\end{verbatim}
+	}}
 \end{example}
@@
 \vspace{1em}
-\noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using {\tt findANonZeroMinor};  
+\noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using \\{\tt findANonZeroMinor};  
 it extracts the last entry of the output, finds its determinant and
@@
 %\vspace{1em}
-As again, we can have the following additional options:
+As before, we can have the following additional options:
 \begin{description}
@@
 
-\begin{example}
+\vspace{1em}
+\begin{example}~~
 
+	~~
+
+	{{\small\color{blue}
+\begin{verbatim}
+i2 : R = ZZ/5[x,y,z]
+
+o2 = R
+
+o2 : PolynomialRing
+
+i3 : I = ideal(random(3,R)-2, random(2,R))
+
+                3    2       2     3     2              2       2     3
+o3 = ideal (- 2x  + x y - x*y  - 2y  - 2x z + 2x*y*z + y z + y*z  - 2z 
+-------------------------------------------------------------------
+         2            2                 2
+- 2, - 2x  - 2x*y + 2y  + x*z - 2y*z - z )
+
+o3 : Ideal of R
+
+i4 : M = jacobian(I)
+
+o4 = {1} | -x2+2xy-y2+xz+2yz    x-2y+z   |
+     {1} | x2-2xy-y2+2xz+2yz+z2 -2x-y-2z |
+     {1} | -2x2+2xy+y2+2yz-z2   x-2y-2z  |
+
+             3       2
+o4 : Matrix R  <--- R
+
+i5 : extendIdealByNonZeroMinor(2,M,I, Strategy => LinearIntersection)
+
+                3    2       2     3     2              2       2     3
+o5 = ideal (- 2x  + x y - x*y  - 2y  - 2x z + 2x*y*z + y z + y*z  - 2z 
+-------------------------------------------------------------------
+         2            2                 2   3    2        2    3  
+- 2, - 2x  - 2x*y + 2y  + x*z - 2y*z - z , x  + x y + 2x*y  - y  +
+-------------------------------------------------------------------
+  2       2    3
+2x z + y*z  - z )
+
+o5 : Ideal of R
+\end{verbatim}	
+}}	
 \end{example}
@@
 However, we can use this function to quickly find interesting minors.
-\begin{example}
+
+\vspace{1em}
+
+\begin{example}~~
+
+	~~
+	{{\small\color{blue}
+	\begin{verbatim}
+		i2 : T = ZZ/101[x1,x2,x3,x4,x5,x6,x7];
+
+		i3 : I =  ideal(x5*x6-x4*x7,x1*x6-x2*x7,x5^2-x1*x7,x4*x5-x2*x7,x4^2-x2*x6,x1*x4-x2*x5,
+		x2*x3^3*x5+3*x2*x3^2*x7+8*x2^2*x5+3*x3*x4*x7-8*x4*x7+x6*x7,
+		x1*x3^3*x5+3*x1*x3^2*x7+8*x1*x2*x5+3*x3*x5*x7-8*x5*x7+x7^2,
+		x2*x3^3*x4+3*x2*x3^2*x6+8*x2^2*x4+3*x3*x4*x6-8*x4*x6+x6^2,
+		x2^2*x3^3+3*x2*x3^2*x4+8*x2^3+3*x2*x3*x6-8*x2*x6+x4*x6,
+		x1*x2*x3^3+3*x2*x3^2*x5+8*x1*x2^2+3*x2*x3*x7-8*x2*x7+x4*x7,
+		x1^2*x3^3+3*x1*x3^2*x5+8*x1^2*x2+3*x1*x3*x7-8*x1*x7+x5*x7);
+
+		o3 : Ideal of T
+
+		i4 : M = jacobian I;
+
+		             7       12
+		o4 : Matrix T  <--- T
+
+		i5 : i = 0;
+
+		i6 : J = I;
+
+		o6 : Ideal of T
+
+		i7 : elapsedTime(while (i < 10) and dim J > 1 do (
+                                i = i+1; 
+                                J = extendIdealByNonZeroMinor(4, M, J)));
+		-- 0.903328 seconds elapsed
+
+		i8 : dim J
+
+		o8 = 1
+
+		i9 : i
+
+		o9 = 5
+
+	\end{verbatim}
+}}
 

### 2020-12-21 FastLinAlg/Paper/RandomRationalPoints.tex
Some Further Changes,Example 3.4 added

@@
  \begin{itemize}
+		\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
  	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
- 	\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
     \item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
-	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by findANonZeroMinor. (\Cref{extendIdealByNonZeroMinor})	
+	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
  \end{itemize}
@@
 %%%
-\subsection*{\tt randomCoordinateChange} This function takes a polynomial ring as an input and outputs the coordinate change map.  
+\subsection*{\tt randomCoordinateChange:} This function takes a polynomial ring as an input and outputs the coordinate change map.  
 	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
@@
 If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
-		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
+		\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
 		This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
@@
 	\end{itemize}
+
+\vspace{1em}
+
+\begin{example}~~
+
+	~~
 
@@
 	\end{verbatim}}
-	}
+	}}
 
+\end{example}
 \section{genericProjection, projectionToHypersurface}\label{projectionfunctions}
@@
 \subsection{genericProjection} 
-This function finds a random (somewhat depending on options) generic projection of the ring or ideal.
+This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
 The typical usages are as follows: 
@@
 $n\in \mathbb{Z}$ is
-an integer specifying how many dimensions to drop.  Note, this function makes no attempt to verify that the projection is actually generic with respect to the ideal.
+an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.
 
@@
 
-Alternately, instead of {\tt I}, we may pass it a quotient ring.
+Alternately, instead of {$I$}, we may pass it a quotient ring.
 It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.
@@
 
-~~
+\newpage
 
@@
 
+\vspace{1em}
+\begin{example}
+	{{\small\color{blue}
+	\begin{verbatim}
+	i2 : R=ZZ/5[x,y,z];
+
+	i3 : I = ideal(random(3,R)-2, random(2,R))
+
+	               3       2     3            2        2       2       2  
+	o3 = ideal (- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  -
+	-------------------------------------------------------------------
+	        2
+	2x*y - y  + x*z - y*z)
+
+	o3 : Ideal of R
+
+	i4 : projectionToHypersurface(I)
+
+	            ZZ                             4    3      2 2     4
+	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y -
+	            5
+	-------------------------------------------------------------------
+	z))
+
+	o4 : Sequence
+
+	i5 : projectionToHypersurface(R/I)
+
+                           	                    R                    
+	o5 = (map(--------------------------------------------------------------
+	    3       2     3            2        2       2       2     
+	(- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  - 2x
+	-------------------------------------------------------------------
+                                    ZZ                       
+	                                    --[y, z]                 
+	                                    5                       
+	--------------------,----------------------------------------------
+	      2               4 2    3 3       5     6     3       2     3 
+	*y - y  + x*z - y*z) y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  
+	-------------------------------------------------------------------
+                    	       ZZ
+                	           --[y, z]
+	                            5
+	---,{x - z, y}), -------------------------------------------------)
+	                  4 2    3 3       5     6     3       2     3
+	- 1              y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  - 1
+
+	o5 : Sequence
+
+	\end{verbatim}
+}}
+\end{example}
+
 %\end{enumerate}
@@
 \noindent where $n$ is a positive integer denoting the number of desired points, and 
-$I$ is an ideal inside a polynomial ring.  If {\tt n} is omitted, then it is assumed to be 1.
+$I$ is an ideal inside a polynomial ring.  If {$ n$} is omitted, then it is assumed to be 1.
 
@@
 	Notice that the speed, or success, varies depending on the strategy.
+\newpage
 
@@
 	We re-compute \cref{BruteForce} this time specifying more attempts.
+
+	~~
 
@@
 
-	\subsection{\tt findANonZeroMinor}\label{findANonZeroMinor}: The typical usage of this function is as follows:
+	\subsection{\tt findANonZeroMinor:}\label{findANonZeroMinor} The typical usage of this function is as follows:
 
@@
 \begin{description}
-	\item[\tt Strategy => Symbol] to specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
+	\item[\tt Strategy => Symbol] To specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
 
@@
 
-\subsection{\tt extendIdealByNonZeroMinor}\label{extendIdealByNonZeroMinor}: The typical usage is
+\subsection{\tt extendIdealByNonZeroMinor:}\label{extendIdealByNonZeroMinor} The typical usage is
 
@@
 
-In this particular example, there tend to be about 5 associated primes when adding the first minor to J, and so one would expect about 5 steps as each minor computed most likely will eliminate one of those primes.
+In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
 

### 2020-12-22 FastLinAlg/Paper/RandomRationalPoints.tex
Format changes

@@
 \documentclass[11pt]{amsart}
-\usepackage{calc,amssymb,amsthm,amsmath,fullpage    }
-%\usepackage{mathtools}
+\usepackage{calc,amssymb,amsthm,amsmath,fullpage}
+% \usepackage{mathtools}
 \RequirePackage[dvipsnames,usenames, table]{xcolor}
@@
 \usepackage{multicol}  
-%\usepackage{etex}
+% \usepackage{etex}
 \usepackage{xspace}
@@
 
-%\usepackage{mabliautoref}
-%\usepackage{colonequals}
-%\frenchspacing
-%\input{kmacros3.sty}
-%\usepackage{stmaryrd}
+% \usepackage{mabliautoref}
+% \usepackage{colonequals}
+% \frenchspacing
+% \input{kmacros3.sty}
+% \usepackage{stmaryrd}
 % 
@@
 \keywords{RandomRationalPoints, Macaulay2}
-%\thanks{Martinova was supported by a University of Utah Mathematics REU fellowship and by the University of Utah ACCESS program.  
-%%Yao was supported by a University of Utah Mathematics REU fellowship.}
+% \thanks{Martinova was supported by a University of Utah Mathematics REU fellowship and by the University of Utah ACCESS program.  
+%% Yao was supported by a University of Utah Mathematics REU fellowship.}
 
@@
 
-%We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
+% We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
 
@@
 %%%   
-%%Codimension is not a function, it is an option.
+%%   Codimension is not a function, it is an option.
 %%%   
@@
 \begin{verbatim}
-	i3 : R = ZZ/11[x,y,z];
+	i3 : R = ZZ/11[x, y, z];
 
@@
 
-	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1)
+                    ZZ
+o4 : RingMap R <--- --[x, y, z]
+                    11
+
+	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1)
 
@@
 
-	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1, Homogeneous=>false)
+	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Homogeneous => false)
 
@@
 
-	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace=>1, Replacement=>Binomial)
+	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Replacement => Binomial)
 
@@
 
-%\begin{enumerate}
+% \begin{enumerate}
 \subsection{genericProjection} 
@@
 
-i1 : R=ZZ/5[x,y,z,w];
+i1 : R = ZZ/5[x, y, z, w];
 
-i2 : I = ideal(x,y^2,w^3+x^2);
+i2 : I = ideal(x, y^2, w^3 + x^2);
 
-i3 : genericProjection(2,I)
+i3 : genericProjection(2, I)
 
             ZZ                                         2          2
-o6 = (map(R,--[z, w],{- x - 2y - z, - y - 2z}), ideal(z  - z*w - w ))                   
+o3 = (map(R,--[z, w],{- x - 2y - z, - y - 2z}), ideal(z  - z*w - w ))                   
             5
 
-o6 : Sequence
-	\end{verbatim}
+o3 : Sequence
+\end{verbatim}}
     }}
@@
 
-i1 : R=ZZ/5[x,y,z,w];
+i1 : R = ZZ/5[x, y, z, w];
 
-i2 : I = ideal(x,y^2,w^3+x^2);
+i2 : I = ideal(x, y^2, w^3 + x^2);
 
-i3 : genericProjection(2,R/I)
+i3 : genericProjection(2, R/I)
 
@@
 o3 : Sequence
-	\end{verbatim}
+\end{verbatim}}
     }}
@@
 
-\begin{example}
+\begin{example}~~
+
   ~~
@@
 
-i2 : R=ZZ/5[x,y,z,w];
+i2 : R = ZZ/5[x, y, z, w];
 
-i3 : I = ideal(x,y^2,w^3+x^2);
+i3 : I = ideal(x, y^2, w^3 + x^2);
 
@@
 
-i4 :genericProjection(2,I, Replacement=>Binomial)
+i4 :genericProjection(2, I, Replacement => Binomial)
 
@@
 o4 : Sequence
-\end{verbatim}
+\end{verbatim}}
     }}
@@
 
-%\vspace{1em} 
+% \vspace{1em} 
 \subsection{\tt projectionToHypersurface} This function creates a projection to a hypersurface. The typical usages are as follows:
@@
 \vspace{1em}
-\begin{example}
+\begin{example}~~
+
+  ~~
+
   {{\small\color{blue}
 \begin{verbatim}
-	i2 : R=ZZ/5[x,y,z];
+	i2 : R = ZZ/5[x, y, z];
 
-	i3 : I = ideal(random(3,R)-2, random(2,R))
+	i3 : I = ideal(random(3, R) - 2, random(2, R))
 
@@
 
-	\end{verbatim}
+\end{verbatim}}
     }}
@@
 
-%\end{enumerate}
+% \end{enumerate}
 \section{randomPoints}\label{randomPoints}
@@
   Notice that the speed, or success, varies depending on the strategy.
-\newpage
 
@@
 
-			i2 : R = ZZ/101[x,y,z];
+			i2 : R = ZZ/101[x, y, z];
 
-			i3 : J = ideal(x^3+y^2+1,z^3-x^2-y^2+2);
+			i3 : J = ideal(x^3 + y^2 + 1, z^3 - x^2 - y^2 + 2);
 
@@
 
-	\end{verbatim}
+\end{verbatim}}
       }}
@@
 \begin{verbatim}
-	i1: R = ZZ/101[x,y,z];
+	i1: R = ZZ/101[x, y, z];
 
-	o1 : PolynomialRing
-
-	i2 : J = ideal "x3+y2+1,z3-x2-y2+2";
+	i2 : J = ideal "x3 + y2 + 1, z3 - x2 - y2 + 2";
 
@@
 
-	i3 : randomPoints(J,Strategy=>BruteForce,PointCheckAttempts=>20000)
+	i3 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
 		-- used 0.84679 seconds
@@
 	o3 : List
-	\end{verbatim}
+\end{verbatim}}
       }}
@@
 \end{description}
-%Outputs
-%:List
-%a list of points in the variety with possible repetitions.
-%Description
-%Text  
-%Gives at most $n$ many point in a variety $V(I)$. 
-%Example
-%R = ZZ/5[t_1..t_3];
-%I = ideal(t_1,t_2+t_3);
-%randomPoints(3, I)
-%randomPoints(4, I, Strategy => Default)
-%randomPoints(4, I, Strategy => GenericProjection)
-%randomPoints(4, I, Strategy => LinearIntersection)
+% Outputs
+% :List
+% a list of points in the variety with possible repetitions.
+% Description
+% Text  
+% Gives at most $n$ many point in a variety $V(I)$. 
+% Example
+% R = ZZ/5[t_1..t_3];
+% I = ideal(t_1,t_2+t_3);
+% randomPoints(3, I)
+% randomPoints(4, I, Strategy => Default)
+% randomPoints(4, I, Strategy => GenericProjection)
+% randomPoints(4, I, Strategy => LinearIntersection)
 
@@
 
-			i2 : R = ZZ/7[x_1..x_10]
+			i2 : R = ZZ/7[x_1..x_10];
 
-			o2 = R
-
-			o2 : PolynomialRing
-
-			i3 : I = ideal(random(2, R), random(3,R));
+			i3 : I = ideal(random(2, R), random(3, R));
 
@@
 
-			i4 : time randomPoints(I,Strategy=>BruteForce, PointCheckAttempts=>20000)
+			i4 : time randomPoints(I, Strategy => BruteForce, PointCheckAttempts => 20000)
 				 -- used 0.0102075 seconds
@@
 
-			i5 : time randomPoints(I,Strategy=>Default)
+			i5 : time randomPoints(I, Strategy => Default)
 				 -- used 0.085741 seconds
@@
 
-			i6 : S = ZZ/211[x_1..x_10]
-
-			o6 = S
-
-			o6 : PolynomialRing
+			i6 : S = ZZ/211[x_1..x_10];
 
@@
 
-			i8 : time randomPoints(J,Strategy=>BruteForce, PointCheckAttempts=>2000000)
+			i8 : time randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 2000000)
 				-- used 25.3479 seconds
@@
 
-			i9 : time randomPoints(J,Strategy=>Default)
+			i9 : time randomPoints(J, Strategy => Default)
 			-- used 0.102918 seconds
@@
 
-			i10 : time randomPoints(J,Strategy=>LinearIntersection)
+			i10 : time randomPoints(J, Strategy => LinearIntersection)
 			-- used 0.0169781 seconds
@@
 
-			i11 : time randomPoints(J,Strategy=>GenericProjection)
+			i11 : time randomPoints(J, Strategy => GenericProjection)
 			-- used 0.364902 seconds
@@
 
-			i12 : time randomPoints(J,Strategy=>LinearIntersection, ExtendField => true)
+			i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
      		-- used 0.0612306 seconds
@@
 \vspace{1em}
-	-- {\tt findANonZeroMinor(n,M,I)}
+-- {\tt findANonZeroMinor(n, M, I)}
 
@@
 \begin{verbatim}
-		i3 : R = ZZ/5[x,y,z]
+		i3 : R = ZZ/5[x, y, z];
 
-		o3 = R
-
-		o3 : PolynomialRing
-
-		i4 : I = ideal(random(3,R)-2, random(2,R))
+		i4 : I = ideal(random(3, R) - 2, random(2, R))
 
@@
 
-		i6 : findANonZeroMinor(2,M,I, Strategy => GenericProjection)
+		i6 : findANonZeroMinor(2, M, I, Strategy => GenericProjection)
 
@@
 		o6 : Sequence
-		\end{verbatim}
+\end{verbatim}}
     }}
@@
 \vspace{1em}
--- {\tt extendIdealByNonZeroMinor(n,M,I)}
+-- {\tt extendIdealByNonZeroMinor(n, M, I)}
 
@@
 
-%\vspace{1em}
+% \vspace{1em}
 As before, we can have the following additional options:
@@
 \begin{verbatim}
-i2 : R = ZZ/5[x,y,z]
+i2 : R = ZZ/5[x, y, z];
 
-o2 = R
-
-o2 : PolynomialRing
-
-i3 : I = ideal(random(3,R)-2, random(2,R))
+i3 : I = ideal(random(3, R) - 2, random(2, R))
 
@@
 
-i5 : extendIdealByNonZeroMinor(2,M,I, Strategy => LinearIntersection)
+i5 : extendIdealByNonZeroMinor(2, M, I, Strategy => LinearIntersection)
 
@@
 o5 : Ideal of R
-\end{verbatim}	
+\end{verbatim}}	
     }}	
@@
 \begin{verbatim}
-		i2 : T = ZZ/101[x1,x2,x3,x4,x5,x6,x7];
+		i2 : T = ZZ/101[x1, x2, x3, x4, x5, x6, x7];
 
@@
 		i7 : elapsedTime(while (i < 10) and dim J > 1 do (
-                                i = i+1; 
+                                i = i + 1; 
                                 J = extendIdealByNonZeroMinor(4, M, J)));
@@
 
-%\end{enumerate}
+% \end{enumerate}
 

### 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
Updated the paper to make it shorter.

@@
 \newtheorem{example}{Example}[section]
+%\usepackage{enumitem}
+%\SetLabelAlign{parleft}{\parbox[t]\textwidth{#1\smallskip}}
 
@@
 \usepackage{enumerate}
+
 \begin{document}
@@
   In this article, we present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
-We provide different strategies like linear intersection and projection to obtain such points.
+We provide different strategies such as linear intersection and projection to obtain such points.
 We also present methods to obtain non-vanishing minors of some given size in a given matrix, by evaluating the matrix at a point.
-This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
+%This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
 \end{abstract}
@@
     We provide some functions that apply different strategies to generate random rational points for the given variety.
-We also provide functions that will expedite the process of determining properties of the singular locus of $X$ when combined with the functions provided in the package {\tt FastLinAlg}.
+We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.
 
@@
     \begin{itemize}    
+    \item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
     \item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
-\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
-\item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
-\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
+    \item {\tt findANonZeroMinor} and {\tt extendIdealByNonZeroMinor}:  The first function finds a submatrix of a given matrix that is nonsingular at a point of a given ideal.
+The second, adds said submatrix to an ideal, which is useful for computing partial Jacobian ideals.
+(\Cref{findANonZeroMinor})
+    %\item :  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
     \end{itemize}
@@
     All polynomial rings considered here will be over finite fields.
-We first briefly describe some helper functions that will be used extensively in the core functions above.
-In the subsequent sections, we explain briefly the core functions.
-We shall also briefly mention the strategies that we have implemented in the execution of both helper and core functions.
+In the subsequent sections, we explain briefly the core and some helper functions and describe the strategies that we have implemented in their execution.
 
@@
 
-\section{A helper function: randomCoordinateChange}{\label{helper}}
+\section{Our primary pupose: {\tt randomPoints}}\label{randomPoints}
+    We start with the core function in this package:  {\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:
 
-% We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
-
-% \item {\tt Codimension}: It takes input an integer.
-Various functions need to know the codimension/height of the scheme/ideal it is working with.
-Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.
-The default value is {\tt null}, in which case the function will compute the codimension.
-%%%   
-%%   Codimension is not a function, it is an option.
-%%%   
-\subsection*{\tt randomCoordinateChange:} This function takes a polynomial ring as an input and outputs the coordinate change map.  
-i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
-
-In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.
-A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.
-Refinements of {\tt randomCoordinateChange} are controlled with the following options.
-
-\begin{itemize}
-\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
-If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
-If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
-\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
-  This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
-\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
-\end{itemize}
-
-\vspace{1em}
-
-\begin{example}~~
-
-  ~~
-
-  {{\small\color{blue}
-\begin{verbatim}
-	i3 : R = ZZ/11[x, y, z];
-
-	i4 : randomCoordinateChange(R)
-
-           ZZ
-o4 = map(R,--[x, y, z],{4x + 5y - 5z, 3x - 4y - 3z, 4x})                     
-           11
-
-                    ZZ
-o4 : RingMap R <--- --[x, y, z]
-                    11
-
-	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1)
-
-	o5 = | x -x-4y-5z y |
-
-	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Homogeneous => false)
-
-	o6 = | x-3 z-5 -x+3y-4z+2 |
-
-	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Replacement => Binomial)
-
-	o7 = | y x+4z z |
-\end{verbatim}}
-    }}
-
-\end{example}
-\section{genericProjection, projectionToHypersurface}\label{projectionfunctions}
-
-Both of these functions provide customizable projection techniques. We describe them here.
-
-% \begin{enumerate}
-\subsection{genericProjection} 
-This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
-The typical usages are as follows: 
-\vspace{1em}
-
--- {\tt genericProjection(n, I)},
-
--- {\tt genericProjection(n, R)},
-
--- {\tt genericProjection(I)},
-
--- {\tt genericProjection(R)}
-
-\vspace{1em}
-\noindent where 
-$I$ is an ideal 
-in a polynomial ring, 
-$R$ can denote a quotient of a polynomial ring and 
-$n\in \mathbb{Z}$ is
-an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.
-
-This gives the projection map from $\mathbb{A}^N \mapsto\mathbb{A}^{N-n}$ and the defining ideal of the projection of $V(I)$. If no integer $n$ is provided then it acts as if $n = 1$.
-
-~~
-
-\begin{example}	
-  ~~
-
-  {{\small\color{blue}
-\begin{verbatim}
-
-i1 : R = ZZ/5[x, y, z, w];
-
-i2 : I = ideal(x, y^2, w^3 + x^2);
-
-i3 : genericProjection(2, I)
-
-            ZZ                                         2          2
-o3 = (map(R,--[z, w],{- x - 2y - z, - y - 2z}), ideal(z  - z*w - w ))                   
-            5
-
-o3 : Sequence
-\end{verbatim}}
-    }}
-\end{example}
-
-Alternately, instead of {$I$}, we may pass it a quotient ring.
-It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.
-It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.
-
-~~
-
-\begin{example}~~
-
-  {{\small\color{blue}
-\begin{verbatim}
-
-i1 : R = ZZ/5[x, y, z, w];
-
-i2 : I = ideal(x, y^2, w^3 + x^2);
-
-i3 : genericProjection(2, R/I)
-
-                            ZZ                          ZZ
-                            --[z, w]                    --[z, w]                        
-                R           5                           5
-o3 = (map(----------------,--------,{2y - w, - y - z}), --------)                       
-             2   3    2        4                            4
-        (x, y , w  + x )      z                            z
-
-o3 : Sequence
-\end{verbatim}}
-    }}
-\end{example}
-
-This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.
-It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.
-
-The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.
-
-\newpage
-
-\begin{example}~~
-
-  ~~
-
-  {{\small\color{blue}
-\begin{verbatim}
-
-i2 : R = ZZ/5[x, y, z, w];
-
-i3 : I = ideal(x, y^2, w^3 + x^2);
-
-o3 : Ideal of R
-
-i4 :genericProjection(2, I, Replacement => Binomial)
-
-            ZZ                                   4
-o4 = (map(R,--[z, w],{x + 2z, - 2y + w}), ideal w )                                     
-            5
-
-o4 : Sequence
-\end{verbatim}}
-    }}
-\end{example}
-
-% \vspace{1em} 
-\subsection{\tt projectionToHypersurface} This function creates a projection to a hypersurface. The typical usages are as follows:
-
-\vspace{1em}
--- {\tt projectionToHypersurface I},
-
--- {\tt projectionToHypersurface R}
-
-\vspace{1em}
-\noindent where $I$ is an ideal in a polynomial ring, $R$ is a quotient of a polynomial ring. The output is
-a list with two entries, the generic projection map, and the ideal if {\tt I} was provided, or the ring if {\tt R} was provided.
-
-It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.
-This can be faster, and can be useful for finding points.
-If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
-
-\vspace{1em}
-\begin{example}~~
-
-  ~~
-
-  {{\small\color{blue}
-\begin{verbatim}
-	i2 : R = ZZ/5[x, y, z];
-
-	i3 : I = ideal(random(3, R) - 2, random(2, R))
-
-	               3       2     3            2        2       2       2  
-	o3 = ideal (- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  -
-	-------------------------------------------------------------------
-	        2
-	2x*y - y  + x*z - y*z)
-
-	o3 : Ideal of R
-
-	i4 : projectionToHypersurface(I)
-
-	            ZZ                             4    3      2 2     4
-	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y -
-	            5
-	-------------------------------------------------------------------
-	z))
-
-	o4 : Sequence
-
-	i5 : projectionToHypersurface(R/I)
-
-                           	                    R                    
-	o5 = (map(--------------------------------------------------------------
-	    3       2     3            2        2       2       2     
-	(- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  - 2x
-	-------------------------------------------------------------------
-                                    ZZ                       
-	                                    --[y, z]                 
-	                                    5                       
-	--------------------,----------------------------------------------
-	      2               4 2    3 3       5     6     3       2     3 
-	*y - y  + x*z - y*z) y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  
-	-------------------------------------------------------------------
-                    	       ZZ
-                	           --[y, z]
-	                            5
-	---,{x - z, y}), -------------------------------------------------)
-	                  4 2    3 3       5     6     3       2     3
-	- 1              y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  - 1
-
-	o5 : Sequence
-
-\end{verbatim}}
-    }}
-\end{example}
-
-% \end{enumerate}
-\section{randomPoints}\label{randomPoints}
-{\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:
-
-\vspace{1em}
+    \vspace{0.5em}
     -- {\tt randomPoints(I)},
@@
 
-\vspace{1em}
+    \vspace{0.5em}
     \noindent where $n$ is a positive integer denoting the number of desired points, and 
@@
 
-\begin{description}		
+    \begin{description}%[]
+        \setlength{\itemsep}{5pt}
     \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection} or {\tt HybridProjectionIntersection}.
@@
     \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.
+It begins with less general projections and linear intersections, and gradually ramps up the randomness.
     \item {\tt BruteForce} simply tries random points and sees if they are on the variety.
@@
     \begin{example}\label{BruteForce}
+        Consider the following example.
         ~~        
-
-    {{\small\color{blue}
+        {\small\color{blue}
+    \begin{quote}
 \begin{verbatim}
-
 i2 : R = ZZ/101[x, y, z];
@@
 o6 : List
-
-\end{verbatim}}
-      }}
-  \end{example}
-
+        \end{verbatim}
+    \end{quote}
+        }
+    \end{example}%
+    %
+    \vspace{-1em}
     \item[\tt ProjectionAttempts => ZZ]
@@
 
-  \vspace{1em}
-\item[\tt MaxCoordinatesToReplace => ZZ] (see \Cref{helper})
-  \vspace{1em}
-\item[\tt Codimension => ZZ] (see \Cref{helper})
+    \item[\tt MaxCoordinatesToReplace => ZZ] This is used for fine tuning the {\tt LinearIntersection} or {\tt GenericProjection} strategies, see \Cref{helper}.
+    \item[\tt Codimension => ZZ] If the codimension is already known, one can provide it so it is not recomputed.
+This is used in the {\tt GenericProjection} and {\tt LinearIntersection} strategies, see \Cref{projectionfunctions}.
 
-  \vspace{1em}
     \item[\tt ExtendField => Boolean]
@@
 
-  \vspace{1em}
     \item[\tt IntersectionAttempts => ZZ]
@@
 
-  \vspace{1em}
     \item[\tt PointCheckAttempts => ZZ]
 
-  When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy or {\tt GenericProjection} strategy, this denotes the number of trials for brute force point checking.
+    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of trials for brute force point checking.
 
     \begin{example}
-
         We re-compute \cref{BruteForce} this time specifying more attempts.
-
+    {{\small\color{blue}
     ~~
-
-    {{\small\color{blue}
+    \begin{quote}
 \begin{verbatim}    
-	i1: R = ZZ/101[x, y, z];
-
-	i2 : J = ideal "x3 + y2 + 1, z3 - x2 - y2 + 2";
-
-	o2 : Ideal of R
-
-	i3 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
+i7 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
     -- used 0.84679 seconds
 
-	o3 = {{-39, -43, 28}}
-
-	o3 : List
-\end{verbatim}}
-      }}
-  \end{example}
-
-  \vspace{1em}
+o7 = {{-39, -43, 28}}
+    \end{verbatim}%        
+\end{quote}%
+}}%
+    \end{example}%
+    \vspace{-1em}
+%
+    %\vspace{1em}
     \item[\tt MaxCoordinatesToTrivialize]
@@
 
-  \vspace{1em}
+%    \vspace{1em}
     \item[\tt NumThreadsToUse => ZZ]
 
-  When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
+    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
     \end{description}
@@
 Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
-
-~
+A quick comment, while in many examples generic projection works well, there are some where it has very slow performance.
+Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation.
 
     \begin{example}
+        We begin with an example over a small field.
+    {{\small\color{blue}
     ~~
-
-  {{\small\color{blue}
+    \begin{quote}
 \begin{verbatim}
-
 i2 : R = ZZ/7[x_1..x_10];
@@
 o5 : List
+    \end{verbatim}
+\end{quote}\vspace{-1em}
+    }}
+    \end{example}
 
+\begin{example}Now we work over a larger field.
+        {{\small\color{blue}
+        ~~
+        \begin{quote}
+\begin{verbatim}
 i6 : S = ZZ/211[x_1..x_10];
@@
 o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
+    \end{verbatim}
+\end{quote}\vspace{-1em}
+        }}
+    \end{example}
 
-			o11 : List
-
+    \begin{example}
+        Finally, we allow our functions to extend our field.
+        {{\small\color{blue}
+        ~~
+        \begin{quote}
+        \begin{verbatim}
 i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
@@
     43a  - 70a  - 96a  + 50a - 103, -29, 0, -37}}
-
-\end{verbatim}}
+        \end{verbatim}
+    \end{quote}
         }}
@@
 
-\section{findANonZeroMinor, extendIdealByNonZeroMinor}
+\section{A helper function: {\tt randomCoordinateChange}}{\label{helper}}
+
+% We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
+
+% \item {\tt Codimension}: It takes input an integer.
+Various functions need to know the codimension/height of the scheme/ideal it is working with.
+Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.
+The default value is {\tt null}, in which case the function will compute the codimension.
+%%%   
+%%   Codimension is not a function, it is an option.
+%%%   
+\subsection*{\tt randomCoordinateChange:}
+
+This function takes a polynomial ring as an input and outputs the coordinate change map.  
+i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
+
+In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.
+A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.
+This is controlled with the following options.
+
+\begin{itemize}
+\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
+If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
+If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
+\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
+  This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
+\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
+\end{itemize} 
+%\par
+%\vspace{-1em}    
+\begin{example}  We demonstrate some of these options.
+%
+  {{\small\color{blue}
+  \begin{quote}
+\begin{verbatim}
+	i3 : R = ZZ/11[x, y, z];
+
+	i4 : randomCoordinateChange(R)
+
+           ZZ
+o4 = map(R,--[x, y, z],{4x + 5y - 5z, 3x - 4y - 3z, 4x})                     
+           11
+
+                    ZZ
+o4 : RingMap R <--- --[x, y, z]
+                    11
+
+	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1)
+
+	o5 = | x -x-4y-5z y |
+
+	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Homogeneous => false)
+
+	o6 = | x-3 z-5 -x+3y-4z+2 |
+
+	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Replacement => Binomial)
+
+	o7 = | y x+4z z |
+\end{verbatim}
+\end{quote}
+    }}
+
+\end{example}
+\section{Other helper functions: {\tt genericProjection}, {\tt projectionToHypersurface}}\label{projectionfunctions}
+
+Both of these functions provide customizable projection techniques. We describe them here.
+
+% \begin{enumerate}
+\subsection{genericProjection} 
+This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
+The typical usages are as follows: 
+\vspace{0.5em}
+
+-- {\tt genericProjection(n, I)},
+
+-- {\tt genericProjection(n, R)},
+
+-- {\tt genericProjection(I)},
+
+-- {\tt genericProjection(R)}
+
+\vspace{0.5em}
+\noindent where 
+$I$ is an ideal 
+in a polynomial ring, 
+$R$ can denote a quotient of a polynomial ring and 
+$n\in \mathbb{Z}$ is
+an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.
+
+This gives the projection map from $\mathbb{A}^N \mapsto\mathbb{A}^{N-n}$ and the defining ideal of the projection of $V(I)$. If no integer $n$ is provided then it acts as if $n = 1$.
+
+\begin{example}	
+    We project a curve in 4-space to 2-space.
+    ~~
+  {{\small\color{blue}
+  \begin{quote}
+\begin{verbatim}
+i1 : R = ZZ/5[x, y, z, w];
+
+i2 : I = ideal(x, y^2, w^3 + x^2);
+
+i3 : genericProjection(2, I)
+
+            ZZ                                         2          2
+o3 = (map(R,--[z, w],{- x - 2y - z, - y - 2z}), ideal(z  - z*w - w ))                   
+            5
+
+o3 : Sequence
+\end{verbatim}
+\end{quote}
+    }}
+\end{example}
+
+Alternately, instead of {$I$}, we may pass it a quotient ring.
+It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.
+It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.
+
+~~
+
+\begin{comment}~~
+
+  {{\small\color{blue}
+\begin{verbatim}
+
+i1 : R = ZZ/5[x, y, z, w];
+
+i2 : I = ideal(x, y^2, w^3 + x^2);
+
+i3 : genericProjection(2, R/I)
+
+                            ZZ                          ZZ
+                            --[z, w]                    --[z, w]                        
+                R           5                           5
+o3 = (map(----------------,--------,{2y - w, - y - z}), --------)                       
+             2   3    2        4                            4
+        (x, y , w  + x )      z                            z
+
+o3 : Sequence
+\end{verbatim}}
+    }}
+\end{comment}
+
+This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.
+It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.
+%The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.
+
+\begin{comment}~~
+
+  ~~
+
+  {{\small\color{blue}
+\begin{verbatim}
+
+i2 : R = ZZ/5[x, y, z, w];
+
+i3 : I = ideal(x, y^2, w^3 + x^2);
+
+o3 : Ideal of R
+
+i4 :genericProjection(2, I, Replacement => Binomial)
+
+            ZZ                                   4
+o4 = (map(R,--[z, w],{x + 2z, - 2y + w}), ideal w )                                     
+            5
+
+o4 : Sequence
+\end{verbatim}}
+    }}
+\end{comment}
+
+% \vspace{1em} 
+\subsection{\tt projectionToHypersurface} This function creates a projection to a hypersurface. The typical usages are as follows:
+
+\vspace{0.5em}
+-- {\tt projectionToHypersurface I},
+
+-- {\tt projectionToHypersurface R} 
+\vspace{0.5em}
+
+\noindent where $I$ is an ideal in a polynomial ring, respectively, $R$ is a quotient of a polynomial ring. The output is
+a list with two entries: the generic projection map and the ideal (respectively the ring).
+
+It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.
+This can be faster, and can be useful for finding points.
+If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
+
+%\vspace{1em}
+\begin{comment}~~
+
+  ~~
+
+  {{\small\color{blue}
+\begin{verbatim}
+	i2 : R = ZZ/5[x, y, z];
+
+	i3 : I = ideal(random(3, R) - 2, random(2, R));
+
+	o3 : Ideal of R
+
+	i4 : projectionToHypersurface(I)
+
+	            ZZ                             4    3      2 2     4
+	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y - z))
+	            5
+
+    o4 : Sequence
+
+	i5 : projectionToHypersurface(R/I)
+
+                           	                    R                    
+	o5 = (map(--------------------------------------------------------------
+	    3       2     3            2        2       2       2     
+	(- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  - 2x
+	-------------------------------------------------------------------
+                                    ZZ                       
+	                                    --[y, z]                 
+	                                    5                       
+	--------------------,----------------------------------------------
+	      2               4 2    3 3       5     6     3       2     3 
+	*y - y  + x*z - y*z) y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  
+	-------------------------------------------------------------------
+                    	       ZZ
+                	           --[y, z]
+	                            5
+	---,{x - z, y}), -------------------------------------------------)
+	                  4 2    3 3       5     6     3       2     3
+	- 1              y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  - 1
+
+	o5 : Sequence
+	\end{verbatim}
+    }}
+\end{comment}
+
+% \end{enumerate}
+
+\section{An application:  {\tt findANonZeroMinor}, {\tt extendIdealByNonZeroMinor}}
 
@@
 
-\vspace{1em}
+\vspace{0.5em}
 -- {\tt findANonZeroMinor(n, M, I)} 
-
-\vspace{1em}
+\vspace{0.5em}
 \noindent where $I$ is an ideal
@@
 \end{description}
-
-\vspace{1em}
-
-\begin{example}~~
-
+\vspace{0.25em}
+\begin{example}
+  We demonstrate this function.
   ~~
   {{\small\color{blue}
+  \begin{quote}
 \begin{verbatim}
@@
 		o6 : Sequence
-\end{verbatim}}
+\end{verbatim}
+\end{quote}\vspace{-1em}    
     }}
@@
 
-\vspace{1em}
+\vspace{0.5em}
 -- {\tt extendIdealByNonZeroMinor(n, M, I)}
 
-\vspace{1em}
+\vspace{0.5em}
 \noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using \\{\tt findANonZeroMinor};  
 it extracts the last entry of the output, finds its determinant and
-adds it to the ideal $I$, thus extending $I$.
+adds it to the ideal $I$, thus extending $I$.   It has the same options as {\tt findANonZeroMinor}.
 
 % \vspace{1em}
-As before, we can have the following additional options:
-\begin{description}
-\item[\tt Strategy => Symbol] This specifies which strategy to use when calling {\tt randomPoints}.
-\item[\tt Homogeneous => Boolean] This is passed in calls to {\tt randomPoints}.
-\item[\tt Verbose => Boolean] This turns on or off verbose output (see \Cref{findANonZeroMinor}).
-\item[\tt MinorPointAttempts => ZZ] This controls how many points at which to check the rank of the matrix.
-\end{description}
 
-\vspace{1em}
-\begin{example}~~
+%\vspace{1em}
+\begin{comment}~~
 
@@
     }}	
-\end{example}
+\end{comment}
 
-One use for this function can be in showing that certain rings are $(R_1)$ (regular in codimension $1$).
-Consider the following example which is $(R_1)$ where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
+One can use this function to show that rings are $(R_1)$, that is, regular in codimension $1$.
+
+\begin{example}
+    Consider the following example (which is $(R_1)$) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
 However, we can use this function to quickly find interesting minors.
-
-\vspace{1em}
-
-\begin{example}~~
-
-  ~~
   {{\small\color{blue}
+  \begin{quote}
 \begin{verbatim}
@@
 
-\end{verbatim}}
+\end{verbatim}
+\end{quote}\vspace{-1em}    
     }}    
-
+    \noindent
+    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
 \end{example}
-
-In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
 

### 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
updated addresses

@@
 \email{sbisui@tulane.edu}
-\address{Tulane University}
+\address{Department of Mathematics, Tulane University, New Orleans, LA 70118}
 \author{Zhan Jiang}
 \email{zoeng@umich.edu}
-\address{University of Michigan}
+\address{Department of Mathematics, University of Michigan, Ann Arbor, MI 48109}
 \author{Sarasij Maitra}
 \email{sm3vg@virginia.edu}
-\address{University of Virginia}
+\address{Department of Mathematics, University of Virginia, Charlottesville, VA 22904}
 \author{Thai Nguyen}
 \email{tnguyen11@tulane.edu}
-\address{Tulane University}
+\address{Department of Mathematics, Tulane University, New Orleans, LA 70118}
 \author{Karl Schwede}
@@
 \address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
-\date{\today}
+%\date{\today}
+
+\thanks{Schwede was supported by NSF Grant \#1801849, NSF FRG Grant \#1952522 and a Fellowship from the Simons Foundation.}
 
@@
     There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.
-    We provide some functions that apply different strategies to generate random rational points for the given variety.
+    We develop functions that apply different strategies to generate random rational and geometric points onn the given variety.
 We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.
@@
 
-\begin{example}Now we work over a larger field.
+\begin{example}
+    Now we work over a larger field.
     {{\small\color{blue}
-        ~~
     \begin{quote}
@@
         {{\small\color{blue}
-        ~~
         \begin{quote}
@@
 
-~~
-
 \begin{comment}~~
@@
 %The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.
-
 \begin{comment}~~

### 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
added some citations

@@
 
-    There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.
-    We develop functions that apply different strategies to generate random rational and geometric points onn the given variety.
+    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.
+We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can built-in the function {\tt randomKRationalPoint} \cite{M2}, our methods frequently appear to be faster.
+
+    We develop functions that apply different strategies to generate random rational and geometric points on the given variety.
 We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.
@@
 Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
-A quick comment, while in many examples generic projection works well, there are some where it has very slow performance.
+A quick comment, while in many examples generic projection works well, there are some examples where it performs quite slowly.
 Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation.
@@
 \begin{example}	
-    We project a curve in 4-space to 2-space.
+    We project a curve in 4-space to one 2-space.
     ~~
@@
 It differs from {\tt genericProjection(codim I - 1, I)} as it only tries to find a hypersurface equation that vanishes along the projection, instead of finding one that vanishes exactly at the projection.
-This can be faster, and can be useful for finding points.
+This can be faster and can be useful for finding points.
 If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
@@
 
-There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}.
+There is some similar functionality computing partial Jacobian ideals obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}, see \cite{FastLinAlgSource}.
 
@@
 
-\vspace{2em}
+\bibliographystyle{skalpha}
+\bibliography{MainBib}
+
 \end{document}

### 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
Minor changes

@@
 We provide different strategies such as linear intersection and projection to obtain such points.
-We also present methods to obtain non-vanishing minors of some given size in a given matrix, by evaluating the matrix at a point.
+We also present methods to obtain non-vanishing minors of a given size in a given matrix, by evaluating the matrix at a point.
 %This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
@@
     There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.
-We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can built-in the function {\tt randomKRationalPoint} \cite{M2}, our methods frequently appear to be faster.
+We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can built-in the function {\tt randomKRationalPoint} \cite{M2}, our methods frequently appear to be faster and apply in the affine setting as well.
 
@@
 o6 = {{2, -30, -8}}
-
-o6 : List
         \end{verbatim}
@@
     \end{verbatim}%        
-\end{quote}%
+\end{quote}\vspace{-1em}%
 }}%
     \end{example}%    
-    \vspace{-1em}
 %
@@
 o5 = {{1, -2, 2, 2, -2, 0, 0, -1, -1, 3}}
-
-o5 : List
     \end{verbatim}
@@
 
-Both of these functions provide customizable projection techniques. We describe them here.
+We include two functions providing customizable projections. We describe them here.
 
@@
             5
-
-o3 : Sequence
 \end{verbatim}
@@
 
-This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.
+This method works by calling {\tt randomCoordinateChange} (\Cref{helper}) before dropping variables.
 It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.
@@
 		{1} | -2x2-xy-2y2-2xz+2yz+2z2 -2x   |
-
-		o6 : Sequence
 \end{verbatim}
-\end{quote}\vspace{-1em}    
+\end{quote}\vspace{-0.5em}    
     }}
@@
     \noindent
-    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
+    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one expects about $5$ steps as each minor will typically eliminate one of those primes.
 \end{example}
@@
 There is some similar functionality computing partial Jacobian ideals obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}, see \cite{FastLinAlgSource}.
+That package now uses the functionality contained here in {\tt RandomRationalPoints} in some of its functions.
 

### 2021-01-21 FastLinAlg/Paper/RandomRationalPoints.tex
Minor Corrections

--- 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
@@
     There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.
-    We develop functions that apply different strategies to generate random rational and geometric points onn the given variety.
+    We develop functions that apply different strategies to generate random rational and geometric points on the given variety.
 We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.
@@
 
-\section{Our primary pupose: {\tt randomPoints}}\label{randomPoints}
+\section{Our primary purpose: {\tt randomPoints}}\label{randomPoints}
     We start with the core function in this package:  {\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:
@@
 
-This function takes a polynomial ring as an input and outputs the coordinate change map.  
+This function takes a polynomial ring as an input and outputs the coordinate change map,
 i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
@@
 
-Alternately, instead of {$I$}, we may pass it a quotient ring.
+Alternatively, instead of {$I$}, we may pass it a quotient ring.
 It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.
@@
 \vspace{0.5em}
+
 \noindent where $I$ is an ideal
-in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime; $M$ is a matrix
+in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime, $M$ is a matrix
 over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.
@@
 \end{description}
-\vspace{0.25em}
+%\vspace{0.25em}
 \begin{example}
   We demonstrate this function.
-  ~~
   {{\small\color{blue}
@@
 \end{verbatim}
-\end{quote}\vspace{-1em}    
+\end{quote}    
     }}
@@
 \end{verbatim}
-\end{quote}\vspace{-1em}    
+\end{quote}
     }}    
@@
 \end{example} 
-
 There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}.
@@
 % \end{enumerate}
-
-\vspace{2em}
+\vspace{1em}
 \end{document}

### 2021-01-21 FastLinAlg/Paper/RandomRationalPoints.tex
fixed spacing issue

--- 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
@@
 		o9 = 5
-
 \end{verbatim}
-\end{quote}\vspace{-1em}    
+\end{quote}
     }}    

### 2021-01-22 FastLinAlg/Paper/RandomRationalPoints.tex
Works Now
minor changes

@@
 \address{Department of Mathematics, University of Virginia, Charlottesville, VA 22904}
-\author{Thai Nguyen}
+\author{Th\'ai Th\`anh Nguy$\tilde{\text{\^e}}$n}
 \email{tnguyen11@tulane.edu}
@@
     \item {\tt GenericProjection} projects to a hypersurface, via 
-        {\tt projectionToHypersurface} and then uses a {\tt BruteForce} strategy.
+        {\tt projectionToHypersurface} and then uses the {\tt BruteForce} strategy.
 
@@
 i2 : R = ZZ/101[x, y, z];
-
 i3 : J = ideal(x^3 + y^2 + 1, z^3 - x^2 - y^2 + 2);
-
 o3 : Ideal of R
-
 i4 : randomPoints(J)
     -- used 0.0181549 seconds
-
 o4 = {{32, 37, -30}}
-
 o4 : List
-
 i5 : time randomPoints(J,Strategy=>BruteForce)
     -- used 0.0146932 seconds
-
 o5 = {}
-
 o5 : List
-
 i6 : randomPoints(J,Strategy=>GenericProjection)
     -- used 0.0146932 seconds
-
 o6 = {{2, -30, -8}}
@@
 
-    When calling the Strategy {\tt GenericProjection} or \\
+    When calling the strategy {\tt GenericProjection} or \\
     {\tt HybridProjectionIntersection} from {\tt randomPoints}, this option denotes the number of trials before giving up.  This option is also passed to {\tt randomPoints} by other functions.
@@
     -- used 0.84679 seconds
-
 o7 = {{-39, -43, 28}}
@@
 However, if the field size is larger, {\tt BruteForce} will perform poorly.
+Even for some simple examples, it could not provide any rational points if the number of trials is not enough.
 Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
@@
 i2 : R = ZZ/7[x_1..x_10];
-
 i3 : I = ideal(random(2, R), random(3, R));
-
 o3 : Ideal of R
-
 i4 : time randomPoints(I, Strategy => BruteForce, PointCheckAttempts => 20000)
     -- used 0.0102075 seconds
-
 o4 = {{-2, -1, 3, -2, -2, 2, 0, -2, -2, -2}}
-
 o4 : List
-
 i5 : time randomPoints(I, Strategy => Default)
     -- used 0.085741 seconds
-
 o5 = {{1, -2, 2, 2, -2, 0, 0, -1, -1, 3}}
@@
 i6 : S = ZZ/211[x_1..x_10];
-
 i7 : J = ideal(random(2, S), random(3, S));
-
 o7 : Ideal of S
-
 i8 : time randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 2000000)
     -- used 25.3479 seconds
-
 o8 = {{-36, 26, -63, -7, -48, 78, -44, -22, 105, 39}}
-
 i9 : time randomPoints(J, Strategy => Default)
 -- used 0.102918 seconds
-
 o9 = {{0, -42, -53, -33, -65, -27, -105, -37, 99, 64}}
-
 o9 : List
-
 i10 : time randomPoints(J, Strategy => LinearIntersection)
 -- used 0.0169781 seconds
-
 o10 = {{-51, -11, -34, -33, -65, 19, -86, 50, -44, -97}}
-
 o10 : List
-
 i11 : time randomPoints(J, Strategy => GenericProjection)
 -- used 0.364902 seconds
-
 o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
@@
 -- used 0.0612306 seconds
-
             5      4      3      2                                      5  
@@
 
-This function takes a polynomial ring as an input and outputs the coordinate change map,
+This function takes a polynomial ring as an input and output the coordinate change map,
 i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
@@
 	i3 : R = ZZ/11[x, y, z];
-
 	i4 : randomCoordinateChange(R)
-
            ZZ
@@
            11
-
                     ZZ
@@
                     11
-
 	i5 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1)
-
 	o5 = | x -x-4y-5z y |
-
 	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Homogeneous => false)
-
 	o6 = | x-3 z-5 -x+3y-4z+2 |
-
 	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Replacement => Binomial)
-
 	o7 = | y x+4z z |
@@
 i1 : R = ZZ/5[x, y, z, w];
-
 i2 : I = ideal(x, y^2, w^3 + x^2);
-
 i3 : genericProjection(2, I)
-
             ZZ                                         2          2
@@
     }}
-\end{example}
-
+\end{example}\vspace{-1em}
 Alternatively, instead of {$I$}, we may pass it a quotient ring.
@@
 \begin{verbatim}
-
 i1 : R = ZZ/5[x, y, z, w];
-
 i2 : I = ideal(x, y^2, w^3 + x^2);
-
 i3 : genericProjection(2, R/I)	
-
                             ZZ                          ZZ
@@
         (x, y , w  + x )      z                            z
-
 o3 : Sequence
@@
 i2 : R = ZZ/5[x, y, z, w];
-
 i3 : I = ideal(x, y^2, w^3 + x^2);
-
 o3 : Ideal of R
-
 i4 :genericProjection(2, I, Replacement => Binomial)
-
             ZZ                                   4
@@
             5
-
 o4 : Sequence
@@
     o4 : Sequence
-
 	i5 : projectionToHypersurface(R/I)
@@
 i2 : R = ZZ/5[x, y, z];
-
 i3 : I = ideal(random(3, R) - 2, random(2, R))
-
                 3    2       2     3     2              2       2     3
@@
 - 2, - 2x  - 2x*y + 2y  + x*z - 2y*z - z )
-
 o3 : Ideal of R
-
 i4 : M = jacobian(I)
-
 o4 = {1} | -x2+2xy-y2+xz+2yz    x-2y+z   |
@@
      {1} | -2x2+2xy+y2+2yz-z2   x-2y-2z  |
-
              3       2
 o4 : Matrix R  <--- R
-
 i5 : extendIdealByNonZeroMinor(2, M, I, Strategy => LinearIntersection)
-
                 3    2       2     3     2              2       2     3
@@
 2x z + y*z  - z )
-
 o5 : Ideal of R
@@
 
-\bibliographystyle{skalpha}
+\bibliographystyle{alpha}
 \bibliography{MainBib}
 
+\vspace{1em}
 \end{document}

### 2021-01-22 FastLinAlg/Paper/RandomRationalPoints.tex
Minor changes

--- 2021-01-21 FastLinAlg/Paper/RandomRationalPoints.tex
@@
 \begin{verbatim}
-
 i1 : R = ZZ/5[x, y, z, w];

### 2021-01-23 FastLinAlg/Paper/RandomRationalPoints.tex
minor changes in the introduction and abstract

@@
 
-\thanks{Schwede was supported by NSF Grant \#1801849, NSF FRG Grant \#1952522 and a Fellowship from the Simons Foundation.}
+\thanks{Schwede was supported by NSF Grants \#1801849, FRG \#1952522 and a Fellowship from the Simons Foundation.}
 
 \begin{abstract}
-  In this article, we present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
+  We present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
 We provide different strategies such as linear intersection and projection to obtain such points.
@@
 
-    All polynomial rings considered here will be over finite fields.
-In the subsequent sections, we explain briefly the core and some helper functions and describe the strategies that we have implemented in their execution.
+    All polynomial rings considered here will be over finite fields. In the subsequent sections, we explain  the core and helper functions and describe the strategies that we have implemented.
 
@@
     \item[\tt MaxCoordinatesToReplace => ZZ] This is used for fine tuning the {\tt LinearIntersection} or {\tt GenericProjection} strategies, see \Cref{helper}.
-    \item[\tt Codimension => ZZ] If the codimension is already known, one can provide it so it is not recomputed.
+    \item[\tt Codimension => ZZ] One can provide the codimension so it is not recomputed.
 This is used in the {\tt GenericProjection} and {\tt LinearIntersection} strategies, see \Cref{projectionfunctions}.

### 2021-03-17 FastLinAlg/Paper/RandomRationalPoints.tex
PaperModified To Incorporate dimViaBezout and MultiplicationTable

@@
 The authors began work on this package at the virtual Cleveland 2020 Macaulay2 workshop.
+The authors are also grateful to the reviewers for suggesting and providing preliminary codes to speed up computations, thereby improving the efficacy of the package substantially.
 
@@
         \setlength{\itemsep}{5pt}
-    \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection} or {\tt HybridProjectionIntersection}.
+    \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection}, {\tt HybridProjectionIntersection} or {\tt MultiplicationTable}.
 
@@
     \item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersection.
+
+    \item {\tt MultiplicationTable} works for homogeneous ideals only.
+It cuts down with a linear space to a zero-dimensional projective scheme and computes how the last two variables act on the quotient ring truncated
+    at the regularity. In case of an absolutely irreducible $I$ over the field, we will find with  high probability a point, 
+    since the determinant will have a linear factor in at least 50\% of the cases.
     \end{itemize}
 
-    Notice that the speed, or success, varies depending on the strategy.
+    Notice that the speed, or success, varies depending on the strategy (see also \Cref{helper}).
 
@@
 Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
+Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck.
+If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}).
 A quick comment, while in many examples generic projection works well, there are some examples where it performs quite slowly.
-Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation.
+Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation primarily.
 
@@
     \end{example}
+To enable speed ups as indicated in the above discussion, the {\tt ProjectionAttempts} has been set to $0$ in the {\tt Default} strategy so that {\tt GenericProjection} is not opted for unless specified by user.
 
@@
 
-\section{A helper function: {\tt randomCoordinateChange}}{\label{helper}}
+\section{Two helper functions: {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
 
-% We start {\tt randomCoordinateChange}, a helper function that also might be useful to others.
+\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension significantly slows down speed.
+Instead, we compute the dimension probabilistically using {\tt dimViaBezout}.
+This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.
+For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.
+If there was no intersection, this function tries a $2$-dimensional linear space, and so on.
+This speeds up many computations.
+The function also takes in optional inputs as described below:
 
-% \item {\tt Codimension}: It takes input an integer.
-Various functions need to know the codimension/height of the scheme/ideal it is working with.
-Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.
-The default value is {\tt null}, in which case the function will compute the codimension.
-%%%   
-%%   Codimension is not a function, it is an option.
-%%%   
+\begin{itemize}
+	\item {\tt DimensionIntersectionAttempts}: 
+	This integer input denotes the number of linear spaces to try before moving to the next dimension.
+	\item {\tt MinimumFieldSize}:
+	If the ambient field is smaller than this integer value, it will automatically be replaced with an extension field.
+The user may set the {\tt MinimumFieldSize} to ensure that the field being worked over is big enough.
+For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be provided.
+\end{itemize}
+
+\begin{example} We illustrate the speed difference in this example.
+	{\color{blue}{
+			\begin{quote}
+				\begin{verbatim}
+				i3 : kk=ZZ/nextPrime 10^2;
+
+				i4 : S=kk[y_0..y_14];
+
+				i5 : I=minors(2,random(S^3,S^{5:-1}));
+
+				o5 : Ideal of S
+
+				i6 : elapsedTime dimViaBezout(I)
+				-- 0.303629 seconds elapsed
+
+				o6 = 7
+
+				i7 : elapsedTime dim I
+				-- 1.47916 seconds elapsed
+
+				o7 = 7
+				\end{verbatim}
+			\end{quote}
+	}}
+\end{example}
+
+As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using this function, and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}.
+Alternatively, all the strategies for {\tt randomPoints} (see \Cref{strategydetails}) compute the dimension using the above function, which significantly accelerates the computations.
+
 \subsection*{\tt randomCoordinateChange:}
@@
 
-	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Homogeneous => false)
+	i6 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, 
+	Homogeneous => false)
 
@@
 
-	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, Replacement => Binomial)
+	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, 
+	Replacement => Binomial)
 
@@
 \end{example}
+
 \section{Other helper functions: {\tt genericProjection}, {\tt projectionToHypersurface}}\label{projectionfunctions}
@@
 \end{verbatim}
-\end{quote}\vspace{-0.5em}    
+\end{quote}  
     }}

### 2021-03-17 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Changes txt File Added

We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples.

Changes Made to Paper:

a) Section 2.1 (Page 2),

b) 2.2 Page 3,

c) Sentence before Example 2.5 added on Page 4

d) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.

MAIN CHANGE:

We implement the dimViaBezout to compute the dimension probabilistically. This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a 2-dimensional linear space, and so on.  This speeds up many computations.
Optional Inputs:

DimensionIntersectionAttempts => ZZ : the number of linear spaces to try before moving to the next dimension
MinimumFieldSize => ZZ: if the ambient field is smaller than this value it will automatically be replaced with an extension. (The user may set the  MinimumFieldSize to ensure that the field being worked over is big enough.  For instance, there are relatively few linear spaces over a field of characteristic 2, and this can cause incorrect results to be provided.)

The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.

Changes Made to the Source Code:

We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout.

New documentation and examples have been provided.

All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known. In case, that is known, it should be specified along with the call of randomPoints to achieve high speeds.

ProjectionAttempts has been set to 0 in the Default Strategy now.

### 2021-03-18 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Some Comments Added

@@
 We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples.
+We would be very happy to invite the reviewer to be a coauthor to the source code (and also for the paper, if desired).
 
@@
 
-b) 2.2 Page 3,
+b) 2.2 Page 3;
 
@@
 This speeds up many computations.
+
 Optional Inputs:
@@
 
+The MultiplicationTable strategy has also beeen added.
+(This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently.
+The previous strategy was using decompose to break up the ideal.)
+
 Changes Made to the Source Code:
@@
 We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout.
+(This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently.
+The previous strategy was using decompose to break up the ideal.)
 

### 2021-03-18 FastLinAlg/Paper/RandomRationalPoints.tex
Minor
Comments Added and minor changes to Paper as well

@@
     \end{verbatim}
-\end{quote}\vspace{-1em}
+\end{quote}\vspace{-2em}
         }}
@@
 To enable speed ups as indicated in the above discussion, the {\tt ProjectionAttempts} has been set to $0$ in the {\tt Default} strategy so that {\tt GenericProjection} is not opted for unless specified by user.
-
 \begin{example}
@@
 
-\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension significantly slows down speed.
-Instead, we compute the dimension probabilistically using {\tt dimViaBezout}.
+\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, significantly slows down speed.
+Instead, we compute the dimension of $V(I)$ probabilistically using {\tt dimViaBezout}.
 This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.
@@
 
-As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using this function, and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}.
-Alternatively, all the strategies for {\tt randomPoints} (see \Cref{strategydetails}) compute the dimension using the above function, which significantly accelerates the computations.
+As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using the output from the above function (or by existing command {\tt codim} beforehand), and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}.
+Alternatively, one can simply rely on any of the strategies for {\tt randomPoints} (see \Cref{strategydetails}) as they compute the dimension of $V(I)$ using {\tt dimViaBezout}, which significantly accelerates the computations.
 
@@
 
-This function takes a polynomial ring as an input and output the coordinate change map,
+This function takes a polynomial ring as an input and outputs the coordinate change map,
 i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
@@
 \begin{example}	
-    We project a curve in 4-space to one 2-space.
+    We project a curve in $4$-space to one in $2$-space.
     ~~

### 2021-04-02 FastLinAlg/Paper/RandomRationalPoints.tex
Remark 2.2.1 added, New References added, Comment File also changed accordingly

@@
 \newtheorem{example}{Example}[section]
+\newtheorem{remark}{Remark}[subsection]
 %\usepackage{enumitem}
@@
     \end{example}
+
+\begin{remark}
+	In the case of an absolutely irreducible hypersurface (defined by $f$ say), there are significant discussion on estimating lower bounds of number of rational points (see for instance, \cite{RationalKPointsSource}, \cite{lang1954number},\cite{ghorpade2002number}, \cite{cafure2006improved}) all of which point to the fact that there is ``good probability" of finding a rational point in this case when we intersect with a random line.
+Heuristically, we can make the following rough estimation: A random line $L$ has $q$ points.
+We want to find the probability that one of these points is  rational for $V(f)$.
+By the proof of \cite[Proposition 2.12]{bothmer2005quick}, we get that the probability that  the chosen point does not lie in the intersection, $L\cap V(f)$, is $1-\frac{1}{q}$.
+We then exhaust this search over all the points on $L$ to get the probability that there is indeed a successful intersection is $1-(1-\frac{1}{q})^{q}$.
+As $q$ gets larger, this value tends to $1-e^{-1}\approx 0.63$, thereby showing a very good chance of successful output.
+
+\end{remark}
 

### 2021-04-02 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Updated

@@
 
+d) Page 5 Remark 2.2.1 added following reviewer's suggestion to discuss a bit about the probability of succesful output.
+
 d) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.
 
-MAIN CHANGE:
+MAIN CHANGES:
 
-We implement the dimViaBezout to compute the dimension probabilistically.
+-- We implement the dimViaBezout to compute the dimension probabilistically.
 This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.
@@
 
-The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.
+-- The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.
 
-The MultiplicationTable strategy has also beeen added.
+-- The MultiplicationTable strategy has also beeen added.
 (This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently.
 The previous strategy was using decompose to break up the ideal.)
+
+-- Following the reviewer's suggestion, we have added Remark 2.2.1 with references that all point to a good probability of finding a rational point by intersecting with a line in the case of hypersurface.
 

### 2021-04-19 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Updated Comment File
Fixes to the paper and to the comments

@@
 
-d) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.
+e) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.
 
@@
 
--- We implement the dimViaBezout to compute the dimension probabilistically.
+-- We have implemented the dimViaBezout to compute the dimension probabilistically.
 This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.
@@
 This speeds up many computations.
+
+    Some comments on implementation.
+The referee suggested using a binary search pattern, however we found that in practice, it is much faster to start with a 1-dimensional linear space, then a 2-dimensional linear space, and so on, until we find an intersection.
+This is because in our examples, checking whether an intersection is empty is much faster when the intersection is indeed empty.
+
+    Furthermore, our implementation is in the affine case and not the projective setting (it did not seem to impact performance, and now it works on a wider range of examples).
 
@@
 For instance, there are relatively few linear spaces over a field of characteristic 2, and this can cause incorrect results to be provided.)
-
--- The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.
 
@@
 
+There are some aspects of this code we do not totally understand.  In particular, we do not understand what is occurring in the lines:
+
+m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));
+and what follows.
+
+Additionally, we have had quite a few examples, especially when running something iteratively, where the saturation:
+Js:=saturateInGenericCoordinates J;
+(which only saturates in the last coordinate)
+will saturate away the points, and thus causes this function to crash.
+
+However, doing a saturation with respect to every variable (even separately) seems to make this function quite a bit slower.
+We are tempted to saturate with respect to a (random) variable, then use a try command to see if that yields a good result, and if it crashed try saturating with respect to a different (random) variable.
+However, we
+
 New documentation and examples have been provided.
@@
 All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known.
-In case, that is known, it should be specified along with the call of randomPoints to achieve high speeds.
+In case, that it is known, it should be specified along with the call of randomPoints.
 
-ProjectionAttempts has been set to 0 in the Default Strategy now.
+ProjectionAttempts has been set to 0 in the Default Strategy now for reasons discussed in

### 2021-04-19 FastLinAlg/Paper/RandomRationalPoints.tex
Fixes to the paper and to the comments

--- 2021-03-18 FastLinAlg/Paper/RandomRationalPoints.tex
@@
 \newtheorem{example}{Example}[section]
+\newtheorem{remark}{Remark}[subsection]
+
+\newcommand{\bA}{\mathbb{A}}
+\newcommand{\bF}{\mathbb{F}}
 %\usepackage{enumitem}
@@
     \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.
-It begins with less general projections and linear intersections, and gradually ramps up the randomness.
+It begins begins with brute force and moves to linear intersections with particularly simple linear forms, and gradually ramps up the randomness.
+If {\tt ProjectionAttempts} is set to a value bigger than zero (it is set to zero by default), then the strategy alternates between generic projects and hybrid projection / intersections.
+For additional discussion, see \cref{rem.ProjectingToAHypersurfaceFirst}.
     \item {\tt BruteForce} simply tries random points and sees if they are on the variety.
@@
         {\tt projectionToHypersurface} and then uses the {\tt BruteForce} strategy.
+Finally, it pulls back back the points that were found to the original variety (via a linear intersection).
+See \cref{rem.ProjectingToAHypersurfaceFirst}.
 
@@
 
-    \item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersection.
-
+    \item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersectio with that hypersurface.
+Finally, it pulls back back the points that were found to the original variety (via a linear intersection).
+See \cref{rem.ProjectingToAHypersurfaceFirst}.
     \item {\tt MultiplicationTable} works for homogeneous ideals only.
-It cuts down with a linear space to a zero-dimensional projective scheme and computes how the last two variables act on the quotient ring truncated
-    at the regularity. In case of an absolutely irreducible $I$ over the field, we will find with  high probability a point, 
-    since the determinant will have a linear factor in at least 50\% of the cases.
+It cuts down with a linear space to a zero-dimensional projective scheme and computes how the last two variables act on the quotient ring truncated at the regularity.
+%In case of an absolutely irreducible $I$ over the field, we will find with  high probability a point, since the determinant will have a linear factor in at least 50\% of the cases.
     \end{itemize}
@@
 If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}).
-A quick comment, while in many examples generic projection works well, there are some examples where it performs quite slowly.
-Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation primarily.
+A quick comment, while in many examples generic projection works well, there are some examples where it is many orders of magnitude slower.
+By default, generic projection is turned off in the default strategy, that is we set {\tt ProjectionAttempts=>0}.
+Increasing this number will include it.
 
@@
     \end{example}
+
+\begin{remark}[Finding a point]
+	In the case of an absolutely irreducible hypersurface in $\bA^n_{\bF_q}$ (defined by $f$ say), there is significant discussion in the literature estimating lower bounds of number of rational points (see for instance, \cite{RationalKPointsSource,lang1954number,ghorpade2002number,cafure2006improved}) all of which point to the fact that there is ``good probability" of finding a rational point in this case when we intersect with a random line.
+Heuristically, we can make the following rough estimation.
+We expect that each equation $f = \lambda$ for $\lambda \in \bF_q$ has approximately the same number of solutions.
+Since each point on $\bF_q^n$ solves exactly one of these equations, we expect that $f = 0$ has approximately $q^{n-1}$ solutions, or in other words, our hypersurface has $q^{n-1}$-points.
+       Now, a random line $L$ has $q$ points.
+We want to find the probability that one of these points is  rational for $V(f)$.
+We would expect that if these points are randomly distributed, then the probability that our line contains one of those points $1-(1-{1 \over q})^q$ which tends to $1-e^{-1} \approx 0.63$ for $q$ large.
+Alternately, one can use the proof of \cite[Proposition 2.12]{bothmer2005quick} for a more precise statement.
+For each point of $L$, we see that the probability that  the chosen point does not lie in the intersection, $L\cap V(f)$, is $1-\frac{1}{q}$.
+We then exhaust this search over all the points on $L$ to get the probability that there is indeed a successful intersection is $1-(1-\frac{1}{q})^{q}$.
+As $q$ gets larger, this value tends to $1-e^{-1}\approx 0.63$.
+
+       Of course, there are certainly examples schemes over $\bF_q$ with no rational points at all, even for curves in $\bA^2$.
+\end{remark}
+
+\begin{remark}[Projecting to a hypersurface first]
+    \label{rem.ProjectingToAHypersurfaceFirst}
+    Suppose $X \subseteq \bA^n$ is an algebraic set.
+In a number of existing algorithms, one first does a generic (or even not very generic) projection $h : \bA^n \to \bA^m$ and so that $h(X)$ is a hypersurface (at least set theoretically).
+Then we find a point $x \in h(X)$ (say as above), and compute $h^{-1}(\{ x \})$ which is a linear space in $\bA^n$ that typically intersects $X$ in a rational point.
+For example, this is done in {\tt randomKRationalPoint} in core Macaulay2.
+Note that projecting to a hypersurface \emph{still is performing an intersection with a linear space}, but it just tries to choose the linear space more intelligently.
+
+    However, while in a number of examples, doing this generic projection first can speed up computations (as opposed to simply intersecting with a linear space, since we have identified a linear space initially), there are numerous cases where computing this hypersurface $h(X)$ can be extremely slow.
+This particularly appears in cases when one is computing successive minors to identify the locus where some variety is nonsingular.
+Thus, by default, we disable the functionality to project to a hypersurface in the default strategy.
+
+    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (ie, with many linear or binomial terms).
+Such simple linear spaces are the ones considered in {\tt randomKRationalPoint} for instance.
+In practice, this seems give approximately the same performance as projecting to a hypersurface, without the chance of the code getting hung up on the generic projection.
+\end{remark}
 
@@
 
-\vspace{1em}
 \end{document}

### 2021-04-20 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Various changes on the paper, and on the main code.

@@
 
-ProjectionAttempts has been set to 0 in the Default Strategy now for reasons discussed in
+ProjectionAttempts has been set to 0 in the Default Strategy now for reasons discussed.
+
+-------------------------------
+
+We also provide some examples demonstrating performance in some somewhat larger examples.
+
+-------------------------
+----Large Example #1-----
+-------------------------
+
+-------------------------
+----Large Example #3-----
+-------------------------
+
+Here is an example (coming out of a jacobian computation) where the rational points tend to be saturated away, which causes the MultiplicationTable strategy to fail (at least based on our understanding).
+
+In this example, the primary speed limitation still appears to be the dimension computation.
+
+T2 = ZZ/101[x1,x2,x3,x4,x5,x6,x7, Z];
+J2 = ideal(x5*x6-x4*x7,x1*x6-x2*x7,x5^2-x1*x7,x4*x5-x2*x7,x4^2-x2*x6,x1*x4-x2*x5,x2*x3^3*x5+3*x2*x3^2*x7*Z+8*x2^2*x5*Z^2+3*x3*
+      x4*x7*Z^2-8*x4*x7*Z^3+x6*x7*Z^3,x1*x3^3*x5+3*x1*x3^2*x7*Z+8*x1*x2*x5*Z^2+3*x3*x5*x7*Z^2-8*x5*x7*Z^3+x7^2*Z^3,x2*x3^3*x4+3*
+      x2*x3^2*x6*Z+8*x2^2*x4*Z^2+3*x3*x4*x6*Z^2-8*x4*x6*Z^3+x6^2*Z^3,x2^2*x3^3+3*x2*x3^2*x4*Z+8*x2^3*Z^2+3*x2*x3*x6*Z^2-8*x2*x6*
+      Z^3+x4*x6*Z^3,x1*x2*x3^3+3*x2*x3^2*x5*Z+8*x1*x2^2*Z^2+3*x2*x3*x7*Z^2-8*x2*x7*Z^3+x4*x7*Z^3,x1^2*x3^3+3*x1*x3^2*x5*Z+8*x1^2
+      *x2*Z^2+3*x1*x3*x7*Z^2-8*x1*x7*Z^3+x5*x7*Z^3,x1*x2^2*x4-x2^3*x5,-x1^2*x3^3*x5^2-6*x1*x3^2*x5^3*Z-6*x3*x5^4*Z^2-3*x1*x3*x5^
+      2*x7*Z^2+16*x5^4*Z^3+8*x1*x5^2*x7*Z^3-4*x5^3*x7*Z^3,-x2^2*x3^3*x4^2-6*x2*x3^2*x4^3*Z-8*x2^3*x4^2*Z^2-6*x3*x4^4*Z^2-3*x2*x3
+      *x4^2*x6*Z^2+16*x4^4*Z^3+8*x2*x4^2*x6*Z^3-4*x4^3*x6*Z^3,-2*x3^3*x5^2*x7^2-x1*x3^3*x7^3-6*x3^2*x5*x7^3*Z-8*x1*x5^2*x6*x7*Z^
+      2-8*x1*x4*x5*x7^2*Z^2-16*x2*x5^2*x7^2*Z^2-8*x1*x2*x7^3*Z^2-3*x3*x7^4*Z^2+8*x7^4*Z^3,-8*x1^5,-x1^2*x3^3*x5^2+x1^3*x3^3*x7-6
+      *x1*x3^2*x5^3*Z+6*x1^2*x3^2*x5*x7*Z+8*x1^3*x4*x5*Z^2-16*x1^2*x2*x5^2*Z^2-6*x3*x5^4*Z^2+8*x1^3*x2*x7*Z^2+3*x1*x3*x5^2*x7*Z^
+      2+3*x1^2*x3*x7^2*Z^2+16*x5^4*Z^3-8*x1*x5^2*x7*Z^3-4*x5^3*x7*Z^3-8*x1^2*x7^2*Z^3+4*x1*x5*x7^2*Z^3,-x1^2*x5^2+x1^3*x7,-6*x1^
+      2*x2*x3^5*x5*x7-30*x1*x2*x3^4*x5^2*x7*Z-48*x1^2*x2^2*x3^2*x5*x7*Z^2-36*x2*x3^3*x5^3*x7*Z^2-24*x1*x2*x3^3*x5*x7^2*Z^2+5*x1*
+      x2^2*x3*x5^2*x7*Z^3+48*x1*x2*x3^2*x5*x7^2*Z^3+47*x2*x3^2*x5^2*x7^2*Z^3-3*x1*x2*x3^2*x7^3*Z^3-48*x1*x2^2*x5*x7^2*Z^4-5*x2*
+      x3*x5^2*x7^2*Z^4-24*x2*x3*x5*x7^3*Z^4+48*x2*x5*x7^3*Z^5-3*x2*x7^4*Z^5,18*x1^2*x2^3*x3^6+24*x1^2*x2^3*x3^5*Z+24*x1^2*x2^2*
+      x3^5*x4*Z+9*x1*x2^3*x3^5*x5*Z+29*x1^2*x2^4*x3^3*Z^2+29*x1^2*x2^2*x3^4*x4*Z^2+19*x1*x2^3*x3^4*x5*Z^2-6*x1*x2^2*x3^4*x4*x5*Z
+      ^2-9*x2^3*x3^4*x5^2*Z^2+18*x1*x2^3*x3^4*x7*Z^2-10*x1^2*x2^4*x3^2*Z^3-24*x1^2*x2^3*x3^2*x4*Z^3-43*x1*x2^4*x3^2*x5*Z^3-45*x1
+      *x2*x3^3*x4^2*x5*Z^3-18*x2^2*x3^3*x4*x5^2*Z^3+43*x1^2*x2^2*x3^3*x6*Z^3+9*x1*x2^2*x3^3*x5*x6*Z^3+24*x1*x2^3*x3^3*x7*Z^3+24*
+      x1*x2^2*x3^3*x4*x7*Z^3-20*x1*x2^4*x3*x5*Z^4-48*x1*x2^3*x3*x4*x5*Z^4-39*x1*x2*x3^2*x4^2*x5*Z^4-15*x2*x3^2*x4^2*x5^2*Z^4+10*
+      x1^2*x2^2*x3^2*x6*Z^4+19*x1*x2^2*x3^2*x5*x6*Z^4-18*x1*x2*x3^2*x4*x5*x6*Z^4+3*x2^2*x3^2*x5^2*x6*Z^4+6*x1^2*x2*x3^2*x6^2*Z^4
+      +29*x1*x2^4*x3*x7*Z^4+29*x1*x2^2*x3^2*x4*x7*Z^4+20*x1*x2*x3*x4^2*x5*Z^5+48*x2*x3*x4^2*x5^2*Z^5+48*x1*x2*x3*x4*x5*x6*Z^5+9*
+      x1*x3*x4^2*x5*x6*Z^5-48*x2^2*x3*x5^2*x6*Z^5-6*x2*x3*x4*x5^2*x6*Z^5-6*x1*x2*x3*x5*x6^2*Z^5-10*x1*x2^4*x7*Z^5-24*x1*x2^3*x4*
+      x7*Z^5+43*x1*x2^2*x3*x6*x7*Z^5-24*x1*x4^2*x5*x6*Z^6+3*x4^2*x5^2*x6*Z^6+48*x1*x2*x5*x6^2*Z^6-6*x2*x5^2*x6^2*Z^6+10*x1*x2^2*
+      x6*x7*Z^6+6*x1*x2*x6^2*x7*Z^6,-x1^3*x4+x1^2*x2*x5,2*x1*x2*x4^2-2*x2^2*x4*x5-x1*x2^2*x6+x2^3*x7,-2*x3^3*x5^2*x7^2-x1*x3^3*
+      x7^3-6*x3^2*x5*x7^3*Z-16*x1*x5^2*x6*x7*Z^2-16*x2*x5^2*x7^2*Z^2-8*x1*x2*x7^3*Z^2-3*x3*x7^4*Z^2+8*x7^4*Z^3,-2*x4^4+2*x2*x4^2
+      *x6,-27*x1*x2^3*x3^5*x5-29*x1*x2^3*x3^4*x5*Z-36*x1*x2^2*x3^4*x4*x5*Z-27*x2^3*x3^4*x5^2*Z+43*x1*x2^2*x3^3*x4*x5*Z^2-9*x1*x2
+      *x3^3*x4^2*x5*Z^2+47*x2^2*x3^3*x4*x5^2*Z^2-9*x1*x2^2*x3^3*x5*x6*Z^2+18*x2^3*x3^3*x5*x7*Z^2+10*x1*x2^2*x3^2*x4*x5*Z^3+24*x1
+      *x2*x3^2*x4^2*x5*Z^3-10*x2^3*x3^2*x5^2*Z^3+38*x2^2*x3^2*x4*x5^2*Z^3-15*x2*x3^2*x4^2*x5^2*Z^3+24*x1*x2^2*x3^2*x5*x6*Z^3-3*
+      x1*x2*x3^2*x4*x5*x6*Z^3-18*x2^2*x3^2*x5^2*x6*Z^3+10*x2^3*x3^2*x5*x7*Z^3+6*x2^2*x3^2*x4*x5*x7*Z^3+20*x2^2*x3*x4*x5^2*Z^4+48
+      *x2*x3*x4^2*x5^2*Z^4+48*x2^2*x3*x5^2*x6*Z^4-6*x2*x3*x4*x5^2*x6*Z^4-20*x2^3*x3*x5*x7*Z^4-48*x2^2*x3*x4*x5*x7*Z^4+9*x2*x3*x4
+      ^2*x5*x7*Z^4-18*x2^2*x3*x5*x6*x7*Z^4-24*x2*x4^2*x5*x7*Z^5+3*x4^3*x5*x7*Z^5+48*x2^2*x5*x6*x7*Z^5-6*x2*x4*x5*x6*x7*Z^5,-3*x1
+      ^2*x2*x3^5*x4*x5-6*x1*x2^2*x3^5*x5^2-12*x1*x2*x3^4*x4*x5^2*Z-9*x2^2*x3^4*x5^3*Z-12*x1^2*x2*x3^4*x4*x7*Z-12*x1*x2^2*x3^4*x5
+      *x7*Z-24*x1^2*x2^2*x3^2*x4*x5*Z^2-48*x1*x2^3*x3^2*x5^2*Z^2-15*x1^2*x3^3*x4^2*x7*Z^2-48*x1*x2*x3^3*x4*x5*x7*Z^2-27*x2^2*x3^
+      3*x5^2*x7*Z^2+48*x1*x2^2*x3*x4*x5^2*Z^3+5*x1^2*x2^2*x3*x4*x7*Z^3+24*x1^2*x3^2*x4^2*x7*Z^3+5*x1*x2^3*x3*x5*x7*Z^3+24*x1*x2*
+      x3^2*x4*x5*x7*Z^3-36*x1*x3^2*x4^2*x5*x7*Z^3+24*x2^2*x3^2*x5^2*x7*Z^3-12*x2*x3^2*x4*x5^2*x7*Z^3-18*x1*x2*x3^2*x4*x7^2*Z^3-
+      18*x2^2*x3^2*x5*x7^2*Z^3-48*x1^2*x2*x4^2*x7*Z^4-24*x1*x2^2*x4*x5*x7*Z^4+48*x1*x3*x4^2*x5*x7*Z^4+48*x1*x2*x3*x4*x7^2*Z^4-18
+      *x1*x3*x4^2*x7^2*Z^4+48*x2^2*x3*x5*x7^2*Z^4-15*x2*x3*x4*x5*x7^2*Z^4+48*x1*x4^2*x7^2*Z^5+24*x2*x4*x5*x7^2*Z^5-3*x4^2*x5*x7^
+      2*Z^5,0,9*x1^3*x2^2*x3^7+9*x1^3*x2*x3^6*x4*Z+36*x1^2*x2^2*x3^6*x5*Z-29*x1^3*x2^3*x3^4*Z^2-24*x1^3*x2*x3^5*x4*Z^2+24*x1^2*
+      x2^2*x3^5*x5*Z^2+39*x1^2*x2*x3^5*x4*x5*Z^2+30*x1*x2^2*x3^5*x5^2*Z^2+3*x1^3*x2*x3^5*x6*Z^2+18*x1^2*x2^2*x3^5*x7*Z^2-29*x1^3
+      *x2^2*x3^3*x4*Z^3-29*x1^2*x2^3*x3^3*x5*Z^3-19*x1^2*x2*x3^4*x4*x5*Z^3-29*x1*x2^2*x3^4*x5^2*Z^3+45*x1*x2*x3^4*x4*x5^2*Z^3+15
+      *x1^2*x2*x3^4*x5*x6*Z^3-24*x1^2*x2^2*x3^4*x7*Z^3+24*x1^2*x2*x3^4*x4*x7*Z^3+6*x1*x2^2*x3^4*x5*x7*Z^3+10*x1^3*x2^2*x3^2*x4*Z
+      ^4-10*x1^2*x2^3*x3^2*x5*Z^4-5*x1^2*x2^2*x3^2*x4*x5*Z^4+24*x1*x2^3*x3^2*x5^2*Z^4-43*x1*x2*x3^3*x4*x5^2*Z^4+24*x1^3*x2^2*x3^
+      2*x6*Z^4+18*x1*x2*x3^3*x5^2*x6*Z^4+29*x1^2*x2^3*x3^2*x7*Z^4-43*x1^2*x2*x3^3*x4*x7*Z^4-6*x1^2*x3^3*x4^2*x7*Z^4+43*x1*x2^2*
+      x3^3*x5*x7*Z^4+48*x1*x2*x3^3*x4*x5*x7*Z^4+12*x1^2*x2*x3^3*x6*x7*Z^4-27*x1*x2^2*x3^3*x7^2*Z^4+20*x1^2*x2^2*x3*x4*x5*Z^5+48*
+      x1^2*x2^2*x3*x5*x6*Z^5-20*x1^2*x2^3*x3*x7*Z^5+5*x1^2*x2^2*x3*x4*x7*Z^5-10*x1^2*x2*x3^2*x4*x7*Z^5+48*x1*x2^3*x3*x5*x7*Z^5+
+      10*x1*x2^2*x3^2*x5*x7*Z^5+39*x1*x2*x3^2*x4*x5*x7*Z^5-9*x1*x3^2*x4^2*x5*x7*Z^5+15*x2*x3^2*x4*x5^2*x7*Z^5-24*x1^2*x2*x3^2*x6
+      *x7*Z^5+3*x1^2*x3^2*x4*x6*x7*Z^5+27*x1*x2*x3^2*x5*x6*x7*Z^5+14*x1*x2^2*x3^2*x7^2*Z^5-27*x1*x2*x3^2*x4*x7^2*Z^5-48*x1^2*x2*
+      x4^2*x7*Z^6+24*x1*x2^2*x4*x5*x7*Z^6-20*x1*x2*x3*x4*x5*x7*Z^6-48*x2*x3*x4*x5^2*x7*Z^6+24*x1^2*x2^2*x6*x7*Z^6-48*x1*x2*x3*x5
+      *x6*x7*Z^6+6*x1*x3*x4*x5*x6*x7*Z^6+20*x1*x2^2*x3*x7^2*Z^6-5*x1*x2*x3*x4*x7^2*Z^6-18*x1*x3*x4^2*x7^2*Z^6+3*x2*x3*x4*x5*x7^2
+      *Z^6+9*x1*x2*x3*x6*x7^2*Z^6+48*x1*x4^2*x7^2*Z^7-24*x2*x4*x5*x7^2*Z^7-3*x4^2*x5*x7^2*Z^7-24*x1*x2*x6*x7^2*Z^7+3*x1*x4*x6*x7
+      ^2*Z^7,3*x1*x3^5*x4*x5^2+3*x1*x2*x3^5*x5*x7+8*x1^2*x2*x3^3*x4*x5*Z+9*x1*x3^4*x5^2*x6*Z+9*x1*x2*x3^4*x7^2*Z+48*x1*x2*x3^2*
+      x4*x5^2*Z^2+24*x1^2*x2*x3^2*x5*x6*Z^2+24*x1^2*x2*x3^2*x4*x7*Z^2+24*x1*x2^2*x3^2*x5*x7*Z^2+x3^3*x4*x5^2*x7*Z^2+9*x1*x3^3*x5
+      *x6*x7*Z^2+9*x1*x3^3*x4*x7^2*Z^2+x2*x3^3*x5*x7^2*Z^2-37*x1^2*x2^2*x4*x5*Z^3+48*x1^2*x3*x4*x5*x6*Z^3+24*x1^2*x3*x4^2*x7*Z^3
+      -24*x1*x3^2*x5*x6*x7*Z^3+3*x3^2*x5^2*x6*x7*Z^3-24*x1*x3^2*x4*x7^2*Z^3+6*x1*x3^2*x6*x7^2*Z^3+3*x2*x3^2*x7^3*Z^3-27*x1^2*x4*
+      x5*x6*Z^4+16*x1^2*x5*x6^2*Z^4+37*x1^2*x4^2*x7*Z^4+16*x2*x4*x5^2*x7*Z^4+16*x1^2*x4*x6*x7*Z^4+8*x2^2*x5*x7^2*Z^4+3*x3*x5*x6*
+      x7^2*Z^4+3*x3*x4*x7^3*Z^4-8*x5*x6*x7^2*Z^5-8*x4*x7^3*Z^5+2*x6*x7^3*Z^5,-2*x5^4+2*x1*x5^2*x7,-2*x1*x2^3*x3^3-3*x1*x2^2*x3^2
+      *x4*Z-24*x1*x2^4*Z^2+x1*x2*x4*x6*Z^3,-x2^2*x4*x5+x1*x2^2*x6,6*x1*x2^3*x3^7+8*x1*x2^3*x3^6*Z+8*x1*x2^2*x3^6*x4*Z+24*x1*x2^4
+      *x3^4*Z^2-24*x1*x2^2*x3^5*x4*Z^2-18*x2^3*x3^5*x7*Z^2+27*x1*x2^4*x3^3*Z^3-45*x1*x2^3*x3^3*x4*Z^3+48*x1*x2^2*x3^4*x6*Z^3+24*
+      x2^3*x3^4*x7*Z^3-30*x2^2*x3^4*x4*x7*Z^3+10*x1*x2^5*x3*Z^4+10*x1*x2^3*x3^2*x4*Z^4+37*x1*x2^2*x3^3*x6*Z^4+2*x1*x2*x3^3*x6^2*
+      Z^4-29*x2^4*x3^2*x7*Z^4-37*x2^3*x3^3*x7*Z^4+27*x2^2*x3^3*x4*x7*Z^4-8*x2*x3^3*x4^2*x7*Z^4+7*x1*x2^5*Z^5+37*x1*x2^4*x4*Z^5-
+      20*x1*x2^3*x3*x6*Z^5+20*x2^4*x3*x7*Z^5+48*x2^3*x3*x4*x7*Z^5+10*x2^2*x3^2*x4*x7*Z^5+24*x2*x3^2*x4^2*x7*Z^5-43*x2^2*x3^2*x6*
+      x7*Z^5-7*x1*x2^3*x6*Z^6+16*x1*x2^2*x6^2*Z^6+7*x2^4*x7*Z^6-27*x2^3*x4*x7*Z^6+8*x2^2*x4^2*x7*Z^6-30*x2^2*x3*x6*x7*Z^6-48*x2*
+      x3*x4*x6*x7*Z^6-6*x2*x3*x6^2*x7*Z^6-7*x2^2*x6*x7*Z^7-37*x2*x4*x6*x7*Z^7+16*x2*x6^2*x7*Z^7-2*x4*x6^2*x7*Z^7,-2*x1*x2*x3^6*
+      x5*x7-3*x2*x3^5*x5^2*x7*Z-48*x1*x2^2*x3^3*x5*x7*Z^2-3*x1*x3^4*x5*x6*x7*Z^2-3*x2*x3^4*x5*x7^2*Z^2-48*x2^2*x3^2*x5^2*x7*Z^3+
+      24*x1^2*x2*x3^2*x6*x7*Z^3+8*x1*x3^3*x5*x6*x7*Z^3-x3^3*x5^2*x6*x7*Z^3+8*x2*x3^3*x5*x7^2*Z^3-7*x1*x3^3*x6*x7^2*Z^3+47*x1*x2^
+      3*x5*x7*Z^4-48*x1*x2*x3*x5*x6*x7*Z^4-48*x2^2*x3*x5*x7^2*Z^4+24*x1*x3^2*x6*x7^2*Z^4+27*x1*x2*x5*x6*x7*Z^5-16*x2*x5^2*x6*x7*
+      Z^5+8*x1^2*x6^2*x7*Z^5+27*x2^2*x5*x7^2*Z^5+16*x1*x2*x6*x7^2*Z^5+3*x3*x6*x7^3*Z^5-8*x6*x7^3*Z^6,-3*x1^2*x2^2*x3^2*x7-6*x1*
+      x2^2*x3*x5*x7*Z-3*x1*x2^2*x7^2*Z^2,-6*x1*x2^2*x3^5*x5^2-6*x1*x2*x3^4*x4*x5^2*Z-6*x2^2*x3^4*x5^3*Z-18*x1*x2^2*x3^4*x5*x7*Z-
+      48*x1*x2^3*x3^2*x5^2*Z^2-12*x1*x2*x3^3*x5^2*x6*Z^2-30*x2^2*x3^3*x5^2*x7*Z^2-10*x1*x2^2*x3*x4*x5^2*Z^3+15*x2^3*x3*x5^3*Z^3+
+      48*x1*x2*x3^2*x5^2*x6*Z^3-18*x2*x3^2*x5^3*x6*Z^3-48*x2^2*x3^2*x5^2*x7*Z^3+18*x2*x3^2*x4*x5^2*x7*Z^3-6*x1*x2*x3^2*x5*x6*x7*
+      Z^3-12*x2^2*x3^2*x5*x7^2*Z^3-5*x1*x2^2*x5^2*x6*Z^4-5*x2*x3*x5^3*x6*Z^4-43*x2^3*x5^2*x7*Z^4+5*x2*x3*x4*x5^2*x7*Z^4-12*x2*x3
+      *x5^2*x6*x7*Z^4+12*x2*x3*x4*x5*x7^2*Z^4);
+
+i55 :  time randomPoints(J2, Strategy=>MultiplicationTable)
+RandomRationalPoints.m2:840:17:(3):[12]: error: expected a polynomial ring
+RandomRationalPoints.m2:840:17:(3):[12]: --entering debugger (type help to see debugger commands)                                
+RandomRationalPoints.m2:840:17-840:79: --source code:                                                                            
+            m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));
+
+i56 : time randomPoints(J2, Strategy=>Default)
+     -- used 4.2498 seconds
+
+o56 = {{0, 0, 23, 0, 0, 42, 29, 0}}
+
+o56 : List

### 2021-04-20 FastLinAlg/Paper/RandomRationalPoints.tex
Various changes on the paper, and on the main code.

@@
     When working over very small fields, frequently {\tt BruteForce} is most efficient.
-This is not surprising as there may not be many points to check, and a surprisingly large percentage of points can be rational in hypersurfaces and complete intersections.
+This is not surprising as there may not be many points to check.
 However, if the field size is larger, {\tt BruteForce} will perform poorly.
 Even for some simple examples, it could not provide any rational points if the number of trials is not enough.
-Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.
-Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck.
+Other strategies work differently on different examples, and the same strategy can sometimes work very quickly even if it typically works very slowly.
+
+    Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck in some cases.
 If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}).
@@
 By default, generic projection is turned off in the default strategy, that is we set {\tt ProjectionAttempts=>0}.
-Increasing this number will include it.
+See \autoref{rem.ProjectingToAHypersurfaceFirst} for additional related discussion.
 

### 2021-04-20 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
additional comments on dimViaBezout

@@
 
-    Some comments on implementation.
-The referee suggested using a binary search pattern, however we found that in practice, it is much faster to start with a 1-dimensional linear space, then a 2-dimensional linear space, and so on, until we find an intersection.
+    Some comments on implementation.  We have implemented this in the affine case and the projective case.
+
+    In the projective case, saturate against all coordinates, instead of just the last coordinate.  Otherwise, we were getting incorrect answers too frequently.
+
+    We also implemented this in the affine setting affine examples where we do not saturate.
+
+    In the projective case, the referee suggested a binary search pattern, and we still use that there.
+In the affine case however, we found that in practice, it is much faster to start with a 1-dimensional linear space, then a 2-dimensional linear space, and so on, until we find an intersection.
 This is because in our examples, checking whether an intersection is empty is much faster when the intersection is indeed empty.
 
-    Furthermore, our implementation is in the affine case and not the projective setting (it did not seem to impact performance, and now it works on a wider range of examples).
+    Unfortunately, there are large classes of examples where either the probablistic code, for either the affine or non-affine examples, are much slower than simply calling dim.
+We would like to implement both simultaneously by multi-threading.
+Unfortunately, canceling tasks in multithreaded examples still creates a great deal of instability in M2.
 
@@
 All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known.
-In case, that it is known, it should be specified along with the call of randomPoints.
+In case that it is known, it should be specified along with the call of randomPoints.
 

### 2021-04-24 FastLinAlg/Paper/RandomRationalPoints.tex
Minor Changes on Bib File and Paper

@@
 
-    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (ie, with many linear or binomial terms).
+    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (i.e., with many linear or binomial terms).
 Such simple linear spaces are the ones considered in {\tt randomKRationalPoint} for instance.

### 2021-04-24 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
A bunch of changes.

@@
 
-    In the projective case, saturate against all coordinates, instead of just the last coordinate.  Otherwise, we were getting incorrect answers too frequently.
+    In the projective case, we saturate against all coordinates, instead of just the last coordinate.
+Otherwise, we were getting incorrect answers too frequently, especially in iterative processes where a point is found, used to identify a submatrix and a minor is computed, a new point is found, a minor computed, etc.
+Perhaps it would instead be better to saturate with respect to ceiling(sqrt(# of variables))?
+In cases where there are many variables, saturating with respect to all of them can be pretty slow.
 
-    We also implemented this in the affine setting affine examples where we do not saturate.
+    We also implemented this in the affine setting affine examples where we do not saturate but instead try to compute whether 1 is in the ideal.
 
@@
 Unfortunately, canceling tasks in multithreaded examples still creates a great deal of instability in M2.
+Hence we have created the option
+    DimensionFunction
+        which by default is DimensionFunction => dimViaBezout and through which the user can specify to use the built in Macaulay2 dimension function, or their own custom dimension function.
 
-Optional Inputs:
+Optional Inputs for dimViaBezout:
 
@@
 We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout.
-(This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently.
-The previous strategy was using decompose to break up the ideal.)
+This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently.
+The previous strategy was using decompose to break up the ideal.
 
@@
 We are tempted to saturate with respect to a (random) variable, then use a try command to see if that yields a good result, and if it crashed try saturating with respect to a different (random) variable.
-However, we
+However, we are not completely sure what the best way forward would be.
 
@@
 
-ProjectionAttempts has been set to 0 in the Default Strategy now for reasons discussed.
+ProjectionAttempts has been set to 0 in the Default Strategy because in too many examples, especially when iteratively computing minors, it runs into a situation where trying to do even a simple projection to a coordinate hangs (takes much too long).
 
@@
 -------------------------
-----Large Example #1-----
+----Largish Example #1-----
 -------------------------
+
+The first example just grabs 15 minors of some matrix.
+Notice that dimDegViaBezout is much faster, as it does much less saturation.
+The dimension function does seem to be the slowdown in this example, and MultiplicationTable is almost immediate if you specify that Codimesion=>8, whereas the default strategy is still on the order of 0.5 seconds.
+
+i2 : loadPackage "FastLinAlg";
+
+i3 : R = (ZZ/101)[YY_1, YY_2, YY_3, YY_4, YY_5, YY_6, YY_7, YY_8, YY_9];
+
+i4 : I2 =  ideal(YY_8^2-YY_7*YY_9,YY_6*YY_8-YY_5*YY_9,YY_3*YY_8-YY_2*YY_9,YY_2*YY_8-YY_1*YY_9,YY_6*YY_7-YY_5*YY_8,YY_3*YY_7-YY_1*YY_9,YY_2*YY_7-YY_1*YY_8,YY_6^2-YY_4*YY_9,YY_5*YY_6-YY_4*YY_8,YY_4*YY_6+YY_1*YY_8-10*YY_1*YY_9-YY_2*YY_9+10*YY_3*YY_9,YY_3*YY_6-YY_8*YY_9-10*YY_9^2,YY_2*YY_6-YY_7*YY_9-10*YY_8*YY_9,YY_1*YY_6-YY_7*YY_8-10*YY_7*YY_9,YY_5^2-YY_4*YY_7,YY_4*YY_5+YY_1*YY_7-10*YY_1*YY_8-YY_1*YY_9+10*YY_2*YY_9,YY_3*YY_5-YY_7*YY_9-10*YY_8*YY_9,YY_2*YY_5-YY_7*YY_8-10*YY_7*YY_9,YY_1*YY_5-YY_7^2-10*YY_7*YY_8,YY_4^2+YY_7^2-YY_9^2,YY_3*YY_4-YY_5*YY_9-10*YY_6*YY_9,YY_2*YY_4-YY_5*YY_8-10*YY_5*YY_9,YY_1*YY_4-YY_5*YY_7-10*YY_5*YY_8,YY_2^2-YY_1*YY_3,YY_1*YY_2-10*YY_1*YY_3-YY_2*YY_3+10*YY_3^2+YY_4*YY_8+10*YY_4*YY_9,YY_1^2-YY_3^2+YY_4*YY_7+20*YY_4*YY_8-YY_4*YY_9);
+
+o4 : Ideal of R
+
+i5 : M2 = jacobian I2;
+
+i6 : J2 = I2 + chooseGoodMinors(15, 7, M2);
+
+o6 : Ideal of R
+
+i7 : time dim J2
+     -- used 0.00395536 seconds
+
+o7 = 1
+
+i8 : time dimViaBezout(J2)
+     -- used 0.965556 seconds
+
+o8 = 1
+
+i14 : time dimDegViaBezout(J2)
+     -- used 0.0362843 seconds
+
+o14 = (1, 7)
+
+o14 : Sequence
+
+i29 : time randomPoints(J2)
+     -- used 1.26618 seconds
+
+o29 = {{0, 0, 0, 0, 0, 0, 30, -3, -30}}
+
+o29 : List
+
+i30 : time randomPoints(J2, Strategy=>MultiplicationTable)
+     -- used 1.03038 seconds
+
+o30 = {{0, 0, 0, 0, 0, 0, -1, -10, 1}}
+
+o30 : List
+
+i31 : time randomPoints(J2, Strategy=>LinearIntersection)
+     -- used 1.90992 seconds
+
+o31 = {{0, 0, 0, 0, 0, 0, 15, 49, -15}}
+
+o31 : List
+
+i32 : time randomPoints(J2, Strategy=>LinearIntersection)
+     -- used 3.00063 seconds
+
+o32 = {{0, 0, 0, 0, 0, 0, 7, -31, -7}}
+
+o32 : List
+
+-------------------------
+----Large Example #2-----
+-------------------------
+
+Here we have an example where dimDegViaBezout sometimes quickly gives the right answer, but more frequently gives the wrong answer.
+The best performance is actually via the non homogeneous version of the function in this case.
+dimDegViaBezout usually also gives the wrong answer (probably due to saturation issues).
+The Macaulay2 dimension function does not terminate on this example.
+We run each version of the function 100 times.
+Note, in this case usually Strategy=>MultiplicationTable typically crashes, 90% of the time (so this might be a case where a try function would be good, or more aggressive saturation).
+
+p=101;
+(ZZ/p)[YY_1..YY_9, Degrees => {9:1}, Heft => {1}, MonomialOrder => VerticalList{MonomialSize => 32, GRevLex => {9:1}, Position => Up}, DegreeRank => 1];
+
+f1 =-2*YY_2*YY_6*YY_7*YY_8^2*YY_9^2-2*YY_4*YY_6^2*YY_7*YY_9^3-4*YY_1*YY_6*YY_7*YY_8*YY_9^3-20*YY_2*YY_6*YY_7*YY_8*YY_9^3+2*YY_4^2*YY_8^2*YY_9^3+2*YY_7^2*YY_8^2*YY_9^3-40*YY_1*YY_6*YY_7*YY_9^4-6*YY_2*YY_6*YY_7*YY_9^4+20*YY_3*YY_6*YY_7*YY_9^4-20*YY_4^2*YY_8*YY_9^4-20*YY_7^2*YY_8*YY_9^4+4*YY_7*YY_8^2*YY_9^4-2*YY_4^2*YY_9^5-2*YY_7^2*YY_9^5-40*YY_7*YY_8*YY_9^5+2*YY_8^2*YY_9^5-4*YY_7*YY_9^6-20*YY_8*YY_9^6-2*YY_9^7;
+f2 = -2*YY_4^7-40*YY_2*YY_4^5*YY_5-4*YY_3*YY_4^5*YY_5+8*YY_1^2*YY_4^3*YY_5^2+21*YY_1*YY_2*YY_4^3*YY_5^2-8*YY_1*YY_3*YY_4^3*YY_5^2-41*YY_1*YY_4^5*YY_6+4*YY_2*YY_4^5*YY_6-40*YY_3*YY_4^5*YY_6+40*YY_1^2*YY_4^3*YY_5*YY_6+4*YY_1*YY_2*YY_4^3*YY_5*YY_6-40*YY_1*YY_3*YY_4^3*YY_5*YY_6-4*YY_1*YY_3*YY_4^3*YY_6^2+4*YY_1^2*YY_4^4*YY_7-40*YY_1*YY_2*YY_4^4*YY_7+8*YY_1^3*YY_4^2*YY_5*YY_7+21*YY_1^2*YY_2*YY_4^2*YY_5*YY_7-8*YY_1*YY_3^2*YY_4^2*YY_5*YY_7-21*YY_2*YY_3^2*YY_4^2*YY_5*YY_7-20*YY_4^4*YY_5*YY_6*YY_7-40*YY_3*YY_4^2*YY_5^2*YY_6*YY_7+4*YY_3*YY_4^2*YY_5*YY_6^2*YY_7-8*YY_1*YY_4^3*YY_5*YY_7^2+40*YY_2*YY_4^3*YY_5*YY_7^2-8*YY_1*YY_3*YY_4*YY_5^2*YY_7^2+41*YY_1*YY_4^3*YY_6*YY_7^2-4*YY_2*YY_4^3*YY_6*YY_7^2-40*YY_1^2*YY_4*YY_5*YY_6*YY_7^2-4*YY_1*YY_2*YY_4*YY_5*YY_6*YY_7^2-19*YY_1*YY_3*YY_4*YY_5*YY_6*YY_7^2+4*YY_1*YY_3*YY_4*YY_6^2*YY_7^2-4*YY_1^2*YY_4^2*YY_7^3+40*YY_1*YY_2*YY_4^2*YY_7^3+20*YY_4^2*YY_5*YY_6*YY_7^3+40*YY_3*YY_5^2*YY_6*YY_7^3-4*YY_3*YY_5*YY_6^2*YY_7^3+2*YY_4^3*YY_7^4+4*YY_3*YY_4*YY_5*YY_7^4+40*YY_3*YY_4*YY_6*YY_7^4+40*YY_1^2*YY_4^4*YY_8+4*YY_1*YY_2*YY_4^4*YY_8+40*YY_4^4*YY_5^2*YY_8-21*YY_3*YY_4^2*YY_5^3*YY_8-2*YY_4^4*YY_5*YY_6*YY_8-12*YY_3*YY_4^2*YY_5^2*YY_6*YY_8-40*YY_3*YY_4^2*YY_5*YY_6^2*YY_8-21*YY_1*YY_4^3*YY_5*YY_7*YY_8-21*YY_1^2*YY_4*YY_5^2*YY_7*YY_8+8*YY_1*YY_2*YY_4*YY_5^2*YY_7*YY_8-42*YY_1*YY_3*YY_4*YY_5^2*YY_7*YY_8-16*YY_1*YY_3*YY_4*YY_5*YY_6*YY_7*YY_8-40*YY_1^2*YY_4^2*YY_7^2*YY_8-4*YY_1*YY_2*YY_4^2*YY_7^2*YY_8-40*YY_4^2*YY_5^2*YY_7^2*YY_8+21*YY_3*YY_5^3*YY_7^2*YY_8+2*YY_4^2*YY_5*YY_6*YY_7^2*YY_8+12*YY_3*YY_5^2*YY_6*YY_7^2*YY_8+40*YY_3*YY_5*YY_6^2*YY_7^2*YY_8-2*YY_4^5*YY_8^2-4*YY_3*YY_4^3*YY_5*YY_8^2-40*YY_3*YY_4^3*YY_6*YY_8^2-4*YY_4^2*YY_5^2*YY_7*YY_8^2-8*YY_3*YY_5^3*YY_7*YY_8^2+21*YY_3*YY_5^2*YY_6*YY_7*YY_8^2+2*YY_4^3*YY_7^2*YY_8^2+4*YY_3*YY_4*YY_5*YY_7^2*YY_8^2+40*YY_3*YY_4*YY_6*YY_7^2*YY_8^2-4*YY_1^2*YY_4^4*YY_9-4*YY_1*YY_3*YY_4^4*YY_9+40*YY_2*YY_3*YY_4^4*YY_9+4*YY_3^2*YY_4^4*YY_9-8*YY_1^3*YY_4^2*YY_5*YY_9+8*YY_1*YY_3^2*YY_4^2*YY_5*YY_9-8*YY_1*YY_4^2*YY_5^3*YY_9+8*YY_3*YY_4^2*YY_5^3*YY_9+40*YY_1^3*YY_4^2*YY_6*YY_9-40*YY_1*YY_3^2*YY_4^2*YY_6*YY_9-20*YY_4^4*YY_5*YY_6*YY_9-40*YY_1*YY_4^2*YY_5^2*YY_6*YY_9+19*YY_3*YY_4^2*YY_5^2*YY_6*YY_9-4*YY_3*YY_4^2*YY_5*YY_6^2*YY_9+2*YY_4^5*YY_7*YY_9+8*YY_1*YY_4^3*YY_5*YY_7*YY_9-40*YY_2*YY_4^3*YY_5*YY_7*YY_9+4*YY_3*YY_4^3*YY_5*YY_7*YY_9+8*YY_1*YY_3*YY_4*YY_5^2*YY_7*YY_9+8*YY_3^2*YY_4*YY_5^2*YY_7*YY_9-21*YY_1*YY_4^3*YY_6*YY_7*YY_9+40*YY_3*YY_4^3*YY_6*YY_7*YY_9-21*YY_1^2*YY_4*YY_5*YY_6*YY_7*YY_9-21*YY_1*YY_3*YY_4*YY_5*YY_6*YY_7*YY_9+40*YY_3^2*YY_4*YY_5*YY_6*YY_7*YY_9+4*YY_1^2*YY_4^2*YY_7^2*YY_9+4*YY_1*YY_3*YY_4^2*YY_7^2*YY_9-40*YY_2*YY_3*YY_4^2*YY_7^2*YY_9+4*YY_3^2*YY_4^2*YY_7^2*YY_9+4*YY_4^2*YY_5^2*YY_7^2*YY_9+20*YY_4^2*YY_5*YY_6*YY_7^2*YY_9-40*YY_3*YY_5^2*YY_6*YY_7^2*YY_9+4*YY_3*YY_5*YY_6^2*YY_7^2*YY_9-2*YY_4^3*YY_7^3*YY_9-4*YY_3*YY_4*YY_5*YY_7^3*YY_9-40*YY_3*YY_4*YY_6*YY_7^3*YY_9+20*YY_4^5*YY_8*YY_9+42*YY_1*YY_4^3*YY_5*YY_8*YY_9+40*YY_3*YY_4^3*YY_5*YY_8*YY_9+42*YY_1^2*YY_4*YY_5^2*YY_8*YY_9+21*YY_3^2*YY_4*YY_5^2*YY_8*YY_9+8*YY_1*YY_4^3*YY_6*YY_8*YY_9-4*YY_3*YY_4^3*YY_6*YY_8*YY_9+4*YY_3^2*YY_4*YY_5*YY_6*YY_8*YY_9+40*YY_4^2*YY_5^2*YY_7*YY_8*YY_9-21*YY_3*YY_5^3*YY_7*YY_8*YY_9-8*YY_3*YY_5^2*YY_6*YY_7*YY_8*YY_9-20*YY_4^3*YY_7^2*YY_8*YY_9-40*YY_3*YY_4*YY_5*YY_7^2*YY_8*YY_9+4*YY_3*YY_4*YY_6*YY_7^2*YY_8*YY_9+4*YY_1^2*YY_4^2*YY_8^2*YY_9+4*YY_3^2*YY_4^2*YY_8^2*YY_9+4*YY_4^2*YY_5^2*YY_8^2*YY_9+40*YY_4^2*YY_5*YY_6*YY_8^2*YY_9-2*YY_4^5*YY_9^2-8*YY_1*YY_4^3*YY_5*YY_9^2+4*YY_3*YY_4^3*YY_5*YY_9^2-8*YY_3^2*YY_4*YY_5^2*YY_9^2-20*YY_1*YY_4^3*YY_6*YY_9^2-40*YY_3^2*YY_4*YY_5*YY_6*YY_9^2-8*YY_1*YY_3*YY_4^2*YY_7*YY_9^2-4*YY_3^2*YY_4^2*YY_7*YY_9^2-8*YY_4^2*YY_5^2*YY_7*YY_9^2+41*YY_4^2*YY_5*YY_6*YY_7*YY_9^2-2*YY_4^3*YY_7^2*YY_9^2-4*YY_3*YY_4*YY_5*YY_7^2*YY_9^2-21*YY_1*YY_3*YY_4^2*YY_8*YY_9^2-40*YY_3^2*YY_4^2*YY_8*YY_9^2+2*YY_4^2*YY_5*YY_6*YY_8*YY_9^2-2*YY_4^3*YY_8^2*YY_9^2-4*YY_3*YY_4*YY_5*YY_8^2*YY_9^2+4*YY_1*YY_3*YY_4^2*YY_9^3+4*YY_4^2*YY_5^2*YY_9^3+20*YY_4^2*YY_5*YY_6*YY_9^3+2*YY_4^3*YY_7*YY_9^3+4*YY_3*YY_4*YY_5*YY_7*YY_9^3+20*YY_4^3*YY_8*YY_9^3+40*YY_3*YY_4*YY_5*YY_8*YY_9^3;
+f3 = -2*YY_1^7+20*YY_1^6*YY_2+2*YY_1^6*YY_3-20*YY_1^5*YY_2*YY_3-2*YY_1^5*YY_3^2+20*YY_1^4*YY_2*YY_3^2+2*YY_1^4*YY_3^3-20*YY_1^3*YY_2*YY_3^3-YY_1^4*YY_4^2*YY_5-20*YY_1^3*YY_2*YY_4^2*YY_5+YY_1^3*YY_3*YY_4^2*YY_5-2*YY_1^5*YY_5^2-20*YY_1^4*YY_2*YY_5^2-2*YY_1^3*YY_3^2*YY_5^2-20*YY_1^2*YY_2*YY_3^2*YY_5^2+2*YY_1^5*YY_4*YY_7+10*YY_1^4*YY_2*YY_4*YY_7-2*YY_1^4*YY_3*YY_4*YY_7+30*YY_1^3*YY_2*YY_3*YY_4*YY_7+2*YY_1^2*YY_2^2*YY_3*YY_4*YY_7-YY_1^2*YY_3*YY_4*YY_5^2*YY_7-10*YY_1*YY_2*YY_3*YY_4*YY_5^2*YY_7+YY_1^4*YY_5*YY_7^2+10*YY_1^3*YY_2*YY_5*YY_7^2+YY_1^3*YY_3*YY_5*YY_7^2+YY_1*YY_2^2*YY_3*YY_5*YY_7^2+40*YY_1^5*YY_4*YY_8+4*YY_1^4*YY_2*YY_4*YY_8-40*YY_1^4*YY_3*YY_4*YY_8-4*YY_1^3*YY_2*YY_3*YY_4*YY_8+10*YY_1^3*YY_4*YY_5^2*YY_8+10*YY_1^2*YY_3*YY_4*YY_5^2*YY_8+30*YY_1^4*YY_5*YY_7*YY_8-YY_1^3*YY_2*YY_5*YY_7*YY_8+10*YY_1^3*YY_3*YY_5*YY_7*YY_8-YY_1^2*YY_2*YY_3*YY_5*YY_7*YY_8+20*YY_1*YY_2^2*YY_3*YY_5*YY_7*YY_8-2*YY_1^4*YY_5*YY_8^2+20*YY_1^3*YY_2*YY_5*YY_8^2+2*YY_1^3*YY_3*YY_5*YY_8^2-20*YY_1^2*YY_2*YY_3*YY_5*YY_8^2-2*YY_1^5*YY_4*YY_9+10*YY_1^4*YY_2*YY_4*YY_9+2*YY_1^3*YY_2^2*YY_4*YY_9+2*YY_1^4*YY_3*YY_4*YY_9-10*YY_1^3*YY_2*YY_3*YY_4*YY_9-YY_1^3*YY_4*YY_5^2*YY_9-10*YY_1^2*YY_2*YY_4*YY_5^2*YY_9-YY_1^4*YY_5*YY_7*YY_9-20*YY_1^3*YY_2*YY_5*YY_7*YY_9+YY_1^2*YY_2^2*YY_5*YY_7*YY_9-YY_1^3*YY_3*YY_5*YY_7*YY_9-YY_1*YY_2^2*YY_3*YY_5*YY_7*YY_9-10*YY_1^4*YY_5*YY_8*YY_9+YY_1^3*YY_2*YY_5*YY_8*YY_9+20*YY_1^2*YY_2^2*YY_5*YY_8*YY_9+10*YY_1^3*YY_3*YY_5*YY_8*YY_9+YY_1^2*YY_2*YY_3*YY_5*YY_8*YY_9+10*YY_1^3*YY_2*YY_5*YY_9^2-YY_1^2*YY_2^2*YY_5*YY_9^2;
+f4 = -4*YY_7*YY_8^2*YY_9^4-2*YY_4^2*YY_9^5-2*YY_7^2*YY_9^5+21*YY_7*YY_8*YY_9^5-2*YY_9^7;
+f5 = 2*YY_1^2*YY_3^5+2*YY_3^7+2*YY_3^5*YY_4*YY_7+40*YY_3^5*YY_4*YY_8+YY_1*YY_3^4*YY_4*YY_9-YY_3^5*YY_4*YY_9+YY_3^4*YY_5*YY_7*YY_9+10*YY_3^4*YY_6*YY_7*YY_9+20*YY_3^4*YY_5*YY_8*YY_9-2*YY_3^4*YY_6*YY_8*YY_9-YY_3^4*YY_5*YY_9^2-10*YY_3^4*YY_6*YY_9^2+YY_3^3*YY_7^2*YY_9^2+40*YY_3^3*YY_7*YY_8*YY_9^2-4*YY_3^3*YY_8^2*YY_9^2+YY_1*YY_3^2*YY_7*YY_9^3-YY_3^3*YY_7*YY_9^3+20*YY_1*YY_3^2*YY_8*YY_9^3-20*YY_3^3*YY_8*YY_9^3-YY_1*YY_3^2*YY_9^4;
+
+i93 : J = ideal(f1,f2,f3, f4);
+
+i103 : time apply(100, t -> dimViaBezout(J))
+     -- used 222.619 seconds
+
+o103 = {6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6}
+
+i104 : time apply(100, t -> dimViaBezout(J, Homogeneous=>false))
+     -- used 62.712 seconds
+
+o104 = {6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6,
+       --------------------------------------------------------------------------------                                                        
+       6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6}
+
+i106 : time apply(100, t -> (dimDegViaBezout(J))#0)
+     -- used 323.897 seconds
+
+o106 = {5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 6, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5,
+       --------------------------------------------------------------------------------                                                        
+       5, 5, 5, 5, 6, 5, 5, 5, 5, 5, 5, 5, 5, 6, 5, 6, 5, 5, 5, 5, 5, 6, 5, 5, 5, 5, 5,
+       --------------------------------------------------------------------------------                                                        
+       5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 6, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5,
+       --------------------------------------------------------------------------------                                                        
+       5, 5, 5, 5, 5, 5, 5, 5, 6, 5, 5, 5, 5, 5, 5, 6, 5, 5, 5, 6}
+
+i111 : time randomPoints(J, Strategy=>MultiplicationTable)
+RandomRationalPoints.m2:847:17:(3):[12]: error: expected a polynomial ring
+RandomRationalPoints.m2:847:17:(3):[12]: --entering debugger (type help to see debugger commands)                                              
+RandomRationalPoints.m2:847:17-847:79: --source code:                                                                                          
+            m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));
+
+ii112 : break
+     -- used 2.37748 seconds
+
+i115 : time randomPoints(J, Strategy=>MultiplicationTable)
+     -- used 2.29754 seconds
+
+o115 = {{0, 17, -17, -10, -35, 37, 0, 34, 1}}
+
+o115 : List
+
+i119 : time randomPoints(J)
+     -- used 2.2024 seconds
+
+o119 = {{-20, 31, 26, -12, -16, -42, 12, -8, 0}}
+
+o119 : List
+
+i135 : time randomPoints(J, Codimension=>3)
+     -- used 0.13706 seconds
+
+o135 = {{-14, -17, -25, -37, 23, -22, 12, 10, 0}}
+
+o135 : List
+
+i136 : time randomPoints(J, Strategy=>MultiplicationTable, Codimension=>3)
+RandomRationalPoints.m2:847:17:(3):[12]: error: expected a polynomial ring
+RandomRationalPoints.m2:847:17:(3):[12]: --entering debugger (type help to see debugger commands)                                              
+RandomRationalPoints.m2:847:17-847:79: --source code:                                                                                          
+            m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));
 
@@
 
-In this example, the primary speed limitation still appears to be the dimension computation.
+In this particular example, dim is much faster than dimViaBezout, or even the referee's original dimDegViaBezout.
+Notice that dimDegViaBezout gives the wrong answer more than half the time in this example, most likely due to saturating away things that it should not...
 
@@
 
-i55 :  time randomPoints(J2, Strategy=>MultiplicationTable)
+i55 : time dim J2
+     -- used 0.0122921 seconds
+
+o55 = 3
+
+i56 : time dimDegViaBezout(J2)
+     -- used 0.723174 seconds                                                                                                                  
+o56 = (2, 54)
+
+i57 : time dimViaBezout(J2)
+     -- used 2.09779 seconds
+
+o57 = 3
+
+i60 :  time randomPoints(J2, Strategy=>MultiplicationTable)
 RandomRationalPoints.m2:840:17:(3):[12]: error: expected a polynomial ring
@@
 
-i56 : time randomPoints(J2, Strategy=>Default)
+i61 : time randomPoints(J2, Strategy=>Default)
      -- used 4.2498 seconds

### 2021-04-24 FastLinAlg/Paper/RandomRationalPoints.tex
paper updates

@@
 For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be provided.
+    \item {\tt Homogeneous}: 
+        By default, the function uses one of two algorithms depending on whether the ideal is homogeneous.
+        There are homogeneous examples where the non-Homogeneous method is faster.
+Setting this to {\tt false} will force non-homogeneous computation even in homogeneous examples.
+Setting it to true will force homogeneous computation, which may behave unexpectedly in non-homogeneous examples.
 \end{itemize}

### 2021-04-29 FastLinAlg/Paper/ChangesMade_ResponseToReviewers.txt
Comments File Copied to Paper Folder as well

--- 2021-04-24 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
@@
-We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples.
-We would be very happy to invite the reviewer to be a coauthor to the source code (and also for the paper, if desired).
+We thank the reviewers for their helpful suggestions, comments preliminary code and examples.
+We would very much like to invite the reviewer to be a coauthor to the source code (and also for the paper, if desired).
+In particular, there are some issues we think we would benefit with working with them on (see below).
 
@@
 
-e) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.
+e) Page 5 dimViaBezout added with example 3.1 and paragraph following it.
+
+f)  Page 5, Remark 2.2.2 added explaining how generic projection works and why we aren't using it currently.
 

### 2021-07-14 FastLinAlg/Paper/RandomRationalPoints.tex
Updated the paper partially to the new version of the code.

--- 2021-04-24 FastLinAlg/Paper/RandomRationalPoints.tex
@@
 \begin{document}
-\title{{RandomRationalPoints} package for \emph{Macaulay2}}
+\title{{RandomPoints} package for \emph{Macaulay2}}
 \author{Sankhaneel Bisui}
@@
 \begin{abstract}
-  We present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
+  We present {\tt RandomPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
 We provide different strategies such as linear intersection and projection to obtain such points.
@@
 
-\keywords{RandomRationalPoints, Macaulay2}
+\keywords{RandomPoints, Macaulay2}
 % \thanks{Martinova was supported by a University of Utah Mathematics REU fellowship and by the University of Utah ACCESS program.  
@@
         \setlength{\itemsep}{5pt}
-    \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection}, {\tt HybridProjectionIntersection} or {\tt MultiplicationTable}.
+    \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce} or {\tt LinearIntersection} \\ .
 
@@
 It begins begins with brute force and moves to linear intersections with particularly simple linear forms, and gradually ramps up the randomness.
-If {\tt ProjectionAttempts} is set to a value bigger than zero (it is set to zero by default), then the strategy alternates between generic projects and hybrid projection / intersections.
-For additional discussion, see \cref{rem.ProjectingToAHypersurfaceFirst}.
     \item {\tt BruteForce} simply tries random points and sees if they are on the variety.
 
-    \item {\tt GenericProjection} projects to a hypersurface, via 
-        {\tt projectionToHypersurface} and then uses the {\tt BruteForce} strategy.
-Finally, it pulls back back the points that were found to the original variety (via a linear intersection).
-See \cref{rem.ProjectingToAHypersurfaceFirst}.
+    \item {\tt LinearIntersection} intersects with a random linear space.
+Within this strategy, one can also specify the option {\tt DecompositionStrategy}.
+Valid values are {\tt Decompose} and {\tt MultiplicationTable}, the latter of which is currently only implemented for homogeneous ideals.
 
-    \item {\tt LinearIntersection} intersects with an appropriately random linear space.
-
-    \item {\tt HybridProjectionIntersection} does a generic projection, followed by a linear intersectio with that hypersurface.
-Finally, it pulls back back the points that were found to the original variety (via a linear intersection).
-See \cref{rem.ProjectingToAHypersurfaceFirst}.
-    \item {\tt MultiplicationTable} works for homogeneous ideals only.
-It cuts down with a linear space to a zero-dimensional projective scheme and computes how the last two variables act on the quotient ring truncated at the regularity.
-%In case of an absolutely irreducible $I$ over the field, we will find with  high probability a point, since the determinant will have a linear factor in at least 50\% of the cases.
+    The point is, after intersecting the linear space and obtaining an ideal defining a set of (possibly thickened) points, we need to find the minimal associated primes.
+By default we use Macaulay2's built-in {\tt decompose} command.
+We also have implemented a {\tt MultiplicationTable} algorithm, as provided by Frank Olaf-Schreyer, which utilizes the fact that we can compute the action of a variable on the residue fields of these points in more than one way.
+This method is frequently faster for smaller numbers of variables.
     \end{itemize}
@@
 
-i6 : randomPoints(J,Strategy=>GenericProjection)
+i6 : randomPoints(J,Strategy=>LinearIntersection)
     -- used 0.0146932 seconds
@@
     \vspace{-1em}
-    \item[\tt ProjectionAttempts => ZZ]
-
-    When calling the strategy {\tt GenericProjection} or \\
-    {\tt HybridProjectionIntersection} from {\tt randomPoints}, this option denotes the number of trials before giving up.  This option is also passed to {\tt randomPoints} by other functions.
-
-    \item[\tt MaxCoordinatesToReplace => ZZ] This is used for fine tuning the {\tt LinearIntersection} or {\tt GenericProjection} strategies, see \Cref{helper}.
-    \item[\tt Codimension => ZZ] One can provide the codimension so it is not recomputed.
-This is used in the {\tt GenericProjection} and {\tt LinearIntersection} strategies, see \Cref{projectionfunctions}.
 
@@
     Intersection with a general linear space will naturally find scheme theoretic points that are not rational over the base field.  
-    Setting {\tt ExtendField => true} will tell the function that such points are valid.
-Setting {\tt ExtendField => false} will tell the function ignore such points.
+    Setting {\tt ExtendField => true} will tell the function that such points are valid.  Setting {\tt ExtendField => false} will tell the function ignore such points.  
+    In fact, {\tt ExtendField => true} will also tell Macaulay2 to use linear spaces defined over a field extension, which can improve randomness properties.
     This sometimes can slow computation, and other times can substantially speed it up when the variety has few rational points.
-In some cases, points over extended fields may also have better randomness properties for applications.
+For some applications, points over extended fields may also have better randomness properties.
 
-    \item[\tt IntersectionAttempts => ZZ]
+    \item[\tt Replacement => Monomial, Binomial, Trinomial, Full]
+        When intersecting with a random linear space, it is frequently much faster to use a linear space defined by relatively sparse equations (ie, equations that do not involve all variables).
+Specifying {\tt Monomial} will mean linear forms such as $ax + b$ are used (for constants $a$ and $b$), involving only one variable.
+Binomial means forms like $ax + by + c$, using two variables.
+Trinomial means forms like $ax + by + cz + d$.
+{\tt Full} means all variables will have coefficients.
+Alternately, one can specify that a certain number of monomial, binomial, trinomial and full forms are used by passing a list, see the documentation for details.
 
-    This option is used by {\tt randomPoints} in some strategies to determine the maximum number of attempts to intersect with a linear space when looking for random rational points.
-%Other functions pass this option through to {\tt randomPoints}.
+    \item[\tt DimensionFunction => Function]
+        Our current implementation does not need to know the dimension of $V(I)$.
+However, there are places where we try to verify the dimension.
+You can pass this function {\tt dim} (the default) or our probabilistic {\tt dimViaBezout} or any other dimension function you might prefer.
 
@@
     \end{example}%
-%
-    %\vspace{1em}
-    \item[\tt MaxCoordinatesToTrivialize]
-
-    When calling {\tt randomPoints} and performing an intersection with a linear space, this is the number of defining equations of the linear space of the form $x_i - a_i$.
-Having a large number of these will provide faster intersections.
 
@@
 
-    \subsection{Comments on performance}
+    \subsection{Comments on performance and implementation}
 
@@
 However, if the field size is larger, {\tt BruteForce} will perform poorly.
-Even for some simple examples, it could not provide any rational points if the number of trials is not enough.
+Even for some simple examples, it could not provide any rational points if the number of trials is not large enough.
 Other strategies work differently on different examples, and the same strategy can sometimes work very quickly even if it typically works very slowly.
 
-    Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck in some cases.
+    The current version of the {\tt LinearIntersection} strategy no longer computes the dimension of the algebraic set.
+Instead, it first finds a point defined by linear equations.
+If the point is on the algebraic set, we are done.
+If not, we throw away one of the forms and so now have a line and we see if this line intersects our algebraic set.
+We continue in this way until we find a point.
+This appears to avoid a number of bottlenecks in our previous implementation since Macaulay2 is very fast at identifying when a linear space and a variety do \emph{not} intersect.
+
+    %Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck in some cases.
 If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}).
@@
 
-    However, while in a number of examples, doing this generic projection first can speed up computations (as opposed to simply intersecting with a linear space, since we have identified a linear space initially), there are numerous cases where computing this hypersurface $h(X)$ can be extremely slow.
+    However, in our experience, doing this generic projection first yields slower results.
+First, one has to compute the dimension.
+Furthermore, there are numerous cases where computing this hypersurface $h(X)$ can be extremely slow.
 This particularly appears in cases when one is computing successive minors to identify the locus where some variety is nonsingular.
-Thus, by default, we disable the functionality to project to a hypersurface in the default strategy.
+Thus we no longer have an option to project to a hypersurface first.
 
     On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (i.e., with many linear or binomial terms).
-Such simple linear spaces are the ones considered in {\tt randomKRationalPoint} for instance.
-In practice, this seems give approximately the same performance as projecting to a hypersurface, without the chance of the code getting hung up on the generic projection.
+Such simple linear spaces are the ones implicitly considered in {\tt randomKRationalPoint} for instance.
+In practice, our approach seems to give at superior performance compared to projecting to a hypersurface, without the chance of the code getting hung up on the generic projection or dimension computations.
 \end{remark}
 
-\section{Two helper functions: {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
+\section{Two related functions: {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
 
-\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, significantly slows down speed.
-Instead, we compute the dimension of $V(I)$ probabilistically using {\tt dimViaBezout}.
+\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, significantly slows down some computations.
+While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
 This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.
@@
 If there was no intersection, this function tries a $2$-dimensional linear space, and so on.
-This speeds up many computations.
+This can speed up a number of computations.
 The function also takes in optional inputs as described below:
@@
 The user may set the {\tt MinimumFieldSize} to ensure that the field being worked over is big enough.
-For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be provided.
+For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be returned to the user.
+    \item {\tt Homogeneous}: 
+    If the ideal is homogeneous, we can use homogeneous linear spaces to compute dimension, which has speed advantages.
+        %By default, the function uses one of two algorithms depending on whether the ideal is homogeneous.
+        %There are homogeneous examples where the non-Homogeneous method is faster.
+Setting this to {\tt false} will force non-homogeneous computation even in homogeneous examples.
+Setting it to true will force homogeneous computation, which may behave unexpectedly in non-homogeneous examples.
 \end{itemize}
@@
 
-As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using the output from the above function (or by existing command {\tt codim} beforehand), and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}.
-Alternatively, one can simply rely on any of the strategies for {\tt randomPoints} (see \Cref{strategydetails}) as they compute the dimension of $V(I)$ using {\tt dimViaBezout}, which significantly accelerates the computations.
-
 \subsection*{\tt randomCoordinateChange:}
@@
 \begin{itemize}
-\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.
-If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications.
-If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
+\item {\tt Replacement}: This works like the {\tt Replacement} option for {\tt RandomPoints}.
 \item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
@@
 
-\section{Other helper functions: {\tt genericProjection}, {\tt projectionToHypersurface}}\label{projectionfunctions}
+\section{Other functions: {\tt genericProjection}, {\tt projectionToHypersurface}}\label{projectionfunctions}
 
@@
 There is some similar functionality computing partial Jacobian ideals obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}, see \cite{FastLinAlgSource}.
-That package now uses the functionality contained here in {\tt RandomRationalPoints} in some of its functions.
+That package now uses the functionality contained here in {\tt RandomPoints} in some of its functions.
 

### 2021-07-15 FastLinAlg/Paper/RandomRationalPoints.tex
Changes to the paper.

@@
 The authors are also grateful to the reviewers for suggesting and providing preliminary codes to speed up computations, thereby improving the efficacy of the package substantially.
+Frank Olaf-Schreyer is also an author on this package, as he provided code related to the {\tt MultiplicationTable} decomposition strategy and suggested using a probabilistic approach to compute dimension.
 
@@
     \item {\tt LinearIntersection} intersects with a random linear space.  
-Within this strategy, one can also specify the option {\tt DecompositionStrategy}.
-Valid values are {\tt Decompose} and {\tt MultiplicationTable}, the latter of which is currently only implemented for homogeneous ideals.
-
-    The point is, after intersecting the linear space and obtaining an ideal defining a set of (possibly thickened) points, we need to find the minimal associated primes.
-By default we use Macaulay2's built-in {\tt decompose} command.
-We also have implemented a {\tt MultiplicationTable} algorithm, as provided by Frank Olaf-Schreyer, which utilizes the fact that we can compute the action of a variable on the residue fields of these points in more than one way.
-This method is frequently faster for smaller numbers of variables.
     \end{itemize}
@@
 
-i4 : randomPoints(J)
-    -- used 0.0181549 seconds
+i4 : time randomPoints(J,Strategy=>BruteForce, PointCheckAttempts=>10)
+        -- used 0.00186098 seconds
 
-o4 = {{32, 37, -30}}
+o4 = {}
 
@@
 
-i5 : time randomPoints(J,Strategy=>BruteForce)
-    -- used 0.0146932 seconds
+i5 : time randomPoints(J)
+        -- used 0.0205099 seconds
 
-o5 = {}
+o5 = {{-1, 0, -1}}
 
@@
 
-i6 : randomPoints(J,Strategy=>LinearIntersection)
-    -- used 0.0146932 seconds
+i6 : time randomPoints(J, Strategy=>LinearIntersection)
+        -- used 0.0334881 seconds
 
-o6 = {{2, -30, -8}}
+o6 = {{0, 10, 48}}
 \end{verbatim}
@@
 For some applications, points over extended fields may also have better randomness properties.
+
+    \item[\tt DecompositionStrategy]  Within the {\tt LinearIntersection} strategy, one can also specify the option {\tt DecompositionStrategy}.
+Valid values are {\tt Decompose} and {\tt MultiplicationTable}, the latter of which is currently only implemented for homogeneous ideals.
+
+    The point is, after intersecting the linear space and obtaining an ideal defining a set of (possibly thickened) points, we need to find the minimal associated primes.
+By default we use Macaulay2's built-in {\tt decompose} command.
+We also have implemented a {\tt MultiplicationTable} algorithm, as provided by Frank Olaf-Schreyer, which utilizes the action of a variable on the residue fields of these points computed in more than one way.
+This method is frequently faster for rings with smaller numbers of variables.
 
@@
 \begin{verbatim}    
-i7 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
-    -- used 0.84679 seconds
+i7 : time randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 10000)
+-- used 1.16294 seconds
 
-o7 = {{-39, -43, 28}}
+o7 = {{-43, 25, 29}}
     \end{verbatim}%        
@@
     \subsection{Comments on performance and implementation}
+    \label{subsec.CommentOnPerformanceAndImplementation}
 
-    When working over very small fields, frequently {\tt BruteForce} is most efficient.
+    When working over very small fields, especially with hypersurfaces, frequently {\tt BruteForce} is most efficient.
 This is not surprising as there may not be many points to check.
@@
 i4 : time randomPoints(I, Strategy => BruteForce, PointCheckAttempts => 20000)
-    -- used 0.0102075 seconds
+        -- used 0.00311884 seconds
 
-o4 = {{-2, -1, 3, -2, -2, 2, 0, -2, -2, -2}}
+o4 = {{-1, -1, 0, 2, 2, -2, -2, -3, -3, -3}}
 
@@
 i5 : time randomPoints(I, Strategy => Default)
-    -- used 0.085741 seconds
+        -- used 0.081349 seconds
 
-o5 = {{1, -2, 2, 2, -2, 0, 0, -1, -1, 3}}
+o5 = {{3, 0, 3, 3, 2, -2, 1, -1, 3, 1}}
 \end{verbatim}
@@
 i8 : time randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 2000000)
-    -- used 25.3479 seconds
+        -- used 17.7988 seconds
 
-o8 = {{-36, 26, -63, -7, -48, 78, -44, -22, 105, 39}}
+o8 = {{15, 67, -27, -103, 56, 66, -23, 28, -50, 13}}
+
+o8 : List
 
 i9 : time randomPoints(J, Strategy => Default)
--- used 0.102918 seconds
+        -- used 0.0864013 seconds
 
-o9 = {{0, -42, -53, -33, -65, -27, -105, -37, 99, 64}}
-
-o9 : List
-
-i10 : time randomPoints(J, Strategy => LinearIntersection)
--- used 0.0169781 seconds
-
-o10 = {{-51, -11, -34, -33, -65, 19, -86, 50, -44, -97}}
-
-o10 : List
-
-i11 : time randomPoints(J, Strategy => GenericProjection)
--- used 0.364902 seconds
-
-o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
+o9 = {{0, 0, 0, 0, 34, 76, 51, 0, 1, 0}}
     \end{verbatim}
@@
     \end{example}
-To enable speed ups as indicated in the above discussion, the {\tt ProjectionAttempts} has been set to $0$ in the {\tt Default} strategy so that {\tt GenericProjection} is not opted for unless specified by user.
+
 \begin{example}
@@
         \begin{quote}
+            {\small
 \begin{verbatim}
-i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
--- used 0.0612306 seconds
+i11 : time randomPoints(J, Strategy => Default, ExtendField => true)
+            -- used 0.144332 seconds
 
-               5      4      3      2                                      5  
-o12 = {{54, 14a  - 48a  - 56a  + 56a  - 13a + 48, -101, 56, -42, -27, - 26a  -
-    -------------------------------------------------------------------------
-       4      3      2
-    43a  - 70a  - 96a  + 50a - 103, -29, 0, -37}}
+              3      2                      3      2                        
+o11 = {{0, - a  + 62a  - 47a - 76, 0, 0, 13a  - 18a  + 63a - 31, 0, 0, 
+             3      2                 3      2
+        - 20a  - 82a  + 35a - 19,  55a  - 64a  - 8a - 50, 1}}
+
+i12 : coefficientRing ring first first o11
+
+o12 = GF 1982119441
+
+i13 : log_211 1982119441
+
+o13 = 4
 \end{verbatim}
+            }
     \end{quote}
         }}
+In this case, we found a degree 4 point.
     \end{example}
 
-\begin{remark}[Finding a point]
+\begin{remark}[Comments on the probability of finding a point]
 	In the case of an absolutely irreducible hypersurface in $\bA^n_{\bF_q}$ (defined by $f$ say), there is significant discussion in the literature estimating lower bounds of number of rational points (see for instance, \cite{RationalKPointsSource,lang1954number,ghorpade2002number,cafure2006improved}) all of which point to the fact that there is ``good probability" of finding a rational point in this case when we intersect with a random line.
@@
 
-       Of course, there are certainly examples schemes over $\bF_q$ with no rational points at all, even for curves in $\bA^2$.
+       Of course, there are schemes over $\bF_q$ with no rational points at all, even for plane curves.
 \end{remark}
@@
 In a number of existing algorithms, one first does a generic (or even not very generic) projection $h : \bA^n \to \bA^m$ and so that $h(X)$ is a hypersurface (at least set theoretically).
-Then we find a point $x \in h(X)$ (say as above), and compute $h^{-1}(\{ x \})$ which is a linear space in $\bA^n$ that typically intersects $X$ in a rational point.
+Then one finds a point $x \in h(X)$ (say as above), and computes $h^{-1}(\{ x \})$ which is a linear space in $\bA^n$ that typically intersects $X$ in a rational point.
 For example, this is done in {\tt randomKRationalPoint} in core Macaulay2.
-Note that projecting to a hypersurface \emph{still is performing an intersection with a linear space}, but it just tries to choose the linear space more intelligently.
+Note that projecting to a hypersurface \emph{still is performing an intersection with a linear space}, since $h^{-1}(\{ x \})$ is linear, but it tries to choose the linear space intelligently.
 
@@
 First, one has to compute the dimension.
-Furthermore, there are numerous cases where computing this hypersurface $h(X)$ can be extremely slow.
+There are also numerous cases where computing this hypersurface $h(X)$ can be quite slow.
 This particularly appears in cases when one is computing successive minors to identify the locus where some variety is nonsingular.
-Thus we no longer have an option to project to a hypersurface first.
 
     On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (i.e., with many linear or binomial terms).
-Such simple linear spaces are the ones implicitly considered in {\tt randomKRationalPoint} for instance.
-In practice, our approach seems to give at superior performance compared to projecting to a hypersurface, without the chance of the code getting hung up on the generic projection or dimension computations.
+Such simple linear spaces are the ones implicitly considered in {\tt randomKRationalPoint} for instance since that generic projection is so simple.
+In practice, our approach seems to give at least as good performance compared to projecting to a hypersurface, without the chance of the code getting hung up on the generic projection or dimension computations.
+We also do successive intersections in a way that avoids computing the dimension as described above in \Cref{subsec.CommentOnPerformanceAndImplementation}.
 \end{remark}
 
-\section{Two related functions: {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
+\section{Other functions {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
 
-\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, significantly slows down some computations.
+\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, is a significant bottleneck.
 While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
@@
 	\item {\tt DimensionIntersectionAttempts}: 
-	This integer input denotes the number of linear spaces to try before moving to the next dimension.
+        Our function actually estimates dimension several times and then averages the result (rounding down) since we tend to overestimate the dimension by the above.
+By default it does this 3 times unless the {\tt Homogeneous} flag is set, in which case it is done 5 times.
 	\item {\tt MinimumFieldSize}:
 	If the ambient field is smaller than this integer value, it will automatically be replaced with an extension field.
+For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be returned to the user.
 The user may set the {\tt MinimumFieldSize} to ensure that the field being worked over is big enough.
-For instance, there are relatively few linear spaces over a field of characteristic $2$, and this can cause incorrect results to be returned to the user.
+If this is not set, the program tries to choose a reasonable minimum field size based on the ambient ring.
     \item {\tt Homogeneous}: 
-    If the ideal is homogeneous, we can use homogeneous linear spaces to compute dimension, which has speed advantages.
+    If the ideal is homogeneous, we can use homogeneous linear spaces to compute dimension.
+However, in practice, this appears slower than the non-homogeneous strategy due to the fact that we must saturate the result.
+Hence this is turned off by default.
         %By default, the function uses one of two algorithms depending on whether the ideal is homogeneous.
@@
 				\begin{verbatim}
-				i3 : kk=ZZ/nextPrime 10^2;
+i2 : S = ZZ/101[y_0..y_9];
 
-				i4 : S=kk[y_0..y_14];
+i3 : I=ideal random(S^1,S^{-2,-2,-2,-3})+(ideal random(2,S))^2;
 
-				i5 : I=minors(2,random(S^3,S^{5:-1}));
+o3 : Ideal of S
 
-				o5 : Ideal of S
+i4 : time dimViaBezout I
+        -- used 0.837359 seconds
 
-				i6 : elapsedTime dimViaBezout(I)
-				-- 0.303629 seconds elapsed
+o4 = 5
 
-				o6 = 7
+i5 : time dim I
+        -- used 36.8496 seconds
 
-				i7 : elapsedTime dim I
-				-- 1.47916 seconds elapsed
+o5 = 5
 
-				o7 = 7
+i6 : time dimViaBezout(I, DimensionIntersectionAttempts=>1)
+        -- used 0.280803 seconds
+
+o6 = 5
 				\end{verbatim}
@@
 	}}
+As you can see doing a single intersection attempt is about three times faster, and it usually gives the right answer (more than 99\% of the time in this particular example, but in others doing the computation in triplicate avoids returning incorrect answers).
 \end{example}
@@
 \item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
-  This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
+  %This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
 \item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
@@
 If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
+The same approach was used in the {\tt point} command in the {\tt Cremona} package \cite{CremonaSource}.
 
@@
 \begin{description}
-\item[\tt Strategy => Symbol] To specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
+\item[\tt Strategy => Symbol] Specify a strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
 
@@
 \item[\tt MinorPointAttempts => ZZ] 
-  This controls how many points at which to check the rank of the matrix.
+  This controls how many points at which to check the rank.
 \end{description}
@@
 
-		i5 : i = 0;
-
-		i6 : J = I;
+i5 : i = 0; J = I;
 
@@
                                 J = extendIdealByNonZeroMinor(4, M, J)));
-		-- 0.903328 seconds elapsed
+           -- 0.640164 seconds elapsed
 

### 2021-07-15 FastLinAlg/Paper/RandomRationalPoints.tex
Improvements to the paper.

@@
 Valid values are {\tt Decompose} and {\tt MultiplicationTable}, the latter of which is currently only implemented for homogeneous ideals.
+Additionally, the {\tt Default} strategy switches back and forth between {\tt Decompose} and {\tt MultiplicationTable} for homogeneous ideals.
+Setting this to {\tt Decompose} in the default strategy will force only {\tt Decompose} to be used.
 
@@
 This method is frequently faster for rings with smaller numbers of variables.
+
+    \item[\tt Homogeneous]  Setting this true specifies that the origin (corresponding to the irrelevant ideal) is not a valid point.
 
@@
     When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of trials for brute force point checking.
+It also controls how many linear spaces to simultaneously study in the {\tt LinearIntersection} strategy.
 
@@
 \begin{example} We illustrate the speed difference in this example.
-	{\color{blue}{
+	{\color{blue}{\small
 			\begin{quote}
@@
 	o6 = | x-3 z-5 -x+3y-4z+2 |
-
-	i7 : matrix randomCoordinateChange(R, MaxCoordinatesToReplace => 1, 
-	Replacement => Binomial)
-
-	o7 = | y x+4z z |
 \end{verbatim}
@@
 \end{example}\vspace{-1em}
+\noindent
 Alternatively, instead of {$I$}, we may pass it a quotient ring.
 It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.
-It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.
+%It is the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.
 
@@
 
-The user may also provide the following additional information:
+Besides the options from {\tt randomPoints} which are automatically passed to that function, the user may also provide the following additional information:
 
 \begin{description}
-\item[\tt Strategy => Symbol] Specify a strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).
-
-\item[\tt Verbose => Boolean]
-  Set the option {\tt Verbose => true} to turn on verbose output.  This may be useful in debugging or in determining why a computation is running slowly.
-
-\item[\tt Homogeneous => Boolean] (see \Cref{helper})
-
     \item[\tt MinorPointAttempts => ZZ] 

### 2021-07-15 FastLinAlg/Paper/RandomRationalPoints.tex
Fixed typo

@@
 The authors are also grateful to the reviewers for suggesting and providing preliminary codes to speed up computations, thereby improving the efficacy of the package substantially.
-Frank Olaf-Schreyer is also an author on this package, as he provided code related to the {\tt MultiplicationTable} decomposition strategy and suggested using a probabilistic approach to compute dimension.
+Frank-Olaf Schreyer is also an author on this package, as he provided code related to the {\tt MultiplicationTable} decomposition strategy and suggested using a probabilistic approach to compute dimension.
 
@@
 By default we use Macaulay2's built-in {\tt decompose} command.
-We also have implemented a {\tt MultiplicationTable} algorithm, as provided by Frank Olaf-Schreyer, which utilizes the action of a variable on the residue fields of these points computed in more than one way.
+We also have implemented a {\tt MultiplicationTable} algorithm, as provided by Frank-Olaf Schreyer, which utilizes the action of a variable on the residue fields of these points computed in more than one way.
 This method is frequently faster for rings with smaller numbers of variables.

### 2021-07-16 FastLinAlg/Paper/RandomRationalPoints.tex

@@
 Valid values are {\tt Decompose} and {\tt MultiplicationTable}, the latter of which is currently only implemented for homogeneous ideals.
-Additionally, the {\tt Default} strategy switches back and forth between {\tt Decompose} and {\tt MultiplicationTable} for homogeneous ideals.
-Setting this to {\tt Decompose} in the default strategy will force only {\tt Decompose} to be used.
-
 The point is, after intersecting the linear space and obtaining an ideal defining a set of (possibly thickened) points, we need to find the minimal associated primes.
@@
 This method is frequently faster for rings with smaller numbers of variables.
+
+    The {\tt Default} strategy switches back and forth between {\tt Decompose} and {\tt MultiplicationTable} for homogeneous ideals (starting with one the function thinks will be fastest).
+Setting this to {\tt Decompose} in the default strategy will force only {\tt Decompose} to be used; setting it to {\tt MultiplicationTable} will force only {\tt MultiplicationTable} to be used (if the ideal is homogeneous).
 
@@
 {\tt Full} means all variables will have coefficients.
-Alternately, one can specify that a certain number of monomial, binomial, trinomial and full forms are used by passing a list, see the documentation for details.
+%Alternately, one can specify that a certain number of monomial, binomial, trinomial and full forms are used by passing a list, see the documentation for details.
 
@@
         Our current implementation does not need to know the dimension of $V(I)$.
-However, there are places where we try to verify the dimension.
+However, there are places where we try to verify the dimension of an ideal before we decompose the ideal.
 You can pass this function {\tt dim} (the default) or our probabilistic {\tt dimViaBezout} or any other dimension function you might prefer.
@@
 We continue in this way until we find a point.
-This appears to avoid a number of bottlenecks in our previous implementation since Macaulay2 is very fast at identifying when a linear space and a variety do \emph{not} intersect.
+This appears to avoid a number of bottlenecks in our previous implementation since Macaulay2 is relatively fast at identifying when a linear space and a variety do \emph{not} intersect.
 
@@
 \begin{example}
-        Finally, we allow our functions to extend our field.
+        Finally, we can allow our functions to extend our field.
         {{\small\color{blue}
@@
 
--- {\tt genericProjection(n, I)},
+-- {\tt genericProjection(n, I)}
 
--- {\tt genericProjection(n, R)},
-
--- {\tt genericProjection(I)},
-
--- {\tt genericProjection(R)}
+-- {\tt genericProjection(n, R)}
 
@@
 This can be faster and can be useful for finding points.
+The same approach was used in the {\tt point} command in the {\tt Cremona} package \cite{CremonaSource}.
 If we already know the codimension is {\tt c}, we can set {\tt Codimension=>c} so the function does not compute it.
-The same approach was used in the {\tt point} command in the {\tt Cremona} package \cite{CremonaSource}.
 
@@
 \begin{example}
-    Consider the following example (which is $(R_1)$) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
+    Consider the following 3-dimensional example (which is $(R_1)$, regular in codimension 1) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
 However, we can use this function to quickly find interesting minors.

### 2021-07-17 FastLinAlg/Paper/RandomRationalPoints.tex
Minor changes to the paper.

@@
     \item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
+    \item {\tt dimViaBezout}:  This tries to compute the dimension of an algebraic set by intersecting with hyperplanes.  (\Cref{subsec.DimViaBezout})
     \item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
@@
 
-\section{Other functions {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
+\section{Useful functions {\tt dimViaBezout} and {\tt randomCoordinateChange}}{\label{helper}}
 
-\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, is a significant bottleneck.
+\subsection{\tt dimViaBezout:} 
+\label{subsec.DimViaBezout}
+We thank the Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal, is a significant bottleneck.
 While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
@@
     \item {\tt Homogeneous}: 
-    If the ideal is homogeneous, we can use homogeneous linear spaces to compute dimension.
-However, in practice, this appears slower than the non-homogeneous strategy due to the fact that we must saturate the result.
-Hence this is turned off by default.
+    If the ideal is homogeneous, we can use homogeneous linear spaces to compute dimension. Sometimes this is faster and other times slower.
         %By default, the function uses one of two algorithms depending on whether the ideal is homogeneous.
@@
 	}}
-As you can see doing a single intersection attempt is about three times faster, and it usually gives the right answer (more than 99\% of the time in this particular example, but in others doing the computation in triplicate avoids returning incorrect answers).
+As you can see doing a single intersection attempt is about three times faster, and it usually gives the right answer (far more than 99\% of the time in this particular example, but in others doing the computation in triplicate avoids returning incorrect answers).
 \end{example}
 
-\subsection*{\tt randomCoordinateChange:}
+\subsection{\tt randomCoordinateChange:} 
+\label{subSec.randomCoordinateChange}
 
-This function takes a polynomial ring as an input and outputs the coordinate change map,
+This function takes a polynomial ring as an input and outputs a coordinate change map,
 i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.
 
-In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.
+In some applications, a full random change of coordinates is not desired, as it might cause code to run very slowly.
 A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.

### 2021-07-17 FastLinAlg/Paper/RandomRationalPoints.tex
Minor changes to the paper.

@@
 Let $X:=V(I)$ denote the corresponding set of rational points in affine $n$-space.
-Finding one such rational point or geometric point (geometric meaning a point in some finite field extension), in an algorithmically efficient manner is our primary motivation.
+Finding one such rational point or geometric point (geometric meaning a point over some finite field extension), in an algorithmically efficient manner is our primary motivation for this package.
+The authors of the package are Sankhaneel Bisui, Zhan Jiang, Sarasij Maitra, Th\'ai Th\`anh Nguy$\tilde{\text{\^e}}$n, Frank-Olaf Schreyer, and Karl Schwede.
 

### 2021-07-19 FastLinAlg/Paper/RandomRationalPoints.tex
Updated abstract.

@@
   We present {\tt RandomPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field.
-We provide different strategies such as linear intersection and projection to obtain such points.
+We provide tools to estimate the dimension of a variety.
 We also present methods to obtain non-vanishing minors of a given size in a given matrix, by evaluating the matrix at a point.
@@
 
+\section{The latest version}
+
+The latest version of the package is available in the {\tt FastLinAlg} branch of
+
+\begin{center}
+    {\scriptsize\tt\url{https://github.com/Macaulay2/Workshop-2020-Cleveland/blob/FastLinAlg/FastLinAlg/M2/RandomPoints.m2}}
+\end{center}
+
 \bibliographystyle{alpha}

### 2021-08-18 FastLinAlg/Paper/RandomRationalPoints.tex
Made requested changes to the paper.

@@
 \begin{document}
-\title{{RandomPoints} package for \emph{Macaulay2}}
+\title{Finding points on varieties with \emph{Macaulay2}}
 \author{Sankhaneel Bisui}
@@
 Let $X:=V(I)$ denote the corresponding set of rational points in affine $n$-space.
-Finding one such rational point or geometric point (geometric meaning a point over some finite field extension), in an algorithmically efficient manner is our primary motivation for this package.
+Finding one such rational point or geometric point (geometric meaning a point over some finite field extension), in an algorithmically efficient manner, is our primary motivation for this package.
 The authors of the package are Sankhaneel Bisui, Zhan Jiang, Sarasij Maitra, Th\'ai Th\`anh Nguy$\tilde{\text{\^e}}$n, Frank-Olaf Schreyer, and Karl Schwede.
 
-    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.
-We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can built-in the function {\tt randomKRationalPoint} \cite{M2}, our methods frequently appear to be faster and apply in the affine setting as well.
+    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it the point we find is not rational.
+We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2}.
+Our methods frequently appear to be faster and apply in the affine setting as well.
 
-    We develop functions that apply different strategies to generate random rational and geometric points on the given variety.
+    We develop functions that apply various strategies to generate random rational and geometric points on the given variety.
 We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.
@@
     \item {\tt findANonZeroMinor} and {\tt extendIdealByNonZeroMinor}:  The first function finds a submatrix of a given matrix that is nonsingular at a point of a given ideal.
-The second, adds said submatrix to an ideal, which is useful for computing partial Jacobian ideals.
+The second adds said submatrix to an ideal, which is useful for computing partial Jacobian ideals.
 (\Cref{findANonZeroMinor})
@@
     \vspace{0.5em}
-    \noindent where $n$ is a positive integer denoting the number of desired points, and 
+    \noindent where $n$ is a positive integer denoting the number of points desired, and 
     $I$ is an ideal inside a polynomial ring.  If {$ n$} is omitted, then it is assumed to be 1.
@@
 
-    The user may also choose to provide some additional information depending on the context which may help in faster computations, or whether a point is found at all.
+    The user may also choose to provide some additional information, which
+    may accelerate the computation and improve the probability that a point is found.
 
@@
     \begin{itemize}
-    \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.
-It begins begins with brute force and moves to linear intersections with particularly simple linear forms, and gradually ramps up the randomness.
+    \item {\tt Default} performs the strategies below in a sequence, with the aim of finding a point quickly.
+It begins begins with brute force and moves to linear intersections with particularly simple linear forms, and gradually ramps up the randomness of the linear forms.
     \item {\tt BruteForce} simply tries random points and sees if they are on the variety.
@@
 
-    Notice that the speed, or success, varies depending on the strategy (see also \Cref{helper}).
+    The speed and the probability of success depend on the strategy (see also \Cref{helper}).
 
@@
     Intersection with a general linear space will naturally find scheme theoretic points that are not rational over the base field.  
-    Setting {\tt ExtendField => true} will tell the function that such points are valid.  Setting {\tt ExtendField => false} will tell the function ignore such points.  
+    Setting {\tt ExtendField => true} will tell the function that such points are valid.  Setting {\tt ExtendField => false} will tell the function to ignore such points.  
     In fact, {\tt ExtendField => true} will also tell Macaulay2 to use linear spaces defined over a field extension, which can improve randomness properties.
@@
 In a number of existing algorithms, one first does a generic (or even not very generic) projection $h : \bA^n \to \bA^m$ and so that $h(X)$ is a hypersurface (at least set theoretically).
-Then one finds a point $x \in h(X)$ (say as above), and computes $h^{-1}(\{ x \})$ which is a linear space in $\bA^n$ that typically intersects $X$ in a rational point.
+Then one finds a point $x \in h(X)$ (say as above), and computes $h^{-1}(\{ x \})$, which is a linear space in $\bA^n$ that typically intersects $X$ in a rational point.
 For example, this is done in {\tt randomKRationalPoint} in core Macaulay2.
-Note that projecting to a hypersurface \emph{still is performing an intersection with a linear space}, since $h^{-1}(\{ x \})$ is linear, but it tries to choose the linear space intelligently.
+Note that projecting to a hypersurface \emph{still is intersecting with a linear space}, since $h^{-1}(\{ x \})$ is linear, but it tries to choose the linear space intelligently.
 
@@
 
-    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (i.e., with many linear or binomial terms).
+    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy we initially try linear spaces  whose defining equations have as few
+    terms as possible   For example, in a ring with 10 variables, we first try binomial linear forms like
+    \[ 
+        -27 x_2 + 38 x_7
+    \]
+        instead of a random linear form like
+    \[
+        -28x_1-27x_2+29x_3+27x_4-28x_5+27x_6+38x_7-13x_8+21x_9-3x_{10}.
+    \]
     Such simple linear spaces are the ones implicitly considered in {\tt randomKRationalPoint} for instance since that generic projection is so simple.
@@
 \label{subsec.DimViaBezout}
-We thank the Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal, is a significant bottleneck.
+We thank the Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal is a significant bottleneck.
 While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
-This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.
-For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.
+This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with random linear spaces of increasing dimension until there is an intersection.
+For example, if the intersection of $V(I)$ with a random line has a point, then we expect that $V(I)$ contains a hypersurface.
 If there was no intersection, this function tries a $2$-dimensional linear space, and so on.
@@
 
-One can use this function to show that rings are $(R_1)$, that is, regular in codimension $1$.
+One can use this function to show that rings are regular in codimension 1, that is, satisfy Serre's condition (R1).
 
 \begin{example}
-    Consider the following 3-dimensional example (which is $(R_1)$, regular in codimension 1) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
+    Consider the following 3-dimensional example (which is regular in codimension 1) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
 However, we can use this function to quickly find interesting minors.
@@
 
-There is some similar functionality computing partial Jacobian ideals obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}, see \cite{FastLinAlgSource}.
+There is some similar functionality for computing partial Jacobian ideals obtained via heuristics (as opposed to actually finding rational or geometric points) in the package {\tt FastLinAlg}, see \cite{FastLinAlgSource}.
 That package now uses the functionality contained here in {\tt RandomPoints} in some of its functions.

### 2021-08-19 FastLinAlg/Paper/RandomRationalPoints.tex
Merged the changes of Sankhaneel and Thai

@@
 
-    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it the point we find is not rational.
-We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2}.
-Our methods frequently appear to be faster and apply in the affine setting as well.
+    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.
+We also note that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2} -- our methods frequently appear to be faster and apply in the affine setting as well.
 
@@
     \item[\tt NumThreadsToUse => ZZ]
+When calling  { \tt randomPoints} and functions that call it with a { \tt BruteForce} strategy, this options allows the user to specify the number of threads to use in brute force point checking.
 
-    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
     \end{description}
@@
 By default, generic projection is turned off in the default strategy, that is we set {\tt ProjectionAttempts=>0}.
-See \autoref{rem.ProjectingToAHypersurfaceFirst} for additional related discussion.
+See \autoref{rem.ProjectingToAHypersurfaceFirst}  l related discussion.
 
@@
 \label{subsec.DimViaBezout}
-We thank the Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal is a significant bottleneck.
+We thank Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal is a significant bottleneck.
 While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
@@
 \begin{example}
-    Consider the following 3-dimensional example (which is regular in codimension 1) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.
-However, we can use this function to quickly find interesting minors.
+    Consider the following 3-dimensional example which is regular in codimension 1.
+Note, in this example, computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ coming from the associated $7 \times 12$ Jacobian matrix.
+However, we can use our function to quickly find interesting minors.
   {{\small\color{blue}

### 2021-08-19 FastLinAlg/Paper/RandomRationalPoints.tex
clarifications, almost done with the package documentation changes

@@
 
-    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.
+    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more random rational or geometric points on a variety quickly.
 We also note that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2} -- our methods frequently appear to be faster and apply in the affine setting as well.

### 2023-05-17 FastLinAlg/Paper/RandomRationalPoints.tex
updated addresses etc

@@
 \author{Sankhaneel Bisui}
-\email{sbisui@tulane.edu}
-\address{Department of Mathematics, Tulane University, New Orleans, LA 70118}
+\email{Sankhaneel.Bisui@umanitoba.ca}
+\address{Department of Mathematics, University of Manitoba, Winnipeg, MB R3T 2M8}
 \author{Zhan Jiang}
 \email{zoeng@umich.edu}
-\address{Department of Mathematics, University of Michigan, Ann Arbor, MI 48109}
+\address{Google, Mountain View, CA}
 \author{Sarasij Maitra}
-\email{sm3vg@virginia.edu}
-\address{Department of Mathematics, University of Virginia, Charlottesville, VA 22904}
+\email{maitra@math.utah.edu}
+\address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
 \author{Th\'ai Th\`anh Nguy$\tilde{\text{\^e}}$n}
-\email{tnguyen11@tulane.edu}
-\address{Department of Mathematics, Tulane University, New Orleans, LA 70118}
+\email{nguyt161@mcmaster.ca}
+\address{Department of Mathematics, McMaster University, Hamilton, Ontario, L8S 4L8}
 \author{Karl Schwede}
@@
 
-\thanks{Schwede was supported by NSF Grants \#1801849, FRG \#1952522 and a Fellowship from the Simons Foundation.}
+\thanks{Schwede was supported by NSF Grants \#1801849, \#2101800, FRG \#1952522 and a Fellowship from the Simons Foundation.}
 
